### 2016-01-15 maths.tex
added tex file for derivations and expressions

\documentclass[12pt]{article}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage[top=20mm, bottom=20mm, left=30mm, right=30mm]{geometry}
\usepackage{soul}
%Gummi|061|=)
\usepackage{setspace}

\doublespacing

\title{\textbf{Derivations and Equations}}
\date{}
\begin{document}

\maketitle

\section{Deriving the bound} % (fold)
\label{sec:derive_bound}
The standard variational formulation is
%
\begin{align}
    \log p(y) &= \int q(z)\log\frac{p(y|z)p(z)}{q(z)}\frac{q(z)}{p(z|y)} dz\notag\\
    &= \underbrace{\mathbb{E}_{q(z)}\big[ \log p(y|z)p(z) - \log q(z) \big]}_{\mathcal{L}_1} + \text{KL}\big[ q(z)\Vert p(z|y) \big]\label{eq:logpy1}
\end{align}
%
The GP-LVM-CMF variational joint in the augmented latent space is:
%
\begin{equation}
    q(z,f,u,X) = \prod_{i=1}^N \{q(z_i|f(X_i))q(f(X_i)|u)\}q(u)q(X)
\end{equation}
%
which induces an intractable variational marginal
%
\begin{equation}
    q(z) = \int \prod_{i=1}^N \{q(z_i|f(X_i))q(f(X_i)|u)\}q(u)q(X) df du dX.
\end{equation}
%
However we can apply the variational formulation again, noting that $q(z) = \frac{q(z|f,u,X)q(f,u,X)}{q(f,u,X|z)}$, and integrating \ref{eq:logpy1} wrt the auxiliary variables, to obtain a further lower bound:
\begin{align}
    \log p(y) &= \int q(z|f,u,X)q(f,u,X)\log \frac{p(y|z)p(z)r(f,u,X|z)}{q(z|f,u,X)q(f,u,X)}\frac{q(f,u,X|z)}{r(f,u,X|z)}\frac{q(z)}{p(z|y)} df du dX dz\notag\\
    &= \mathbb{E}_{q(z,f,u,X)}\big[ \log p(y|z)p(z)r(f,u,X|z) - \log q(z|f,X)q(f,u,X) \big]\label{eq:L_aux}\\
    &\qquad+ \mathbb{E}_q(z)\big[ \text{KL}[q(f,u,X|z)\Vert r(f,u,X|z)] \big] + \text{KL}\big[ q(z)\Vert p(z|y) \big]
\end{align}
%
in which the new auxiliary lower bound, $\mathcal{L}_{aux}$ is given by the expression \ref{eq:L_aux}, and where we have introduced the auxiliary distribution $r(f,u,X|z)$ which serves to approximate the variational posterior, $q(f,u,X|z)$, of the auxiliary variables conditioned on the latent variables.\\
We may re-express $\mathcal{L}_{aux}$ in a way which makes use of the analytical expression for the K-L divergence between two Gaussians, $q(f,u,x)$ and $r(f,u,X|z)$ and, in the case that the prior of the generative model, $p(z)$, is also Gaussian distributed - as is the case for the continuous latent variable MLP model we'll consider first - then the bound contains a second Guassian KL term:
%
\begin{equation}
    \mathcal{L}_{aux} = \underbrace{\mathbb{E}_{q(z)}\big[ \log p(y|z) \big]}_{A} - \underbrace{\mathbb{E}_{q(f,u,X)}\big[ \text{KL}[q(z|f,X)\Vert p(z)] \big]}_{B} - \underbrace{\mathbb{E}_{q(z)}\big[ \text{KL}[q(f,u,X)\Vert r(f,u,X|z)] \big]}_{C}.
\end{equation}
%
%
\section{Expressions}
\subsection{Generative model}
%

\subsection{Variational model}
%
\begin{align}
    q(X) &= \prod_{i=1}^N \mathcal{N}(x_{i,:}\mid 0, I_R) = \prod_{i=1}^N \prod_{r=1}^R \mathcal{N}(x_{i,r}\mid 0,1)\\
    q(u) &= \prod_{m=1}^M \mathcal{N}(u_{:,m}\mid 0, K_{X_uX_u})\\
    q(f(X)|u) &= \mathcal{N}(f(X)\mid K_{XX_u}K_{X_uX_u}^{-1}u, K_{XX}-K_{XX_u}K_{X_uX_u}^{-1}K_{X_uX})\\
    q(z|f(X)) &= \prod_{i=1}^N \mathcal{N}(z_i\mid f(X_i), \sigma^2)
\end{align}

\subsection{Auxiliary model}
%
\begin{align}
    r(f,u,X|z) &= q(f(X)|u,X)r(u,X|z)\\
    r(u,X|z) &=
\end{align}

\end{document}

### 2016-01-16 maths.tex
slight changes to python code. Still needs renaming of variables to match with tex doc.

@@
 in which the new auxiliary lower bound, $\mathcal{L}_{aux}$ is given by the expression \ref{eq:L_aux}, and where we have introduced the auxiliary distribution $r(f,u,X|z)$ which serves to approximate the variational posterior, $q(f,u,X|z)$, of the auxiliary variables conditioned on the latent variables.\\
-We may re-express $\mathcal{L}_{aux}$ in a way which makes use of the analytical expression for the K-L divergence between two Gaussians, $q(f,u,x)$ and $r(f,u,X|z)$ and, in the case that the prior of the generative model, $p(z)$, is also Gaussian distributed - as is the case for the continuous latent variable MLP model we'll consider first - then the bound contains a second Guassian KL term:
+We may re-express $\mathcal{L}_{aux}$ in a way which makes use of the analytical expression for the K-L divergence between two Gaussians, $q(f,u,x)$ and $r(f,u,X|z)$ and, in the case that the prior of the generative model, $p(z)$, is also Gaussian distributed - as is the case for the continuous latent variable MLP model we'll consider first - then the bound contains a second Gaussian KL term:
 %

### 2016-01-16 maths.tex
update to maths

--- 2016-01-15 maths.tex
@@
 \usepackage{soul}
+\usepackage{bbm}
 %Gummi|061|=)
 \usepackage{setspace}
+\usepackage[usenames, dvipsnames]{color}
 
+\newcommand{\chris}[1]{\textcolor{ForestGreen}{#1}} 
+\newcommand{\Tau}{\mathcal{T}}
+\newcommand{\Kff}{K_{X_fX_f}}
+\newcommand{\Kuu}{K_{X_uX_u}}
+\newcommand{\Kuf}{K_{X_uX_f}}
+\newcommand{\Kfu}{K_{X_fX_u}}
+\newcommand{\Ex}{\mathbb{E}}
+\newcommand{\No}{\mathcal{N}}
+\newcommand{\chol}{\mathrm{chol}}
 \doublespacing
@@
 \begin{align}
-    q(X) &= \prod_{i=1}^N \mathcal{N}(x_{i,:}\mid 0, I_R) = \prod_{i=1}^N \prod_{r=1}^R \mathcal{N}(x_{i,r}\mid 0,1)\\
-    q(u) &= \prod_{m=1}^M \mathcal{N}(u_{:,m}\mid 0, K_{X_uX_u})\\
-    q(f(X)|u) &= \mathcal{N}(f(X)\mid K_{XX_u}K_{X_uX_u}^{-1}u, K_{XX}-K_{XX_u}K_{X_uX_u}^{-1}K_{X_uX})\\
+q(X) &= \prod_{i=1}^N \mathcal{N}(x_{:,n}\mid 0, I_R) = \prod_{n=1}^N \mathcal{N}(x_{n,:}\mid 0,1)
+     \\
+    q(u) &= \prod_{m=1}^M \mathcal{N}(u_{:,m}\mid \kappa_{:,m}, \Kuu)\\
+    q(f(X)|u) &= \mathcal{N}(f(X)\mid \Kfu\Kuu^{-1}u, \Kff-\Kuf\Kuu^{-1}\Kuf)\\
     q(z|f(X)) &= \prod_{i=1}^N \mathcal{N}(z_i\mid f(X_i), \sigma^2)
+\end{align}
+
+\chris{I think the variational distributions need to have variational parameters. Also there is a single guassian process, right? If so q(u) is just one gaussian, not a product of gaussians}
+\begin{align}
+    q(X) &= \prod_{i=1}^N \mathcal{N}(X_i\mid \lambda_i, \Lambda_i)\\
+    q(u) &= \mathcal{N}(u \mid \kappa, K_{X_uX_u})\\
 \end{align}
@@
     r(f,u,X|z) &= q(f(X)|u,X)r(u,X|z)\\
-    r(u,X|z) &=
+    r(u,X|z) &= r(u| z) \prod_i r(X_i|z) = \mathcal{N}(u; \upsilon, \Upsilon ) \prod_i \mathcal{N}( X_i, \tau_i, \Tau_i )
 \end{align}
 
+\subsection{}
+
+\begin{align}
+\Ex_q(z) \mathcal{L}(z) &= \Ex_q(z,f,u,X) \mathcal{L}(z,f,u,X) \\
+&= \Ex_{\No(\eta;0,1)\No(\xi;0,1)\No(\alpha;0,1)\prod\No(\beta_i,0,1)} \mathcal{L}((f,\sigma^2),(\mu,\Sigma)(\upsilon,\Upsilon),(\tau,\Tau))
+\end{align}
+
+where  $z = f + \mathbf{I}\sigma\eta $
+
+where $\Sigma = \Kff - \Kfu\Kuu^{-1}\Kuf$
+
+where $\mu   = \Kfu\Kuu^{-1}u$
+
+where $f = \mu + \chol(\Sigma)\xi $
+
+where $u = \upsilon + \chol(\Upsilon)\upsilon$
+
+where $X_i = \tau_i + \chol(\Tau_i)\beta_i$
+
 \end{document}

### 2016-01-16 GP-LVM-CMF.py
various hacking

@@ -5,26 +5,34 @@ import theano as th
 import theano.tensor as T
 from theano.tensor import slinalg, nlinalg
 
-eps_y = 1e-4
-class kernel:
-    def RBF(self, sf2, l, X1, X2 = None):
-        _X2 = X1 if X2 is None else X2
-        dist = ((X1 / l)**2).sum(1)[:, None] + ((_X2 / l)**2).sum(1)[None, :] - 2*(X1 / l).dot((_X2 / l).T)
-        RBF = sf2 * T.exp(-dist / 2.0)
-        return (RBF + eps_y * T.eye(X1.shape[0])) if X2 is None else RBF
-    def RBFnn(self, sf2, l, X):
-        return sf2 + eps_y
-    def LIN(self, sl2, X1, X2 = None):
-        _X2 = X1 if X2 is None else X2
-        LIN = sl2 * (X1.dot(_X2.T) + 1)
-        return (LIN + eps_y * T.eye(X1.shape[0])) if X2 is None else LIN
-    def LINnn(self, sl2, X):
-        return sl2 * (T.sum(X**2, 1) + 1) + eps_y
 
+class kernelFactory:
+    def __init__(self, kernelType_, eps_=1e-4):
+        self.kernelType = kernelType_         
+        self.eps        = eps_
+    
+    def kernel( self, X1, X2, theta, name_ ):
+        _X2 = X1 if X2 is None else X2
+        if self.kernelType == 'RBF':
+            dist = ((X1 / theta[0])**2).sum(1)[:, None] + ((_X2 / theta[0])**2).sum(1)[None, :] - 2*(X1 / theta[0]).dot((_X2 / theta[0]).T)
+            K = theta[1] * T.exp(-dist / 2.0)
+            K (K + self.eps * T.eye(X1.shape[0])) if X2 is None else K
+            K.name = name + '(RBF)'
+        elif self.kernelType == 'RBFnn':
+            K = theta[0] + self.eps
+            K.name = name + '(RBFnn)'            
+        elif self.kernelType ==  'LIN':
+            K = theta[0] * (X1.dot(_X2.T) + 1)
+            (K + self.eps_y * T.eye(X1.shape[0])) if X2 is None else K
+            K.name = name + '(LIN)'
+        elif self.kernelType ==  'LINnn':
+            K * (T.sum(X1**2, 1) + 1) + self.eps
+            K.name = name + '(LINnn)'
+        return K
 
[…]

### 2016-01-16 GP-LVM-CMF.py
more hacking

@@ -32,54 +32,89 @@ class kernelFactory:
 
 class SGPDV:
 
-    def __init__( self, N_, M_, dimX_, learningRate_, batchSize_, kernelType_='RBF', theta_init, sigma_init ):
+    def __init__( self, dataSize, induceSize, batchSize, dimLatent, learningRate, theta_init, sigma_init, kernelType_='RBF' ):
 
-        self.N             = N_# number of observations
-        self.M             = M_# Number of inducing ponts
-        self.dimX          = dimX_ # Dimensionality of the latent co-ordinates
-
-        self.lowerBound    = -np.inf       # Lower bound
-        self.learningRate  = learningRate_ # Learning rate
-        self.batchSize     = batchSize_    #
-        self.kernelType    = kernelType_
-
-        # Numpy values
-        upsilon_np    = [0]*self.N # mean of r(u|z)
-        Upsilon_np    = [0]*self.N # variance of r(u|z)
-        tau_np        = np.zeros((self.dimX,1))  # mean of r(X|z)
-        Tau_np      = np.eye(self.dimX)        # variance of r(X|z)
-        phi_np[i]  = np.zeros((self.dimX,1))  # mean of q(X)
-        Phi_np[i]  = np.eye(self.dimX)        # variance of q(X)
-        Xu_np[i]   = np.zeros((self.dimX,1))  # These are the locations of the inducing points
-
-        sigma = T.sahred( np.array( sigma_init )) # standard deviation of q(z|f)
-        theta = np.array( theta_init) # kernel parameters
+        self.N = dataSize   # number of observations
+        self.M = induceSize # Number of inducing ponts
+        self.B = batchSize    #
+        self.R = dimLatent # Dimensionality of the latent co-ordinates
         
-        # Theano variables
-        self.upsilon = T.shared( upsilon_np ) # mean of r(u|z)
-        self.Upsilon = T.shared( Upsilon_np )# variance of r(u|z)
-        self.tau     = [0]*self.N # mean of r(X|z)
-        self.Tau     = [0]*self.N # variance of r(X|z)
-        self.phi     = [0]*self.N # mean of q(X)
-        self.Phi     = [0]*self.N # variance of q(X)
-        for i in range( self.N ):
[…]

### 2016-01-17 GP-LVM-CMF.py
inor mods

@@ -85,16 +85,14 @@ class SGPDV:
         self.xi.name    = 'xi'
     
         # Latent co-ordinates
-        self.Xu = T.shared( M_R_vec )
+        self.Xu = T.dmatrix( M_R_vec )
         self.Xf = T.shared( B_R_vec )
         self.Xu.name = 'Xu'
         self.Xf.name = 'Xf'
 
         self.currentBatch = T.ivector( B_vec )
         cPhi = slinalg.cholesky( self.Phi )
-        for n in range( self.B ):
-            i = self.currentBatch[n]
-            self.Xf[n,:] = self.phi[i,:] + ( cPhi  * self.beta[n,:].T ).T
+        self.Xf = self.phi[self.currentBatch,:] + ( cPhi  * self.beta.T ).T
         
         # Kernels
         kfactory = kernelFactory( self.kernelType )
@@ -113,15 +111,14 @@ class SGPDV:
         self.f  = T.shared( Q_B_mat )
         self.mu = T.shared( Q_B_mat )
 
-        for i in range( self.Q ):
-            # Sample u_q from q(u) = N(u_q; kappa_q, Kuu )
-            self.u[i,:]  = self.kappa[i,:] + cKuu * self.alpha
-            # compute mean of f            
-            self.mu[i,:] = self.Kfu * ( iKuu * self.u[i,:].T ).T
-            # Sample f from q(f|u,X) = N( mu_q, Sigma )            
-            self.f[i,:]  = self.mu + ( cSigma * self.xi[i,:].T ).T        
-            # Sample z from q(z|f) = N(z,f,I*sigma^2)            
-            self.z[i,:]  = self.f + self.sigma * self.eta[i,:]
+        # Sample u_q from q(u) = N(u_q; kappa_q, Kuu )
+        self.u  = self.kappa[i,:] + cKuu * self.alpha
+        # compute mean of f            
+        self.mu = self.Kfu * ( iKuu * self.u.T ).T
+        # Sample f from q(f|u,X) = N( mu_q, Sigma )            
+        self.f  = self.mu + ( cSigma * self.xi.T ).T        
+        # Sample z from q(z|f) = N(z,f,I*sigma^2)            
+        self.z  = self.f + self.sigma * self.eta[i,:]
         
         self.u.name     = 'u'
         self.Sigma.name = 'Sigma'
@@ -131,7 +128,6 @@ class SGPDV:
             
         # TODO add more stuff to this list
[…]

### 2016-01-17 maths.tex
math changes
Added generative model to tex doc

@@
 \begin{equation}
-    \mathcal{L}_{aux} = \underbrace{\mathbb{E}_{q(z)}\big[ \log p(y|z) \big]}_{A} - \underbrace{\mathbb{E}_{q(f,u,X)}\big[ \text{KL}[q(z|f,X)\Vert p(z)] \big]}_{B} - \underbrace{\mathbb{E}_{q(z)}\big[ \text{KL}[q(f,u,X)\Vert r(f,u,X|z)] \big]}_{C}.
+    \mathcal{L}_{aux} = \underbrace{\mathbb{E}_{q(z)}\big[ \log p(y|z) \big]}_{A} - \underbrace{\mathbb{E}_{q(f,u,X)}\big[ \text{KL}[q(z|f,X)\Vert p(z)] \big]}_{B} - \underbrace{\mathbb{E}_{q(z)}\big[ \text{KL}[q(f,u,X)\Vert r(f,u,X|z)] \big]}_{C}.\label{eq:L_aux_ABC}
 \end{equation}
@@
 %
-
+\begin{equation}
+    p(z) = \mathcal{N}(z\mid 0, I)
+\end{equation}
+%
+\subsubsection{Continuous data $\implies$ Gaussian MLP likelihood}
+%
+\begin{align}
+    \log p(y|z) &= \log\mathcal{N}(x\mid \mu, \sigma^2 I)\\
+    \text{where } \mu &= W_2 h + b_2\\
+    \log\sigma^2 &= W_3 h + b_3\\
+    h &= \tanh(W_1 z + b_1)
+\end{align}
+Term $B$ in (\ref{eq:L_aux_ABC}) then becomes:
+%
+\begin{align}
+    &-\frac{1}{2}\sum_{i=1}^N \mathbb{E}_{q(f,u,X)}\big[ 1 + \log\sigma^2 - f(X_i)^2 - \sigma^2 \big]\\
+    &= -\frac{1}{2}\sum_{i=1}^N \mathbb{E}_{q(u)q(X)}\big[ 1 + \log\sigma^2 - (K_{f_iu}K_{uu}^{-1}u)^2 - k_{f_if_i} + K_{f_iu}K_{uu}^{-1}K_{uf_i}-\sigma^2 \big]\\
+    &= -\frac{N}{2}(1 + \log\sigma^2 - \sigma^2) + \frac{1}{2}\sum_{i=1}^N\mathbb{E}_{q(X)}\big[ K_{f_iu}K_{uu}^{-1}\big( \kappa\kappa^\top + K_{uu} \big)K_{uu}^{-1}K_{uf_i} + K_{f_if_i} - K_{f_iu}K_{uu}^{-1}K_{uf_i} \big]\\
+    %&= -\frac{N}{2}(1 + \log\sigma^2 - \sigma^2) + \frac{1}{2}\psi_0 - \frac{1}{2}\text{tr}\big( K_{uu}^{-1}\Psi_2 \big) + \frac{1}{2}\text{tr}\big( K_{uu}^{-1}(\kappa\kappa^\top + K_{uu})K_{uu}^{-1}\Psi_2 \big)
+\end{align}
+%
+%where
+%\begin{align}
+%    \psi_0 &= \text{tr}\big( \langle K_{ff}\rangle_{q(X)} \big)\\
+%    \Psi_2 &= \langle K_{uf}K_{fu}\rangle_{q(X)}
+%\end{align}
+%
+\subsubsection{Discrete data $\implies$ Bernoulli MLP likelihood}
+%
+\begin{align}
+    \log p(y|z) &= \sum_{p=1}^P y_p\log \hat{z}_p + (1-y_p)\log (1-\hat{z}_p)\\
+    \text{where } \hat{z} &= f_\sigma(W_2 h + b2)\\
+     h &= \tanh(W_1 z + b_1).
+\end{align}
 \subsection{Variational model}
@@
 \begin{align}
-    q(X) &= \prod_{i=1}^N \mathcal{N}(X_i\mid \lambda_i, \Lambda_i)\\
+    q(X) &= \prod_{i=1}^N \mathcal{N}(X_i\mid \phi_i, \Phi_i)\\
     q(u) &= \mathcal{N}(u \mid \kappa, K_{X_uX_u})\\
@@
 \Ex_q(z) \mathcal{L}(z) &= \Ex_q(z,f,u,X) \mathcal{L}(z,f,u,X) \\
-&= \Ex_{\No(\eta;0,1)\No(\xi;0,1)\No(\alpha;0,1)\prod\No(\beta_i,0,1)} \mathcal{L}((f,\sigma^2),(\mu,\Sigma)(\upsilon,\Upsilon),(\tau,\Tau))
+&= \Ex_{\No(\eta;0,1)\No(\xi;0,1)\No(\alpha;0,1)\prod\No(\beta_i,0,1)} \mathcal{L}_1(\eta,\xi,\alpha,\beta) \\
+\mathcal{L}_1(\eta,\xi,\alpha,\beta) &=  \mathcal{L}(z,f,u,X)
 \end{align}
@@
 
-where $u = \upsilon + \chol(\Upsilon)\upsilon$
+where $u = \upsilon + \chol(\Upsilon)\alpha$
 
-where $X_i = \tau_i + \chol(\Tau_i)\beta_i$
+where $X_i = \tau_i + \chol(\Tau_i)\beta_i$.
+
+Term $C$ in (\ref{eq:L_aux_ABC}) then becomes:
+%
+\begin{align}
+    &\mathbb{E}_{q(z)}\big[ \text{KL}[q(u)\Vert r(u|z)] \big] = \mathbb{E}_{q(z)}\big[ \text{KL}[\prod_{i=1}^Nq(X_i)\Vert \prod_{i=1}^Nr(X_i|z)] \big]\\
+    &= \frac{1}{2}\mathbb{E}_{q(z)}\big[ (\upsilon - \kappa)^\top\Upsilon^{-1}(\upsilon - \kappa) + \text{tr}\big( \Upsilon^{-1}K_{uu} + \log\Upsilon - \log K_{uu} \big) \big] - MQ\\
+    &\frac{1}{2}\sum_{i=1}^N \big\{ \mathbb{E}_{q(z)}\big[ (\phi_i - \tau_i)^\top\Tau_i^{-1}(\phi_i - \tau_i) + \text{tr}\big( \Tau_i^{-1}\Phi_i + \log\Tau_i - \log\Phi_i \big) \big] \big\} - NR.
+\end{align}
 

### 2016-01-17 maths.tex
Completed all log-likelihood expressions, minor changes to tex doc.

@@
     &\mathbb{E}_{q(z)}\big[ \text{KL}[q(u)\Vert r(u|z)] \big] = \mathbb{E}_{q(z)}\big[ \text{KL}[\prod_{i=1}^Nq(X_i)\Vert \prod_{i=1}^Nr(X_i|z)] \big]\\
-    &= \frac{1}{2}\mathbb{E}_{q(z)}\big[ (\upsilon - \kappa)^\top\Upsilon^{-1}(\upsilon - \kappa) + \text{tr}\big( \Upsilon^{-1}K_{uu} + \log\Upsilon - \log K_{uu} \big) \big] - MQ\\
-    &\frac{1}{2}\sum_{i=1}^N \big\{ \mathbb{E}_{q(z)}\big[ (\phi_i - \tau_i)^\top\Tau_i^{-1}(\phi_i - \tau_i) + \text{tr}\big( \Tau_i^{-1}\Phi_i + \log\Tau_i - \log\Phi_i \big) \big] \big\} - NR.
+    &= \frac{1}{2}\mathbb{E}_{q(z)}\big[ (\upsilon - \kappa)^\top\Upsilon^{-1}(\upsilon - \kappa) + \text{tr}\big( \Upsilon^{-1}K_{uu} + \log\Upsilon - \log K_{uu} \big) \big] - \frac{MQ}{2}\\
+    &\frac{1}{2}\sum_{i=1}^N \big\{ \mathbb{E}_{q(z)}\big[ (\phi_i - \tau_i)^\top\Tau_i^{-1}(\phi_i - \tau_i) + \text{tr}\big( \Tau_i^{-1}\Phi_i + \log\Tau_i - \log\Phi_i \big) \big] \big\} - \frac{NR}{2}.
 \end{align}

### 2016-01-17 GP-LVM-CMF.py

@@ -100,26 +100,26 @@ class SGPDV:
         self.Kff = kfactory.kernel( self.Xf, self.Xf, self.theta, 'Kff' )
         self.Kfu = kfactory.kernel( self.Xf, self.Xu, self.theta, 'Kfu' )
 
-        cKuu = slinalg.cholesky( self.Kuu )
-        iKuu = nlinalg.matrix_inverse( self.Kuu )
+        self.cKuu = slinalg.cholesky( self.Kuu )
+        self.iKuu = nlinalg.matrix_inverse( self.Kuu )
 
         # Variational distribution
-        self.Sigma = self.Kff - self.Kfu * iKuu * self.Kfu.T
-        cSigma = slinalg.cholesky( Sigma )
+        self.Sigma = self.Kff - T.dot(self.Kfu, T.dot(self.iKuu, T.dot(self.Kfu.T)))
+        Sigma = slinalg.cholesky( Sigma )
 
         self.u  = T.shared( Q_M_mat )
         self.f  = T.shared( Q_B_mat )
         self.mu = T.shared( Q_B_mat )
 
         # Sample u_q from q(u) = N(u_q; kappa_q, Kuu )
-        self.u  = self.kappa + ( cKuu * self.alpha.T ).T
-        # compute mean of f            
-        self.mu = self.Kfu * ( iKuu * self.u.T ).T
-        # Sample f from q(f|u,X) = N( mu_q, Sigma )            
-        self.f  = self.mu + ( cSigma * self.xi.T ).T 
-        # Sample z from q(z|f) = N(z,f,I*sigma^2)     
-        self.z  = self.f + ( self.sigma * self.eta.T ).T
-        
+        self.u  = self.kappa + ( T.dot(cKuu,self.alpha.T) ).T
+        # compute mean of f
+        self.mu = self.Kfu * ( T.dot(iKuu,self.u.T) ).T
+        # Sample f from q(f|u,X) = N( mu_q, Sigma )
+        self.f  = self.mu + ( T.dot(cSigma,self.xi.T) ).T
+        # Sample z from q(z|f) = N(z,f,I*sigma^2)
+        self.z  = self.f + ( T.dot(self.sigma,self.eta.T) ).T
+
         self.u.name     = 'u'
         self.Sigma.name = 'Sigma'
         self.mu.name    = 'mu'
@@ -138,27 +138,48 @@ class SGPDV:
     def L( self ):
 
[…]

### 2016-01-17 GP-LVM-CMF.py
Unneeded stuff deleted, randomise functions added

@@ -32,7 +32,7 @@ class kernelFactory:
 
 class SGPDV:
 
-    def __init__( self, dataSize, induceSize, batchSize, dimX, dimZ, learningRate, theta_init, sigma_init, kernelType_='RBF' ):
+    def __init__( self, dataSize, induceSize, batchSize, dimX, dimZ, theta_init, sigma_init, kernelType_='RBF' ):
 
         self.N = dataSize   # number of observations
         self.M = induceSize # Number of inducing ponts
@@ -41,7 +41,6 @@ class SGPDV:
         self.Q = dimZ
 
         self.lowerBound    = -np.inf      # Lower bound
-        self.learningRate  = learningRate # Learning rate
         self.kernelType    = kernelType
         self.p_z_gaussian  = True
 
@@ -53,6 +52,7 @@ class SGPDV:
         Q_B_mat = np.zeros((self.Q,self.B))
         M_M_mat = np.zeros((self.M,self.M))
         B_vec   = np.zeros((self.B,1), dtype=np.int32 )
+        
         # variational and auxilery parameters
         self.upsilon = T.shared( Q_M_mat ) # mean of r(u|z)
         self.Upsilon = T.shared( M_M_mat ) # variance of r(u|z)
@@ -65,7 +65,7 @@ class SGPDV:
         self.Upsilon.name = 'Upsilon'
         self.tau.name     = 'tau'
         self.Tau.name     = 'Tau'
-        self.phi.name     = 'tau'
+        self.phi.name     = 'phi'
         self.Phi.name     = 'Phi'
 
         # Other parameters
@@ -126,90 +126,132 @@ class SGPDV:
         self.f.name     = 'f'
         self.z.name     = 'z'
 
-        # TODO add more stuff to this list
-        self.gradientVariables = [ self.theta, self.sigma, self.upsilon, self.Upsilon ]
+        # This should be all the shared variables
+        self.gradientVariables = [ self.theta, self.sigma, self.phi, self.Phi, self.kappa, self.upsilon, self.Upsilon, self.tau, self.Tau ]
 
+    def randomise( self, sig ):
+        
+        upsilon = np.random.normal( 0, sig, (self.Q, self.M) )
+        Upsilon = np.random.normal( 0, sig, (self.M, self.M) )
+        tau     = np.random.normal( 0, sig, (self.N, self.R) )
+        Tau     = np.random.normal( 0, sig, (self.R, self.R) )
[…]

### 2016-02-02 icml2016/paper.tex
First commit of the working paper, along with the rest of the directory it sits in that includes the style file and example.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%% ICML 2016 EXAMPLE LATEX SUBMISSION FILE %%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% Use the following line _only_ if you're still using LaTeX 2.09.
%\documentstyle[icml2016,epsf,natbib]{article}
% If you rely on Latex2e packages, like most moden people use this:
\documentclass{article}

% use Times
\usepackage{times}
% For figures
\usepackage{graphicx} % more modern
%\usepackage{epsfig} % less modern
\usepackage{subfigure}

% For citations
\usepackage{natbib}

% For algorithms
\usepackage{algorithm}
\usepackage{algorithmic}

\usepackage{amsmath}
\usepackage{amssymb}

% As of 2011, we use the hyperref package to produce hyperlinks in the
% resulting PDF.  If this breaks your system, please commend out the
% following usepackage line and replace \usepackage{icml2016} with
% \usepackage[nohyperref]{icml2016} above.
\usepackage{hyperref}

% Packages hyperref and algorithmic misbehave sometimes.  We can fix
% this with the following command.
\newcommand{\theHalgorithm}{\arabic{algorithm}}

% Employ the following version of the ``usepackage'' statement for
% submitting the draft version of the paper for review.  This will set
% the note in the first column to ``Under review.  Do not distribute.''
\usepackage{icml2016}

% Employ this version of the ``usepackage'' statement after the paper has
% been accepted, when creating the final version.  This will set the
% note in the first column to ``Proceedings of the...''
%\usepackage[accepted]{icml2016}

% The \icmltitle you define below is probably too long as a header.
% Therefore, a short form for the running title is supplied here:
\icmltitlerunning{Submission and Formatting Instructions for ICML 2016}

\begin{document}

\twocolumn[
\icmltitle{Submission and Formatting Instructions for \\
           International Conference on Machine Learning (ICML 2016)}

% It is OKAY to include author information, even for blind
% submissions: the style file will automatically remove it for you
% unless you've provided the [accepted] option to the icml2016
% package.
\icmlauthor{Your Name}{email@yourdomain.edu}
\icmladdress{Your Fantastic Institute,
            314159 Pi St., Palo Alto, CA 94306 USA}
\icmlauthor{Your CoAuthor's Name}{email@coauthordomain.edu}
\icmladdress{Their Fantastic Institute,
            27182 Exp St., Toronto, ON M6H 2T1 CANADA}

% You may provide any keywords that you
% find helpful for describing your paper; these are used to populate
% the "keywords" metadata in the PDF but will not be shown in the document
\icmlkeywords{boring formatting information, machine learning, ICML}

\vskip 0.3in
]

\begin{abstract}
The purpose of this document is to provide both the basic paper template and
submission guidelines. Abstracts should be a single paragraph, between 4--6 sentences long, ideally.  Gross violations will trigger corrections at the camera-ready phase.
\end{abstract}

\section{Introduction}
\label{sec:intro}

We approach the problem of capturing the statistical dependence structure of the conditionally independent variational factors in a flexible, Bayesian nonparametric manner. We augment

From the perspective of the variational principle, we propose a drop-in for mean field in which a rich, variational distribution is specified in the expanded space of variables we wish to infer augmented with auxiliary inference variables. This can be seen as specifying a hierarchical variational distribution (CITE RANAGNATH 2015)... integrating over the auxiliary inference parameters recovers a marginal distribution as an infinite mixture of augmented distributions...

From the perspective of variational autoencoding, the GP-LVM-CMF scheme consists of a non-parametric inference network, or \emph{encoder}, which maps the augmented data-auxiliary variable space to the latent variable space. ... should be more flexible than rigidly encoding directly from the non-augmented data space (WHY EXACTLY???)

The marginal variational distribution over the variables of interest, $q(Z,\eta)$, is itself intractable precluding the analyticity of the variational entropy... therefore we lower bound again...

SVI framework permits a learning procedure that requires only unbiased estimates of the lower bound (or rather its gradient), removing the issue of the lack of analyticity of the lower bound and, for no extra cost, granting a highly data-scalable mini-batch update scheme. There has been much recent work which empirically demonstrates the success of ... despite the presence of two sources of stochasticity.

Controlling the variance of the stochastic lower bounds: reparameterization trick... theoretical Lipschitz result, ease of implementation.

l;asdf. We additionally introduce a kernel based approach to back-constrained amortization in which random functions are trained to capture the encoding process. WHY IS THIS GOOD: provable guarantees of kernel mappings, ameliorates pathological curvature??
This also provokes a second look at the VAE in which, in addition to replacing the decoder with a Bayesian non-parametric mapping, the same adjustment is made for the decoding part of the architecture.

We showcase the GP-LVM-CMF scheme ...
Experiments:

\section{Background principles}
\label{sec:background}

\subsection{Variational Inference}
\label{subsec:vi}

\begin{itemize}
  \item Fixed form and factorised... fixed-point coordinate ascent update scheme. Loss of posterior statistical dependence - particularly bad for factor models.
  \item Black-box approaches and SVI. Reparam trick for variance reduction, autoencoding and back-constraints for amortization.
  \item Augmented inference space, hierarchical variational distribution.
\end{itemize}

\subsection{The Bayesian GP-LVM}
\label{subsec:gplvm}

\section{GP-LVM-Conditional Mean Field Variational Inference} % (fold)
\label{sec:gplvmcmf}

-------------THIS SHOULD POSSIBLY GO IN THE INTRO--------------

\begin{itemize}
  \item Swapping out the directly parameterised $q(z)$ of a fixed-form VB distribution for the augmented GP-LVM-CMF distribution leads to an intractable lower bound.
  \item While it is true that we have already forsaken analyticity by performing mini-batch stochastic estimates of the lower bound - so it doesn't worsen things in the sense of analyticity that we can only sample the entropy, we still have variance control to consider.
  \item By judiciously further lower bounding we show that we can not only avoid such risks of large stoch gradient variance, but do so in a way which is more Bayesian than the typical case.
  \item This procedure results in an \emph{auxiliary} lower bound which was inspired by independent approaches of Salimans and Welling and an UNPUBLISHED TECHNICAL NOTE on reinterpreting the sparse variational Gaussian process framework of Titsias.
  \item In the present work we shed new light on this auxiliary approach, and argue that it as a more correct (in the Bayesian sense) approach to inference than simply sampling the entropy term in the bound, as well as than MCMC.
\end{itemize}

----------------------------------------------------------------

We begin by applying the standard variational principle to lower bound the marginal log likelihood as follows
%
\begin{align}
  \log p_\theta(Y) &= \mathbb{E}_{q_\phi(Z)}[\log p_\theta(Y|Z) - \log q_\phi(Z)]\notag\\
  &+ \text{KL}[q_\phi(Z)\Vert p_\theta(Z|Y)]\label{eq:var_principle}\\
  &\geq \mathbb{E}_{q_\phi(Z)}[\log p_\theta(Y|Z) - \log q_\phi(Z)] =: \mathcal{L}_1\label{eq:bound1}
\end{align}
 %
 where we denote the set of all fixed point generative (resp. inference) model parameters to be learned by $\theta$ (resp. $\phi$), and we denote the set of all latent variables and parameters subject to inference by $Z$ (though later we will use $\eta$ to ...).
 Integrating \ref{eq:bound1} with respect to the auxiliary variables, noting that the marginal variational distribution over $Z$ can be expressed as $q(Z) = q(Z|F(),U,X)q(F(),U,X)/q(F(),U,X|Z)$, and introducing an auxiliary model $r_\psi(F(),U,X)$, we arrive at the following \emph{auxiliary lower bound}:
 %
 \begin{align}
  \mathcal{L}_1 &\geq \mathbb{E}_{q(Z,F(),U,X)}\big[ \log p(Y|Z)p(Z)r(F(),U,X|Z)\notag\\
   &- \log q(Z|F(),X)q(F(),U,X) \big]\label{eq:lowerlower}\\
    &+ \mathbb{E}_{q(Z)}\big[ \text{KL}[q(F(),U,X|Z)\Vert r(F(),U,X|Z)] \big] =: \mathcal{L}_{aux}.\label{eq:L_aux}
 \end{align}
%
To explain this bound, we find it useful to refer to $q(F(),U,X)$ (resp. $q(F(),U,X|Z)$) as the \emph{variational prior} (resp.) \emph{variational posterior} distributions over the auxiliary inference variables, the latter of which constitutes a posterior belief over the auxiliary variables having observed a (set of) sample(s), $Z$, generated under the measure $q$ endowed by the inference network.

This lower bound is indistinguishable from the one we would arrive at if we were to apply the variational principle to the problem of trying to estimate (lower bound) the log evidence of the augmented model $p(Z)p(Y|Z)r(U|Z)$: getting r exactly right (i.e. choosing $r(U|Z)=q(U|Z)$) would recover $\mathcal{L}_1$ but intractability precludes knowing $q(U|Z)$ so we must replace it with a tractable surrogate, and perform inference in the augmented space rather than in the marginal $Z$ space. --> Find the form of the surrogate $r$ which leads to a variational inference task in an augmented space which is as equivalent as possible to performing variational inference in the collapsed space.

+++++++++++++++++++++++++++++++++++++++
+++++++++++++++++++++++++++++++++++++++
+++++++++++++++++++++++++++++++++++++++

\subsection{Scalable and Amortized Inference}
\label{subsec:scale}

Having forsaken analyticity in favour a rich and flexibly variational distribution, we must resort to gradient based methods to optimising the auxiliary lower bound with respect to the set of all parameters; generative, variational and auxiliary $\omega=\{\theta,\phi,\psi\}$, with a simulation-based stochastic approach in which we seek to optimise an MC estimate of the auxiliary lower bound:
%
\begin{align}
  \sum_{j=1}^J [\log p(Y|Z^{(j)})p(Z^{(j)})...]\label{eq:stoch_L_aux}
\end{align}
%
SVI, while introduced as a means of performing variational inference on a stochastic lower bound when the source of stochasticity is the mini-batch approach, popularised the fact that one only requires an unbiased estimate of the lower bound (in addition to satisfying some basic conditions) to converge via stochastic gradient descent on the (local) minima of the true lower bound. There has been much recent success in the way of performing so called \emph{doubly stochastic} variational inference........
Inspired by the success of the approach in the VAE / stochastic backprop in deep networks of learning generative and inference model parameters concurrently, we perform stochastic gradient-based optimization in the joint parameter space of $\omega$. The gradients produced by naively differentiating the MC estimate of (\ref{eq:stoch_L_aux}) will suffer from prohibitively high variance...

variance reduction ... reparam ... A but we also note that the additional lower bounding step (\ref{eq:lowerlower}) allows the resulting auxiliary lower bound (\ref{eq:L_aux}) to be written as

We can rearrange (\ref{eq:L_aux}) as follows:
%
\begin{align}
  \mathcal{L}_{aux} &= \mathbb{E}_{q(Z)}[ \log p(Y|Z) ] - \mathbb{E}_{q(F(),U,X)}[ \text{KL}[q(Z|F(),U,X)\Vert p(Z)] ]\notag\\
  &- \mathbb{E}_{q(Z)}[ \text{KL}[q(F(),U,X)\Vert r(F(),U,X|Z)] ].\label{eq:L_aux_ABC}.
\end{align}
%
As noted by Kingma and Welling, in the case of Gaussian $p(Z)$ and $q(Z|F(),U,X)$, the first K-L divergence can be evaluated analytically and, but the lower lower means we can get a second analytical term... This was also noticed by Tran et al., in their lower lower bound...

Variational compression ... sparse variational GPs ... set $r(F()|U,X,Z):=q(F()|U,X)$ which means that $r(F(),U,X|Z) = q(F()|U,X)r(U,X|Z)$. Full implications of this??? Reduces configuration space... The aptness of this hinges on the veracity of the assumption that ... are sufficient statistics for ... in expectation under ...

\end{document}

### 2016-02-03 GP_LVM_CMF.py
more bug fixed for kernel and mop encodings

@@ -90,24 +90,32 @@ class SGPDV(object):
         elif kernelType == 'RBFnn':
             self.numberOfKernelParameters = 1
         else:
-            RuntimeError('Unrecognised kernel type')
+            raise RuntimeError('Unrecognised kernel type')
 
         self.lowerBound = -np.inf # Lower bound
 
         # Suitably sized zero matrices
         B_R_mat = np.zeros((self.B, self.R), dtype=precision)
         B_B_mat = np.zeros((self.B, self.B), dtype=precision)
+        H_P_mat = np.zeros((self.H, self.P), dtype=precision)
+        H_B_mat = np.zeros((self.H, self.B), dtype=precision)
         M_R_mat = np.zeros((self.M, self.R), dtype=precision)
+        M_H_mat = np.zeros((self.M, self.H), dtype=precision)
         N_R_mat = np.zeros((self.N, self.R), dtype=precision)
         N_N_mat = np.zeros((self.N, self.N), dtype=precision)
         Q_M_mat = np.zeros((self.Q, self.M), dtype=precision)
         Q_B_mat = np.zeros((self.Q, self.B), dtype=precision)
+        R_H_mat = np.zeros((self.R, self.H), dtype=precision)
+        R_vec   = np.zeros((self.R, 1), dtype=precision)
+        H_vec   = np.zeros((self.H, 1), dtype=precision) 
+        M_vec   = np.zeros((self.M, 1), dtype=precision)
         B_vec      = np.zeros((self.B,), dtype=np.int32 )
         BR_BR_mat  = np.zeros((self.B*self.R, self.B*self.R), dtype=precision)
         NR_NR_mat  = np.zeros((self.N*self.R, self.N*self.R), dtype=precision)
         H_QpP_mat  = np.zeros((self.H, (self.Q+self.P) ), dtype=precision )
         H_vec      = np.zeros((self.H, 1), dtype=precision )
         QM_QM_mat  = np.zeros((self.Q*self.M, self.Q*self.M), dtype=precision)
+        One_One_mat= np.zeros((1,1), dtype=precision)
         #Mini batch indicator varible
         self.currentBatch = th.shared(B_vec, name='currentBatch')
 
@@ -151,33 +159,31 @@ class SGPDV(object):
 
         elif self.encoderType == 'MLP':
 
[…]

### 2016-02-03 icml2016/paper.tex
Added to paper.

@@
 
-\subsection{Scalable and Amortized Inference}
+\subsection{Scalable and Amortised Inference}
 \label{subsec:scale}
@@
 
+Amortised inference ... VAE, back-constraints, ... we test several options: MLP vs kernel, q(X) vs r(U,X)...
+
+\section{TO DO:}
+\begin{itemize}
+  \item Graphs for the 3 models.
+\end{itemize}
+
+\section{NOTES}
+
+From another perspective, the partial feedback of of z to the aux params (captured in r) means that there is a term in the cost function (aux lower bound) which is a direct effect of the generative power of aux params to generate z - an indirect is achieved when z is first passed through the generative model.
+So direct feedback is a reinforce step - it reinforces our current belief over z.
+It endows the scheme with a memory!
+
+The above is for the aux lower lower bound in general, but we have even more motivation to use it in the case of the GP-LVM-CMF: want to avoid having to do optimization to make a prediction of a new data point!
+
 \end{document}

### 2016-02-03 GP_LVM_CMF.py
Added 'NoEncoding' options, alternate L function, debugging

@@ -55,13 +55,10 @@ class SGPDV(object):
             dimZ,                   # Dimensionality of the latent variables
             data,                   # [NxP] matrix of observations
             kernelType='RBF',
-            encoderType=None,      # None = undefined, 'MLP' = neural network, 'kernel' = GP
-            encode_qX=False,
-            encode_rX=False,
-            encode_ru=False,
-            encoder_type='MLP',     # 0 = undefined, 1 = neural network, 2 = GP
+            encoderType_qX='FreeForm',  # 'FreeForm', 'MLP', 'Kernel'.
+            encoderType_rX='FreeForm',  # 'FreeForm', 'MLP', 'Kernel', 'NoEncoding'.
+            encoderType_ru='FreeForm',  # 'FreeForm', 'MLP', 'NoEncoding'
             z_optimise=False,
-            phi_optimise=True,
             numberOfEncoderHiddenUnits=0
 
         ):
@@ -76,10 +73,9 @@ class SGPDV(object):
         self.Q = dimZ
         self.H = numberOfEncoderHiddenUnits
 
-        self.encoderType = encoderType
-        self.encode_qX   = encode_qX
-        self.encode_rX   = encode_rX
-        self.encode_ru   = encode_ru
+        self.encoderType_qX = encoderType_qX
+        self.encoderType_rX = encoderType_rX
+        self.encoderType_ru = encoderType_ru
         self.z_optimise  = z_optimise
 
         self.y = th.shared(data)
@@ -101,8 +97,6 @@ class SGPDV(object):
         H_B_mat = np.zeros((self.H, self.B), dtype=precision)
         M_R_mat = np.zeros((self.M, self.R), dtype=precision)
         M_H_mat = np.zeros((self.M, self.H), dtype=precision)
-        N_R_mat = np.zeros((self.N, self.R), dtype=precision)
-        N_N_mat = np.zeros((self.N, self.N), dtype=precision)
         Q_M_mat = np.zeros((self.Q, self.M), dtype=precision)
         Q_B_mat = np.zeros((self.Q, self.B), dtype=precision)
         R_H_mat = np.zeros((self.R, self.H), dtype=precision)
@@ -111,11 +105,15 @@ class SGPDV(object):
         M_vec   = np.zeros((self.M, 1), dtype=precision)
         B_vec      = np.zeros((self.B,), dtype=np.int32 )
[…]

### 2016-02-03 IBP_factor_model.py

@@ -19,12 +19,10 @@ class IBP_Factor(SGPDV):
             dimZ,                   # Dimensionality of the latent variables
             data,                   # [NxP] matrix of observations
             kernelType='RBF',
-            encode_qX=False,
-            encode_rX=False,
-            encode_ru=False,
-            encoderType='MLP',     # 0 = undefined, 1 = neural network, 2 = GP
+            encoderType_qX='FreeForm',  # 'FreeForm', 'MLP', 'Kernel'.
+            encoderType_rX='FreeForm',  # 'FreeForm', 'MLP', 'Kernel', 'NoEncoding'.
+            encoderType_ru='FreeForm',  # 'FreeForm', 'MLP', 'NoEncoding'
             z_optimise=False,
-            phi_optimise=True,
             numHiddenUnits_encoder=0,
             numHiddentUnits_decoder=10,
             continuous=True
@@ -38,12 +36,10 @@ class IBP_Factor(SGPDV):
             dimZ,
             data,
             kernelType,
-            encode_qX,
-            encode_rX,
-            encode_ru,
-            encoderType,
+            encoderType_qX,  # 'FreeForm', 'MLP', 'Kernel'.
+            encoderType_rX,  # 'FreeForm', 'MLP', 'Kernel', 'NoEncoding'.
+            encoderType_ru,  # 'FreeForm', 'MLP', 'NoEncoding'
             z_optimise,
-            phi_optimise,
             numHiddenUnits_encoder
             )
 
@@ -59,32 +55,21 @@ class IBP_Factor(SGPDV):
         scalar    = np.zeros(1, dtype=floatX)
 
         #self.A       = th.shared( K_D_mat )
-        self.Phi_IBP = th.shared( K_D_D_ten )
-        self.phi_IBP = th.shared( K_D_mat)
-        self.tau_IBP = th.shared( K_2_mat )
-        self.mu_IBP  = th.shared( N_K_mat )
-        self.log_alpha_IBP   = th.shared(scalar)
-        self.log_sigma_y = th.shared(scalar)
-        self.log_sigma_A = th.shared(scalar)
-
-
-        self.Phi_IBP.name = 'Phi_IBP'
-        self.phi_IBP.name = 'phi_IBP'
-        self.tau_IBP.name = 'tau_IBP'
[…]

### 2016-02-03 trainmnist_cond_MF.py

@@ -5,7 +5,6 @@ import numpy as np
 import argparse
 import gzip, cPickle
 
-
 parser = argparse.ArgumentParser()
 
 args = parser.parse_args()
@@ -25,25 +24,38 @@ n_iter = 500
 dimZ = 2
 dimX = 5
 HU_decoder = 400
+batchSize = 100
 
-
-<<<<<<< HEAD
-batch_size = 100
-
-n_induce = 5
+numberOfInducingPoints = 5
 learning_rate = 1e-3
 
 r_is_nnet = False
 backConstrainX = False
 
 print "Initialising"
-va = VA(n_induce, batch_size, dimX, dimZ, x_train, HU_decoder, kernelType_='RBF', continuous_=True, backConstrainX=backConstrainX, r_is_nnet=r_is_nnet )
 
-va.construct_L()
+va = VA(
+    numberOfInducingPoints, # Number of inducing ponts in sparse GP
+    batchSize,              # Size of mini batch
+    dimX,                   # Dimensionality of the latent co-ordinates
+    dimZ,                   # Dimensionality of the latent variables
+    data,                   # [NxP] matrix of observations
+    kernelType='RBF',
+    encoderType_qX='MLP',  # 'FreeForm', 'MLP', 'Kernel'.
+    encoderType_rX='MLP',  # 'FreeForm', 'MLP', 'Kernel', 'NoEncoding'.
+    encoderType_ru='FreeForm',  # 'FreeForm', 'MLP', 'NoEncoding'
+    z_optimise=False,
+    numHiddenUnits_encoder=0,
+    numHiddentUnits_decoder=10,
+    continuous=True
+)
 
-va.setHyperparameters(0.01, 5*np.ones((2,)),
+va.construct_L_using_r()
+
+va.setKernelParameters(0.01, 5*np.ones((2,)),
     1e-100, 0.5,
     [1e-10,1e-10], [10,10] )
+
 va.randomise()
 va.sample()
 
@@ -55,7 +67,7 @@ f.write(str(temp))
 f.close()
 
 print "Training"
-lowerBounds = va.train_adagrad( n_iter, learning_rate )
+lowerBounds = va.train_adagrad( n_iter, learningRate=learning_rate )
 
 
 

### 2016-02-03 GP_LVM_CMF.py
minor changes.

@@ -107,7 +107,7 @@ class SGPDV(object):
         Q_B_mat = np.zeros((self.Q, self.B), dtype=precision)
         R_H_mat = np.zeros((self.R, self.H), dtype=precision)
         R_vec   = np.zeros((self.R, 1), dtype=precision)
-        H_vec   = np.zeros((self.H, 1), dtype=precision) 
+        H_vec   = np.zeros((self.H, 1), dtype=precision)
         M_vec   = np.zeros((self.M, 1), dtype=precision)
         B_vec      = np.zeros((self.B,), dtype=np.int32 )
         BR_BR_mat  = np.zeros((self.B*self.R, self.B*self.R), dtype=precision)
@@ -164,7 +164,7 @@ class SGPDV(object):
             self.W1_qX = th.shared(H_P_mat, name='W1_qX')
             self.b1_qX = th.shared(H_vec,   name='b1_qX', broadcastable=(False,True))
             self.W2_qX = th.shared(R_H_mat, name='W2_qX')
-            self.b2_qX = th.shared(R_vec,   name='b2_qX', broadcastable=(False,True)) 
+            self.b2_qX = th.shared(R_vec,   name='b2_qX', broadcastable=(False,True))
             self.W3_qX = th.shared(H_vec.T, name='W3_qX')
             self.b3_qX = th.shared(One_One_mat, name='b3_qX', broadcastable=(False,True))
 
@@ -261,7 +261,7 @@ class SGPDV(object):
             self.rX_vars = [self.Tau, self.tau_full]
 
         elif self.encoderType == 'MLP':
-            
+
             self.W1_rX = th.shared(H_QpP_mat, name='W1_rX')
             self.b1_rX = th.shared(H_vec,     name='b1_rX', broadcastable=(False,True))
             self.W2_rX = th.shared(R_H_mat,   name='W2_rX')
@@ -296,7 +296,7 @@ class SGPDV(object):
 
         elif self.encoderType == 'kernel':
 
-            #Tau_r [BxB] =  
+            #Tau_r [BxB] =
             Tau_r = kfactory.kernel(T.concatenate((self.z,self.y_miniBatch.T)).T, None, self.log_omega, 'Tau_r')
             (cTau_r,iTau_r,logDetTau_r) = cholInvLogDet(Tau_r, self.B, self.jitter)
 
@@ -396,14 +396,14 @@ class SGPDV(object):
             elif var.name.startswith('W1') or \
                  var.name.startswith('W2') or \
                  var.name.startswith('W3'):
-                
+
[…]

### 2016-02-03 GP_LVM_CMF.py
trying to merge.

@@ -927,9 +927,7 @@ class VA(SGPDV):
             encode_qX=False,
             encode_rX=False,
             encode_ru=False,
-            encoder_type='MLP',
             z_optimise=False,
-            phi_optimise=True,             # 0 = undefined, 1 = neural network, 2 = GP
             numHiddenUnits_encoder=0,
             numHiddentUnits_decoder=10,
             continuous=True
@@ -949,7 +947,6 @@ class VA(SGPDV):
             encode_ru,
             encoder_type,           # 0 = undefined, 1 = neural network, 2 = GP
             z_optimise,
-            phi_optimise,
             numHiddenUnits_encoder
         )
 

### 2016-02-03 trainmnist_cond_MF.py

@@ -27,7 +27,6 @@ dimX = 5
 HU_decoder = 400
 
 
-<<<<<<< HEAD
 batch_size = 100
 
 n_induce = 5
@@ -56,6 +55,9 @@ f.close()
 
 print "Training"
 lowerBounds = va.train_adagrad( n_iter, learning_rate )
+
+print "Testing"
+
 
 
 

### 2016-02-08 icml2016/paper.tex
Additions to paper.

@@
 
-We approach the problem of capturing the statistical dependence structure of the conditionally independent variational factors in a flexible, Bayesian nonparametric manner. We augment
+We approach the problem of capturing the statistical dependence structure of the conditionally independent variational factors in a flexible, Bayesian nonparametric manner.
+Considering at first a latent variable model with a latent variable for each datum as well as some global parameters, a traditional (mean field) variational approach is to learn a factored approximation to the posterior in which each latent variable is independently distributed and governed by its own variational parameter.
+Clearly this is an overly restrictive approximation since conditioning on the data statistically couples the latent variables.
+With careful design some statistical dependence may be retained with so-called \emph{structured mean field} approximations (SAUL AND JORDAN, WIERGERNINCK), but the design is highly model specific and not AUTOMATICALLY SCALABLE?
+A natural improvement is to couple the variational factors via a common variational prior, leading to a hierarchical variational approximation (RANGANATH 2015) which are richer and more expressive.
+HOWEVER THE NECESSARY CRUCIAL DESIGN CHOICES MAKE THIS PROCEDURE DIFFICULT TO AUTOMATE AND SCALE?
 
-From the perspective of the variational principle, we propose a drop-in for mean field in which a rich, variational distribution is specified in the expanded space of variables we wish to infer augmented with auxiliary inference variables.
-This can be seen as specifying a hierarchical variational distribution (CITE RANAGNATH 2015)...
+Rather than designing model specific, parametric variational priors..., we take inspiration from the Bayesian non-parametric dimensionality reduction method of..
+GP-LVM...
+to learn a flexible, non-linear mapping from a shared latent auxiliary space to each conditionally independent variational factor.
+The dependence is induced by a fixed number of latent coordinates, the locations of which are learned so as to induce the optimal dependence structure.
+We refer to this scheme as the GP-LVM-conditional mean field
+
+From the perspective of the variational principle, we propose a drop-in replacement for mean field inference in which a rich, variational distribution is specified in the expanded space of variables we wish to infer augmented with auxiliary inference variables.
+This can be seen as specifying a non-parametric hierarchical variational model (CITE RANAGNATH 2015)...
 integrating over the auxiliary inference parameters recovers a marginal distribution as an infinite mixture of augmented distributions...
 
-From the perspective of variational autoencoding, the GP-LVM-CMF scheme consists of a non-parametric inference network, or \emph{encoder}, which maps the augmented data-auxiliary variable space to the latent variable space.
+From the perspective of recognition models and variational autoencoding, the GP-LVM-CMF scheme consists of a non-parametric inference network, or \emph{encoder}, which maps the augmented data-auxiliary variable space to the latent variable space.
 ...
@@
 \begin{itemize}
-  \item Fixed form and factorised... fixed-point coordinate ascent update scheme. Loss of posterior statistical dependence - particularly bad for factor models.
+  \item Fixed form and factorised... fixed-point coordinate ascent update scheme. Loss of posterior statistical dependence - particularly bad for factor models... --> motivating the IBP experiments.
   \item Black-box approaches and SVI. Reparam trick for variance reduction, autoencoding and back-constraints for amortization.
@@
 
+The GP-LVM was proposed by Lawrence as a Bayesian non-linear dimensionality reduction model and can be viewed as a multi-output GP regression model in which the inputs are unobserved and so treated as latent variables.
+In the original framework these inputs were optimised rather than integrated out for tractability.
+
+Denoting the $d^{th}$ column of the data matrix, $Y$, as $\mathbf{y}_d$ the likelihood function for the GP-LVM is
+%
+\begin{align}
+  p(Y|X) &= \prod_{d=1}^D p(\mathbf{y}_d | X),\\
+  p(\mathbf{y}_d) &= \mathcal{N}(\mathbf{y}_d|\mathbf{0}, K_{ff}+\beta^{-1}I)
+\end{align}
+%
+where the GPs are modelled as independent across the data dimensions.
+
+\begin{itemize}
+  \item GP-LVM
+  \item variational compression and SVGP.
+  \item Bayesian GP-LVM
+  \item Provides an extremely general, Bayesian and rich way of auxiliary object upon which to condition the mean field factors in order to couple them statistically.
+\end{itemize}
+
 \section{GP-LVM-Conditional Mean Field Variational Inference} % (fold)
 \label{sec:gplvmcmf}
+
+As is the case for the SVGP and the Bayesian GP-LVM, the inference model of the GP-LVM-CMF constructs a GP mapping to $N$ output values which is governed by a fixed number of auxiliary inducing variables.
+However the motivations for this conditional inducing structure are distinct in these three cases.
+For the sparse GP the desire is scalability; in particular to circumvent the costly $\mathcal{O}(N^3$ complexity of the covariance matrix inversion, which is then limited to $\mathcal{O}(N^2M)$ through variational compression.
+For the Bayesian GP-LVM the motivation is to be able to tractably integrate over the uncertainty associated with the unknown latent input locations.
+For the GP-LVM-CMF however there are no observations of the GP function values since they do not exist in the generative model, so their purpose here is to act as \emph{variational (pseudo) data} upon which the rest of the function values can be conditioned in the standard GP formulation.
 
@@
 
+Hensman SVI GVGP does do some amortising since the global inducing variables provide a statistical conduit...
+
 Having forsaken analyticity in favour a rich and flexibly variational distribution, we must resort to gradient based methods to optimising the auxiliary lower bound with respect to the set of all parameters; generative, variational and auxiliary $\omega=\{\theta,\phi,\psi\}$, with a simulation-based stochastic approach in which we seek to optimise an MC estimate of the auxiliary lower bound:
@@
 
+\subsection{Design Choices}
+\label{subsec:choices}
+
+The alternative configurations...
+
+\section{Related Work}
+\label{sec:related}
+
+Coupling mean field factors: Tran's Copula approach, the other copula approach ??, Michael Jordan paper.
+
+The independent works of Kingma and Welling and Rezende et al.
+popularised the modern approach of autoencoding in the variational framework, and spurred on a great research drive in variational approaches to deep learning.
+Highly sophisticated models such as DRAW - a method to....
+- have at their core the VAE...
+Grosse et al considered an alternative variational lower bound based on importance weights to train a VAE.
+However in all these approaches, the
+
+deep gp - used autoencoding as backconstraints to make it scale (finite number of parameters)
+
+Salimans and Welling considered a Markov chain inference network in which the final ($T^th$) Markov transition in the chain produced the latent variable (sample), such that previous $T-1$ variables were treated as auxiliary variables.
+Collapsing over this expanded auxiliary space then resulted in a rich yet intractable mixture...
+and tractability was recovered through the use of an auxiliary distribution also of a Markov structure.
+
+Most similar in spirit to GP-LVM-CMF approach is that of the \emph{Variational Gaussian Process} developed independently by Tran et al., which was also used conditional GPs (conditioned on aux params) in the inference model to couple the factors.
+The key difference is that each latent variable is encoded with a \emph{separate} conditional GP...
+and the coupling between the latent variables was induced by evaluating each conditional GP at the same input location.
+This leads to coupling whereby each GP tends to be positive in similar regions ?????.
+In our approach the coupling is induced via the shared inducing coordinates...
+more faithfully to the spirit of the GP-LVM.
+
+\section{Experimental Setup}
+\label{sec:experiments}
+
+...
+free form q(X) means having to optimise the auxiliary latent coordinate of the test location just to be able to evaluate.
+Including an optimisation procedure in the predictive step is an interesting idea, but was avoided here for lack of space and clarity of discussion, so investigation was deferred to a future study.
+
+\section{Discussion and Future Work}
+\label{sec:discussion}
+
 \section{TO DO:}
@@
   \item Graphs for the 3 models.
+  \item Algorithm in section \ref{sec:gplvmcmf}
 \end{itemize}

### 2016-02-08 GP_LVM_CMF_2.py
committing before merge.

@@ -519,7 +519,7 @@ class SGPDV(object):
 
         # generate standard gaussian random varibales        
         def rnd( rv ):
-            rv.set_value( np.random.randn( rv.get_value().shape ) )
+            rv.set_value( np.random.randn( rv.get_value().shape ).astype(th.config.floatX) )
         
         rnd( self.alpha )
         rnd( self.beta )
@@ -617,15 +617,15 @@ class SGPDV(object):
             if self.gradientVariables[i] == self.Upsilon:
                 self.Upsilon_lower = values[i]
                 #print  np.dot(self.Upsilon_lower, self.Upsilon_lower.T)
-                self.Upsilon.set_value( np.dot(self.Upsilon_lower, self.Upsilon_lower.T) )
+                self.Upsilon.set_value( np.dot(self.Upsilon_lower, self.Upsilon_lower.T).astype(th.config.floatX) )
             elif self.gradientVariables[i] == self.Phi_full:
                 self.Phi_full_lower = values[i]
-                self.Phi_full.set_value( np.dot(self.Phi_full_lower, self.Phi_full_lower.T) )
+                self.Phi_full.set_value( np.dot(self.Phi_full_lower, self.Phi_full_lower.T).astype(th.config.floatX) )
             elif self.gradientVariables[i] == self.Tau_full:
                 self.Tau_full_lower = values[i]
-                self.Tau_full.set_value( np.dot(self.Tau_full_lower, self.Tau_full_lower.T) )
+                self.Tau_full.set_value( np.dot(self.Tau_full_lower, self.Tau_full_lower.T).astype(th.config.floatX) )
             else:
-                self.gradientVariables[i].set_value( values[i] )
+                self.gradientVariables[i].set_value( values[i].astype(th.config.floatX) )
 
     def getVariableValues(self):
 
@@ -642,7 +642,7 @@ class SGPDV(object):
                 values[i] = self.gradientVariables[i].get_value()
         return values
 
-    def getTestLowerBound(self):
+    def getTestPredictiveLhood(self):
         return 0
 
     def copyParameters(self, other):

### 2016-02-08 trainmnist_cond_MF.py

@@ -99,5 +99,5 @@ vatest = va = VA(
 
 vatest.copyParameters(va)
 
-testLogLhood = vatest.getTestLowerBound()
+testPredLhood = vatest.getTestPredictiveLhood()
 print testLogLhood

### 2016-02-08 fastlin/choleskyGrad.c
fastlin and sandbox added

@@ -1,0 +1,16 @@
+
+#include <stdint.h>
+#include "choleskyGrad.h"
+
+extern void dpofrt_(
+        char *uplo, const int32_t *n, double *a, const int32_t *lda,
+        double *c, const int32_t *ldc, const int32_t *info);
+
+int choleskyGrad( char UPLO, double* L, double* F, int N ) {
+    
+    int32_t info;
+        
+    dpofrt_(&UPLO, &N, L, &N, F, &N, &info);
+    
+    return info;
+}

### 2016-02-08 fastlin/choleskyGrad.h

@@ -1,0 +1,2 @@
+
+int choleskyGrad( char UPLO, double* L, double* F, int N );

### 2016-02-08 fastlin/myCholesky.py

@@ -1,0 +1,132 @@
+# -*- coding: utf-8 -*-
+"""
+Created on Sat Feb  6 19:19:05 2016
+
+@author: clloyd
+"""
+
+from six.moves import xrange
+
+import numpy
+import scipy.linalg
+from theano.tensor import as_tensor_variable
+from theano.gof import Op, Apply
+
+
+class myCholesky(Op):
+    """
+    Return a triangular matrix square root of positive semi-definite `x`.
+    L = cholesky(X, lower=True) implies dot(L, L.T) == X.
+    """
+    # TODO: inplace
+    # TODO: for specific dtypes
+    # TODO: LAPACK wrapper with in-place behavior, for solve also
+
+    __props__ = ('lower', 'destructive')
+
+    def __init__(self, lower=True):
+        self.lower = lower
+        self.destructive = False
+
+    def infer_shape(self, node, shapes):
+        return [shapes[0]]
+
+    def make_node(self, x):
+        x = as_tensor_variable(x)
+        assert x.ndim == 2
+        return Apply(self, [x], [x.type()])
+
+    def perform(self, node, inputs, outputs):
+        x = inputs[0]
+        z = outputs[0]
+        z[0] = scipy.linalg.cholesky(x, lower=self.lower).astype(x.dtype)
+
+    def grad(self, inputs, gradients):
+        return [myCholeskyGrad(self.lower)(inputs[0], self(inputs[0]),
+                                         gradients[0])]
+
+class myCholeskyGrad(Op):
+    """
+    """
+
+    __props__ = ('lower', 'destructive')
+
+    def __init__(self, lower=True):
+        self.lower = lower
+        self.destructive = False
+
+    def make_node(self, x, l, dz):
+        x = as_tensor_variable(x)
+        l = as_tensor_variable(l)
+        dz = as_tensor_variable(dz)
+        assert x.ndim == 2
+        assert l.ndim == 2
+        assert dz.ndim == 2
+        assert l.owner.op.lower == self.lower, (
+            "lower/upper mismatch between Cholesky op and CholeskyGrad op"
+        )
+        return Apply(self, [x, l, dz], [x.type()])
+
+    def perform(self, node, inputs, outputs):
+        """
+        Implements the "reverse-mode" gradient [1]_ for the
+        Cholesky factorization of a positive-definite matrix.
+        References
+        ----------
+        .. [1] S. P. Smith. "Differentiation of the Cholesky Algorithm".
[…]

### 2016-02-08 fastlin/myEig.py

@@ -1,0 +1,152 @@
+# -*- coding: utf-8 -*-
+"""
+Created on Sat Feb  6 16:39:41 2016
+
+@author: clloyd
+"""
+
+from __future__ import print_function
+
+import numpy
+from six.moves import xrange
+
+import theano
+from theano.tensor import as_tensor_variable
+from theano.gof import Op, Apply
+from theano.gradient import DisconnectedType
+
+class myEigh(Op):
+    """
+    Return the eigenvalues and eigenvectors of a Hermitian or symmetric matrix.
+    """
+
+    _numop = staticmethod(numpy.linalg.eigh)
+    __props__ = ('UPLO',)
+
+    def __init__(self, UPLO='L'):
+        assert UPLO in ['L', 'U']
+        self.UPLO = UPLO
+
+    def make_node(self, x):
+        x = as_tensor_variable(x)
+        assert x.ndim == 2
+        # Numpy's linalg.eigh may return either double or single
+        # presision eigenvalues depending on installed version of
+        # LAPACK.  Rather than trying to reproduce the (rather
+        # involved) logic, we just probe linalg.eigh with a trivial
+        # input.
+        w_dtype = self._numop([[numpy.dtype(x.dtype).type()]])[0].dtype.name
+        w = theano.tensor.vector(dtype=w_dtype)
+        v = theano.tensor.matrix(dtype=x.dtype)
+        return Apply(self, [x], [w, v])
+
+    def perform(self, node, inputs, outputs):
+        (x,) = inputs
+        (w, v) = outputs
+        w[0], v[0] = self._numop(x, self.UPLO)
+
+    def infer_shape(self, node, shapes):
+        n = shapes[0][0]
+        return [(n,), (n, n)]
+
+    def grad(self, inputs, g_outputs):
+        r"""The gradient function should return
+           .. math:: \sum_n\left(W_n\frac{\partial\,w_n}
+                           {\partial a_{ij}} +
+                     \sum_k V_{nk}\frac{\partial\,v_{nk}}
+                           {\partial a_{ij}}\right),
+        where [:math:`W`, :math:`V`] corresponds to ``g_outputs``,
+        :math:`a` to ``inputs``, and  :math:`(w, v)=\mbox{eig}(a)`.
+        Analytic formulae for eigensystem gradients are well-known in
+        perturbation theory:
+           .. math:: \frac{\partial\,w_n}
+                          {\partial a_{ij}} = v_{in}\,v_{jn}
[…]

### 2016-02-08 sandbox/vectorTimesScalar.py

@@ -1,0 +1,89 @@
+# -*- coding: utf-8 -*-
+"""
+Created on Mon Feb  8 18:49:48 2016
+
+@author: clloyd
+"""
+
+import numpy
+import theano
+from theano import gof
+import theano.tensor as T
+
+class VectorTimesScalar(gof.Op):
+    __props__ = ()
+
+    def make_node(self, x, y):
+        # Validate the inputs' type
+        if x.type.ndim != 1:
+            raise TypeError('x must be a 1-d vector')
+        if y.type.ndim != 0:
+            raise TypeError('y must be a scalar')
+
+        # Create an output variable of the same type as x
+        output_var = x.type()
+
+        return gof.Apply(self, [x, y], [output_var])
+
+    def c_code_cache_version(self):
+        return (1, 0)
+
+    def c_code(self, node, name, inp, out, sub):
+        x, y = inp
+        z, = out
+
+        # Extract the dtypes of the inputs and outputs storage to
+        # be able to declare pointers for those dtypes in the C
+        # code.
+        dtype_x = node.inputs[0].dtype
+        dtype_y = node.inputs[1].dtype
+        dtype_z = node.outputs[0].dtype
+
+        itemsize_x = numpy.dtype(dtype_x).itemsize
+        itemsize_z = numpy.dtype(dtype_z).itemsize
+
+        fail = sub['fail']
+
+        c_code = """
+        // Validate that the output storage exists and has the same
+        // dimension as x.
+        if (NULL == %(z)s ||
+            PyArray_DIMS(%(x)s)[0] != PyArray_DIMS(%(z)s)[0])
+        {
+            /* Reference received to invalid output variable.
+            Decrease received reference's ref count and allocate new
+            output variable */
+            Py_XDECREF(%(z)s);
+            %(z)s = (PyArrayObject*)PyArray_EMPTY(1,
+                                                PyArray_DIMS(%(x)s),
+                                                PyArray_TYPE(%(x)s),
+                                                0);
+
+            if (!%(z)s) {
+                %(fail)s;
+            }
+        }
+
+        // Perform the vector multiplication by a scalar
+        {
+            /* The declaration of the following variables is done in a new
+            scope to prevent cross initialization errors */
+            npy_%(dtype_x)s* x_data_ptr =
+                            (npy_%(dtype_x)s*)PyArray_DATA(%(x)s);
[…]

### 2016-02-08 sandbox/vectorTimesVector.py

@@ -1,0 +1,34 @@
+# -*- coding: utf-8 -*-
+"""
+Created on Mon Feb  8 18:04:54 2016
+
+@author: clloyd
+"""
+
+import theano
+from theano import gof
+
+class VectorTimesVector(gof.COp):
+    __props__ = ()
+
+    func_file = "./vectorTimesVector.c"
+    func_name = "APPLY_SPECIFIC(vector_times_vector_fred)"
+
+    def __init__(self):
+        super(VectorTimesVector, self).__init__(self.func_file,
+                                                self.func_name)
+
+    def make_node(self, x, y):
+        # Validate the inputs' type
+        if x.type.ndim != 1:
+            raise TypeError('x must be a 1-d vector')
+        if y.type.ndim != 1:
+            raise TypeError('y must be a 1-d vector')
+
+        # Create an output variable of the same type as x
+        output_var = theano.tensor.TensorType(
+                        dtype=theano.scalar.upcast(x.dtype, y.dtype),
+                        broadcastable=[False])()
+
+        return gof.Apply(self, [x, y], [output_var])
+

### 2016-02-15 GP_LVM_CMF.py
minor changes

@@ -430,7 +430,7 @@ class SGPDV(object):
         self.gradientVariables.extend(self.rX_vars)
         self.gradientVariables.extend(self.ru_vars)
 
-        # self.profmode = th.ProfileMode(optimizer='fast_run', linker=th.gof.OpWiseCLinker())
+        self.profmode = th.ProfileMode(optimizer='fast_run', linker=th.gof.OpWiseCLinker())
 
     def randomise(self, sig=1, rndQR=False):
 
@@ -702,7 +702,7 @@ class SGPDV(object):
 
         updates = self.optimiser.updatesIgrad_model(gradColl, self.gradientVariables)
 
-        self.updateFunction = th.function([], None, updates=updates, no_default_updates=True)  # ,  mode=self.profmode)
+        self.updateFunction = th.function([], None, updates=updates, no_default_updates=True,  mode=self.profmode)
 
     def train(self, numberOfEpochs=1, learningRate=1e-3, fudgeFactor=1e-6, maxIters=np.inf):
 

### 2016-02-15 trainmnist_cond_MF.py

@@ -94,25 +94,25 @@ va.train(numberOfEpochs=numberOfEpochs, maxIters=10)
 
 #lowerBounds = va.train_adagrad( n_iter, learningRate=learning_rate )
 #
-print "Testing"
-vatest = va = VA(
-    numberOfInducingPoints, # Number of inducing ponts in sparse GP
-    batchSize,              # Size of mini batch
-    dimX,                   # Dimensionality of the latent co-ordinates
-    dimZ,                   # Dimensionality of the latent variables
-    x_test,                   # [NxP] matrix of observations
-    kernelType=kernelType,
-    encoderType_qX=encoderType_qX,  # 'FreeForm', 'MLP', 'Kernel'.
-    encoderType_rX=encoderType_rX,  # 'FreeForm', 'MLP', 'Kernel', 'NoEncoding'.
-    encoderType_ru=encoderType_ru,  # 'FreeForm', 'MLP', 'NoEncoding'
-    Xu_optimise=Xu_optimise,
-    numHiddenUnits_encoder=numHiddenUnits_encoder,
-    numHiddentUnits_decoder=numHiddentUnits_decoder,
-    continuous=True
-)
+#print "Testing"
+#vatest = va = VA(
+#    numberOfInducingPoints, # Number of inducing ponts in sparse GP
+#    batchSize,              # Size of mini batch
+#    dimX,                   # Dimensionality of the latent co-ordinates
+#    dimZ,                   # Dimensionality of the latent variables
+#    x_test,                   # [NxP] matrix of observations
+#    kernelType=kernelType,
+#    encoderType_qX=encoderType_qX,  # 'FreeForm', 'MLP', 'Kernel'.
+#    encoderType_rX=encoderType_rX,  # 'FreeForm', 'MLP', 'Kernel', 'NoEncoding'.
+#    encoderType_ru=encoderType_ru,  # 'FreeForm', 'MLP', 'NoEncoding'
+#    Xu_optimise=Xu_optimise,
+#    numHiddenUnits_encoder=numHiddenUnits_encoder,
+#    numHiddentUnits_decoder=numHiddentUnits_decoder,
+#    continuous=True
+#)
 
-va.construct_L_using_r()
-vatest.copyParameters(va)
+#va.construct_L_using_r()
+#vatest.copyParameters(va)
 
-testLogLhood = vatest.getMCLogLikelihood(numTestSamples)
+#testLogLhood = vatest.getMCLogLikelihood(numTestSamples)
 print testLogLhood

### 2016-02-15 fastlin/myCholeskyGrad.c
Minor change to make .c chol grad compile under linux

@@ -95,8 +95,7 @@ int APPLY_SPECIFIC(apply_cholesky_grad)(PyArrayObject* input0,
         if(info != 0) {
             rCode = 3;
             Py_XDECREF(F);
-            std::string ermsg = std::string("choleskyGrad returned error code: ") + std::to_string(info);
-            PyErr_Format(PyExc_ValueError, ermsg.c_str());
+            PyErr_Format(PyExc_ValueError, "cholgrad library returned error code %d\n", info);
         }
         // We don't need the reference to L anymore
         Py_XDECREF(L);

### 2016-02-15 GP_LVM_CMF.py
Changes to make work with GPU (allow 64bit -> 32 downcasting

@@ -143,8 +143,12 @@ class SGPDV(object):
         self.y_miniBatch.name = 'y_miniBatch'
 
         # This is for numerical stability of cholesky
-        self.jitterDefault = np.float32(0.0117)
-        self.jitterGrowthFactor = np.float32(1.1)
+	if th.config.floatX == 'float32'
+	    # Need a lot more jitter for 32 bit computation       
+	    self.jitterDefault = np.float32(0.1)
+	else
+	    self.jitterDefault = np.float32(0.0001)
+	self.jitterGrowthFactor = np.float32(1.1)
         self.jitter = th.shared(np.asarray(self.jitterDefault, dtype=precision), name='jitter')
 
         kfactory = kernelFactory(kernelType)
@@ -431,7 +435,7 @@ class SGPDV(object):
         self.gradientVariables.extend(self.rX_vars)
         self.gradientVariables.extend(self.ru_vars)
 
-        self.profmode = th.ProfileMode(optimizer='fast_run', linker=th.gof.OpWiseCLinker())
+        #self.profmode = th.ProfileMode(optimizer='fast_run', linker=th.gof.OpWiseCLinker())
 
     def randomise(self, sig=1, rndQR=False):
 
@@ -708,7 +712,7 @@ class SGPDV(object):
 
         updates = self.optimiser.updatesIgrad_model(gradColl, self.gradientVariables)
 
-        self.updateFunction = th.function([], None, updates=updates, no_default_updates=True,  mode=self.profmode)
+        self.updateFunction = th.function([], None, updates=updates, no_default_updates=True)#,  mode=self.profmode)
 
     def train(self, numberOfEpochs=1, learningRate=1e-3, fudgeFactor=1e-6, maxIters=np.inf):
 
@@ -778,8 +782,9 @@ class SGPDV(object):
                 val = func()
                 passed = True
             except np.linalg.LinAlgError:
-                print 'Increasing value of jitter'
-                self.jitter.set_value(self.jitter.get_value() * self.jitterGrowthFactor)
+		self.jitter.set_value(self.jitter.get_value() * self.jitterGrowthFactor)                
+		print 'Increasing value of jitter. Jitter now: ' + str(self.jitter.get_value())
+                
         if reset:
             self.jitter.set_value(self.jitterDefault)
         return val

### 2016-02-15 trainmnist_cond_MF.py

@@ -116,4 +116,4 @@ va.train(numberOfEpochs=numberOfEpochs, maxIters=10)
 #vatest.copyParameters(va)
 
 #testLogLhood = vatest.getMCLogLikelihood(numTestSamples)
-print testLogLhood
+#print testLogLhood

### 2016-02-16 fastlin/myCholeskyGrad.c
added 64 bit -> 32 changes properly this time

@@ -107,11 +107,13 @@ int APPLY_SPECIFIC(apply_cholesky_grad)(PyArrayObject* input0,
             // If so, decrement the count to that memory
             Py_XDECREF(*output0);
         }
-        // Gradient has same type of thing we're taking gradient of, i.e. input 1
-        PyArray_Descr* output_descr = PyArray_DescrFromObject((PyObject*) input0, NULL);
-        // Copy the F to the right type format for the output: Can use C or F memory styles
-        *output0 = (PyArrayObject*) PyArray_FromAny((PyObject*)F, output_descr, 2, 2, NPY_ARRAY_F_CONTIGUOUS, NULL);
-        if(*output0 == NULL){
+	
+
+        // Get descriptor of output type
+        PyArray_Descr* output_descr =  PyArray_DescrFromType(TYPENUM_OUTPUT_0);
+	// Copy the F to the right type format for the output: Can use C or F memory styles
+        *output0 = (PyArrayObject*) PyArray_FromAny((PyObject*)F, output_descr, 2, 2, NPY_ARRAY_F_CONTIGUOUS|NPY_ARRAY_FORCECAST, NULL);
+        if(*output0 == NULL) {
             rCode = 4;
             PyErr_Format(PyExc_ValueError, "Could not allocate output storage");
         }

### 2016-02-20 icml2016/example_paper.tex
changed diag to log diag

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%% ICML 2016 EXAMPLE LATEX SUBMISSION FILE %%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% Use the following line _only_ if you're still using LaTeX 2.09.
%\documentstyle[icml2016,epsf,natbib]{article}
% If you rely on Latex2e packages, like most moden people use this:
\documentclass{article}

% use Times
\usepackage{times}
% For figures
\usepackage{graphicx} % more modern
%\usepackage{epsfig} % less modern
\usepackage{subfigure}

% For citations
\usepackage{natbib}

% For algorithms
\usepackage{algorithm}
\usepackage{algorithmic}

% As of 2011, we use the hyperref package to produce hyperlinks in the
% resulting PDF.  If this breaks your system, please commend out the
% following usepackage line and replace \usepackage{icml2016} with
% \usepackage[nohyperref]{icml2016} above.
\usepackage{hyperref}

% Packages hyperref and algorithmic misbehave sometimes.  We can fix
% this with the following command.
\newcommand{\theHalgorithm}{\arabic{algorithm}}

% Employ the following version of the ``usepackage'' statement for
% submitting the draft version of the paper for review.  This will set
% the note in the first column to ``Under review.  Do not distribute.''
%\usepackage{icml2016}

% Employ this version of the ``usepackage'' statement after the paper has
% been accepted, when creating the final version.  This will set the
% note in the first column to ``Proceedings of the...''
\usepackage{proceed}

% The \icmltitle you define below is probably too long as a header.
% Therefore, a short form for the running title is supplied here:
\icmltitlerunning{Submission and Formatting Instructions for ICML 2016}

\begin{document}

\begin{abstract} 
The purpose of this document is to provide both the basic paper template and
submission guidelines. Abstracts should be a single paragraph, between 4--6 sentences long, ideally.  Gross violations will trigger corrections at the camera-ready phase.
\end{abstract}

\section{Electronic Submission}
\label{submission}

Submission to ICML 2016 will be entirely electronic, via a web site
(not email).  Information about the submission process and \LaTeX\ templates
are available on the conference web site at:
\begin{center}
\textbf{\texttt{http://icml.cc/2016/}}
\end{center}
Send questions about submission and electronic templates to
\texttt{icml2016pc@gmail.com}.

The guidelines below will be enforced for initial submissions and
camera-ready copies.  Here is a brief summary:
\begin{itemize}
\item Submissions must be in PDF.
\item The maximum paper length is \textbf{8 pages excluding references and acknowledgements, and 10 pages
  including references and acknowledgements} (pages 9 and 10 must contain only references and acknowledgements).
\item Do \textbf{not include author information or acknowledgements} in your initial
submission. 
\item Your paper should be in \textbf{10 point Times font}.
\item Make sure your PDF file only uses Type-1 fonts.
\item Place figure captions {\em under} the figure (and omit titles from inside
the graphic file itself).  Place table captions {\em over} the table.
\item References must include page numbers whenever possible and be as complete
as possible.  Place multiple citations in chronological order.  
\item Do not alter the style template; in particular, do not compress the paper
format by reducing the vertical spaces.
\item Keep your abstract brief and self-contained, one
   paragraph and roughly 4--6 sentences.  Gross violations will require correction at the camera-ready phase.
  Title should have content words capitalized.

\end{itemize}

\subsection{Submitting Papers}

{\bf Paper Deadline:} The deadline for paper submission to ICML 2016
is at \textbf{23:59 Universal Time (3:59 p.m.\ Pacific Standard Time) on February 5, 2016}.
If your full submission does not reach us by this time, it will 
not be considered for publication. There is no separate abstract submission.

{\bf Anonymous Submission:} To facilitate blind review, no identifying
author information should appear on the title page or in the paper
itself.  Section~\ref{author info} will explain the details of how to
format this.

{\bf Simultaneous Submission:} ICML will not accept any paper which,
at the time of submission, is under review for another conference or
has already been published. This policy also applies to papers that
overlap substantially in technical content with conference papers
under review or previously published. ICML submissions must not be
submitted to other conferences during ICML's review period. Authors
may submit to ICML substantially different versions of journal papers
that are currently under review by the journal, but not yet accepted
at the time of submission. Informal publications, such as technical
reports or papers in workshop proceedings which do not appear in
print, do not fall under these restrictions.

\medskip

To ensure our ability to print submissions, authors must provide their
manuscripts in \textbf{PDF} format.  Furthermore, please make sure
that files contain only Type-1 fonts (e.g.,~using the program {\tt
  pdffonts} in linux or using File/DocumentProperties/Fonts in
Acrobat).  Other fonts (like Type-3) might come from graphics files
imported into the document.

Authors using \textbf{Word} must convert their document to PDF.  Most
of the latest versions of Word have the facility to do this
automatically.  Submissions will not be accepted in Word format or any
format other than PDF. Really. We're not joking. Don't send Word.

Those who use \textbf{\LaTeX} to format their accepted papers need to pay close
attention to the typefaces used.  Specifically, when producing the PDF by first
converting the dvi output of \LaTeX\ to Postscript the default behavior is to
use non-scalable Type-3 PostScript bitmap fonts to represent the standard
\LaTeX\ fonts. The resulting document is difficult to read in electronic form;
the type appears fuzzy. To avoid this problem, dvips must be instructed to use
an alternative font map.  This can be achieved with the following two commands:

{\footnotesize
\begin{verbatim}
dvips -Ppdf -tletter -G0 -o paper.ps paper.dvi
ps2pdf paper.ps
\end{verbatim}}
Note that it is a zero following the ``-G''.  This tells dvips to use
the config.pdf file (and this file refers to a better font mapping).

A better alternative is to use the \textbf{pdflatex} program instead of
straight \LaTeX. This program avoids the Type-3 font problem, however you must
ensure that all of the fonts are embedded (use {\tt pdffonts}). If they are
not, you need to configure pdflatex to use a font map file that specifies that
the fonts be embedded. Also you should ensure that images are not downsampled
or otherwise compressed in a lossy way.

Note that the 2016 style files use the {\tt hyperref} package to
make clickable links in documents.  If this causes problems for you,
add {\tt nohyperref} as one of the options to the {\tt icml2016}
usepackage statement.

\subsection{Reacting to Reviews}

We will continue the ICML tradition in which the authors are given the
option of providing a short reaction to the initial reviews. These
reactions will be taken into account in the discussion among the
reviewers and area chairs.

\subsection{Submitting Final Camera-Ready Copy}

The final versions of papers accepted for publication should follow the
same format and naming convention as initial submissions, except of
course that the normal author information (names and affiliations)
should be given.  See Section~\ref{final author} for details of how to
format this.

The footnote, ``Preliminary work.  Under review by the International
Conference on Machine Learning (ICML).  Do not distribute.'' must be
modified to ``\textit{Proceedings of the
$\mathit{33}^{rd}$ International Conference on Machine Learning},
New York, NY, USA, 2016.  JMLR: W\&CP volume 48. 
Copyright 2016 by the author(s).''

For those using the \textbf{\LaTeX} style file, simply change
$\mathtt{\backslash usepackage\{icml2016\}}$ to

$$\mathtt{\backslash usepackage[accepted]\{icml2016\}}$$

\noindent
Authors using \textbf{Word} must edit the
footnote on the first page of the document themselves.

Camera-ready copies should have the title of the paper as running head
on each page except the first one.  The running title consists of a
single line centered above a horizontal rule which is $1$ point thick.
The running head should be centered, bold and in $9$ point type.  The
rule should be $10$ points above the main text.  For those using the
\textbf{\LaTeX} style file, the original title is automatically set as running
head using the {\tt fancyhdr} package which is included in the ICML
2016 style file package.  In case that the original title exceeds the
size restrictions, a shorter form can be supplied by using

\verb|\icmltitlerunning{...}|

just before $\mathtt{\backslash begin\{document\}}$.
Authors using \textbf{Word} must edit the header of the document themselves.

\section{Format of the Paper}

All submissions must follow the same format to ensure the printer can
reproduce them without problems and to let readers more easily find
the information that they desire.

\subsection{Length and Dimensions}

Papers must not exceed eight (8) pages, including all figures, tables,
and appendices, but excluding references and acknowledgements. When references and acknowledgements are included,
the paper must not exceed ten (10) pages.
Acknowledgements should be limited to grants and people who contributed to the paper.
Any submission that exceeds 
this page limit or that diverges significantly from the format specified 
herein will be rejected without review.

The text of the paper should be formatted in two columns, with an
overall width of 6.75 inches, height of 9.0 inches, and 0.25 inches
between the columns. The left margin should be 0.75 inches and the top
margin 1.0 inch (2.54~cm). The right and bottom margins will depend on
whether you print on US letter or A4 paper, but all final versions
must be produced for US letter size.

The paper body should be set in 10~point type with a vertical spacing
of 11~points. Please use Times typeface throughout the text.

\subsection{Title}

The paper title should be set in 14~point bold type and centered
between two horizontal rules that are 1~point thick, with 1.0~inch
between the top rule and the top edge of the page. Capitalize the
first letter of content words and put the rest of the title in lower
case.

\subsection{Author Information for Submission}
\label{author info}

To facilitate blind review, author information must not appear.  If
you are using \LaTeX\/ and the \texttt{icml2016.sty} file, you may use
\verb+\icmlauthor{...}+ to specify authors.  The author information
will simply not be printed until {\tt accepted} is an argument to the
style file. Submissions that include the author information will not
be reviewed.

\subsubsection{Self-Citations}

If your are citing published papers for which you are an author, refer
to yourself in the third person. In particular, do not use phrases
that reveal your identity (e.g., ``in previous work \cite{langley00}, we 
have shown \ldots'').

Do not anonymize citations in the reference section by removing or
blacking out author names. The only exception are manuscripts that are
not yet published (e.g. under submission). If you choose to refer to
such unpublished manuscripts \cite{anonymous}, anonymized copies have 
to be submitted
as Supplementary Material via CMT. However, keep in mind that an ICML
paper should be self contained and should contain sufficient detail
for the reviewers to evaluate the work. In particular, reviewers are
not required to look a the Supplementary Material when writing their
review.

\subsubsection{Camera-Ready Author Information}
\label{final author}

If a paper is accepted, a final camera-ready copy must be prepared.
%
For camera-ready papers, author information should start 0.3~inches
below the bottom rule surrounding the title. The authors' names should
appear in 10~point bold type, electronic mail addresses in 10~point
small capitals, and physical addresses in ordinary 10~point type.
Each author's name should be flush left, whereas the email address
should be flush right on the same line. The author's physical address
should appear flush left on the ensuing line, on a single line if
possible. If successive authors have the same affiliation, then give
their physical address only once.

A sample file (in PDF) with author names is included in the ICML2016 
style file package.

\subsection{Abstract}

The paper abstract should begin in the left column, 0.4~inches below
the final address. The heading `Abstract' should be centered, bold,
and in 11~point type. The abstract body should use 10~point type, with
a vertical spacing of 11~points, and should be indented 0.25~inches
more than normal on left-hand and right-hand margins. Insert
0.4~inches of blank space after the body. Keep your abstract brief and 
self-contained,
limiting it to one paragraph and roughly 4--6 sentences.  Gross violations will require correction at the camera-ready phase.

\subsection{Partitioning the Text}

You should organize your paper into sections and paragraphs to help
readers place a structure on the material and understand its
contributions.

\subsubsection{Sections and Subsections}

Section headings should be numbered, flush left, and set in 11~pt bold
type with the content words capitalized. Leave 0.25~inches of space
before the heading and 0.15~inches after the heading.

Similarly, subsection headings should be numbered, flush left, and set
in 10~pt bold type with the content words capitalized. Leave
0.2~inches of space before the heading and 0.13~inches afterward.

Finally, subsubsection headings should be numbered, flush left, and
set in 10~pt small caps with the content words capitalized. Leave
0.18~inches of space before the heading and 0.1~inches after the
heading.

Please use no more than three levels of headings.

\subsubsection{Paragraphs and Footnotes}

Within each section or subsection, you should further partition the
paper into paragraphs. Do not indent the first line of a given
paragraph, but insert a blank line between succeeding ones.

You can use footnotes\footnote{For the sake of readability, footnotes
should be complete sentences.} to provide readers with additional
information about a topic without interrupting the flow of the paper. 
Indicate footnotes with a number in the text where the point is most
relevant. Place the footnote in 9~point type at the bottom of the
column in which it appears. Precede the first footnote in a column
with a horizontal rule of 0.8~inches.\footnote{Multiple footnotes can
appear in each column, in the same order as they appear in the text,
but spread them across columns and pages if possible.}

\begin{figure}[ht]
\vskip 0.2in
\begin{center}
\centerline{\includegraphics[width=\columnwidth]{icml_numpapers}}
\caption{Historical locations and number of accepted papers for International
  Machine Learning Conferences (ICML 1993 -- ICML 2008) and
  International Workshops on Machine Learning (ML 1988 -- ML
  1992). At the time this figure was produced, the number of
  accepted papers for ICML 2008 was unknown and instead estimated.}
\label{icml-historical}
\end{center}
\vskip -0.2in
\end{figure}

\subsection{Figures}

You may want to include figures in the paper to help readers visualize
your approach and your results. Such artwork should be centered,
legible, and separated from the text. Lines should be dark and at
least 0.5~points thick for purposes of reproduction, and text should
not appear on a gray background.

Label all distinct components of each figure. If the figure takes the
form of a graph, then give a name for each axis and include a legend
that briefly describes each curve. Do not include a title inside the
figure; instead, the caption should serve this function.

Number figures sequentially, placing the figure number and caption
{\it after\/} the graphics, with at least 0.1~inches of space before
the caption and 0.1~inches after it, as in
Figure~\ref{icml-historical}.  The figure caption should be set in
9~point type and centered unless it runs two or more lines, in which
case it should be flush left.  You may float figures to the top or
bottom of a column, and you may set wide figures across both columns
(use the environment {\tt figure*} in \LaTeX), but always place
two-column figures at the top or bottom of the page.

\subsection{Algorithms}

If you are using \LaTeX, please use the ``algorithm'' and ``algorithmic'' 
environments to format pseudocode. These require 
the corresponding stylefiles, algorithm.sty and 
algorithmic.sty, which are supplied with this package. 
Algorithm~\ref{alg:example} shows an example.

\begin{algorithm}[tb]
   \caption{Bubble Sort}
   \label{alg:example}
\begin{algorithmic}
   \STATE {\bfseries Input:} data $x_i$, size $m$
   \REPEAT
   \STATE Initialize $noChange = true$.
   \FOR{$i=1$ {\bfseries to} $m-1$}
   \IF{$x_i > x_{i+1}$} 
   \STATE Swap $x_i$ and $x_{i+1}$
   \STATE $noChange = false$
   \ENDIF
   \ENDFOR
   \UNTIL{$noChange$ is $true$}
\end{algorithmic}
\end{algorithm}

\subsection{Tables}

You may also want to include tables that summarize material. Like 
figures, these should be centered, legible, and numbered consecutively. 
However, place the title {\it above\/} the table with at least 
0.1~inches of space before the title and the same after it, as in 
Table~\ref{sample-table}. The table title should be set in 9~point 
type and centered unless it runs two or more lines, in which case it
should be flush left.

% Note use of \abovespace and \belowspace to get reasonable spacing 
% above and below tabular lines.

\begin{table}[t]
\caption{Classification accuracies for naive Bayes and flexible 
Bayes on various data sets.}
\label{sample-table}
\vskip 0.15in
\begin{center}
\begin{small}
\begin{sc}
\begin{tabular}{lcccr}
\hline
\abovespace\belowspace
Data set & Naive & Flexible & Better? \\
\hline
\abovespace
Breast    & 95.9$\pm$ 0.2& 96.7$\pm$ 0.2& $\surd$ \\
Cleveland & 83.3$\pm$ 0.6& 80.0$\pm$ 0.6& $\times$\\
Glass2    & 61.9$\pm$ 1.4& 83.8$\pm$ 0.7& $\surd$ \\
Credit    & 74.8$\pm$ 0.5& 78.3$\pm$ 0.6&         \\
Horse     & 73.3$\pm$ 0.9& 69.7$\pm$ 1.0& $\times$\\
Meta      & 67.1$\pm$ 0.6& 76.5$\pm$ 0.5& $\surd$ \\
Pima      & 75.1$\pm$ 0.6& 73.9$\pm$ 0.5&         \\
\belowspace
Vehicle   & 44.9$\pm$ 0.6& 61.5$\pm$ 0.4& $\surd$ \\
\hline
\end{tabular}
\end{sc}
\end{small}
\end{center}
\vskip -0.1in
\end{table}

Tables contain textual material that can be typeset, as contrasted 
with figures, which contain graphical material that must be drawn. 
Specify the contents of each row and column in the table's topmost
row. Again, you may float tables to a column's top or bottom, and set
wide tables across both columns, but place two-column tables at the
top or bottom of the page.

\subsection{Citations and References}

Please use APA reference format regardless of your formatter
or word processor. If you rely on the \LaTeX\/ bibliographic 
facility, use {\tt natbib.sty} and {\tt icml2016.bst} 
included in the style-file package to obtain this format.

Citations within the text should include the authors' last names and
year. If the authors' names are included in the sentence, place only
the year in parentheses, for example when referencing Arthur Samuel's
pioneering work \yrcite{Samuel59}. Otherwise place the entire
reference in parentheses with the authors and year separated by a
comma \cite{Samuel59}. List multiple references separated by
semicolons \cite{kearns89,Samuel59,mitchell80}. Use the `et~al.'
construct only for citations with three or more authors or after
listing all authors to a publication in an earlier reference \cite{MachineLearningI}.

Authors should cite their own work in the third person
in the initial version of their paper submitted for blind review.
Please refer to Section~\ref{author info} for detailed instructions on how to
cite your own papers.

Use an unnumbered first-level section heading for the references, and 
use a hanging indent style, with the first line of the reference flush
against the left margin and subsequent lines indented by 10 points. 
The references at the end of this document give examples for journal
articles \cite{Samuel59}, conference publications \cite{langley00}, book chapters \cite{Newell81}, books \cite{DudaHart2nd}, edited volumes \cite{MachineLearningI}, 
technical reports \cite{mitchell80}, and dissertations \cite{kearns89}.

Alphabetize references by the surnames of the first authors, with
single author entries preceding multiple author entries. Order
references for the same authors by year of publication, with the
earliest first. Make sure that each reference includes all relevant
information (e.g., page numbers).

\subsection{Software and Data}

We strongly encourage the publication of software and data with the
camera-ready version of the paper whenever appropriate.  This can be
done by including a URL in the camera-ready copy.  However, do not
include URLs that reveal your institution or identity in your
submission for review.  Instead, provide an anonymous URL or upload
the material as ``Supplementary Material'' into the CMT reviewing
system.  Note that reviewers are not required to look a this material
when writing their review.

% Acknowledgements should only appear in the accepted version. 
\section*{Acknowledgements}

\textbf{Do not} include acknowledgements in the initial version of
the paper submitted for blind review.

If a paper is accepted, the final camera-ready version can (and
probably should) include acknowledgements. In this case, please
place such acknowledgements in an unnumbered section at the
end of the paper. Typically, this will include thanks to reviewers
who gave useful comments, to colleagues who contributed to the ideas, 
and to funding agencies and corporate sponsors that provided financial 
support.

% In the unusual situation where you want a paper to appear in the
% references without citing it in the main text, use \nocite
\nocite{langley00}

\bibliography{example_paper}
\bibliographystyle{icml2016}

\end{document}

% This document was modified from the file originally made available by
% Pat Langley and Andrea Danyluk for ICML-2K. This version was
% created by Lise Getoor and Tobias Scheffer, it was slightly modified  
% from the 2010 version by Thorsten Joachims & Johannes Fuernkranz, 
% slightly modified from the 2009 version by Kiri Wagstaff and 
% Sam Roweis's 2008 version, which is slightly modified from 
% Prasad Tadepalli's 2007 version which is a lightly 
% changed version of the previous year's version by Andrew Moore, 
% which was in turn edited from those of Kristian Kersting and 
% Codrina Lauth. Alex Smola contributed to the algorithmic style files.

### 2016-02-20 icml2016/paper.tex
changed diag to log diag

@@
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%% ICML 2016 EXAMPLE LATEX SUBMISSION FILE %%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\documentclass[]{article}
+\usepackage{proceed2e}
+% Set the typeface to Times Roman
+\usepackage{times}
 
-% Use the following line _only_ if you're still using LaTeX 2.09.
-%\documentstyle[icml2016,epsf,natbib]{article}
-% If you rely on Latex2e packages, like most moden people use this:
-\documentclass{article}
-
-% use Times
-\usepackage{times}
 % For figures
@@
 \newcommand{\theHalgorithm}{\arabic{algorithm}}
+\newcommand{\Ex}{\mathbb{E}}
+\newcommand{\KL}{\textrm{KL}}
+% I would prefer f or \mathbf{f} rather than F()
+\newcommand{\f}{f}
 
-% Employ the following version of the ``usepackage'' statement for
-% submitting the draft version of the paper for review.  This will set
-% the note in the first column to ``Under review.  Do not distribute.''
-\usepackage{icml2016}
+% Just a working title/suggestion
+\title{Blackbox variational inference using \\ the Gaussian process latent variable model}
 
-% Employ this version of the ``usepackage'' statement after the paper has
-% been accepted, when creating the final version.  This will set the
-% note in the first column to ``Proceedings of the...''
-%\usepackage[accepted]{icml2016}
+%\author{} % LEAVE BLANK FOR ORIGINAL SUBMISSION.
+          % UAI  reviewing is double-blind.
 
-% The \icmltitle you define below is probably too long as a header.
-% Therefore, a short form for the running title is supplied here:
-\icmltitlerunning{Submission and Formatting Instructions for ICML 2016}
+\author{ {\bf Rory Beard} \\
+\And
+{\bf Chris Lloyd}  \\
+\And
+{\bf Stephen J. Roberts}   \\
+\AND
+\textnormal{Machine Learning Research Group} \\
+\textnormal{Department of Engineering Science}\\
+\textnormal{University of Oxford} \\
+\url{{rbeard, clloyd, sjrob}@robots.ox.ac.uk}
+}
 
@@
 
-\twocolumn[
-\icmltitle{Submission and Formatting Instructions for \\
-           International Conference on Machine Learning (ICML 2016)}
-
-% It is OKAY to include author information, even for blind
-% submissions: the style file will automatically remove it for you
-% unless you've provided the [accepted] option to the icml2016
-% package.
-\icmlauthor{Your Name}{email@yourdomain.edu}
-\icmladdress{Your Fantastic Institute,
-            314159 Pi St., Palo Alto, CA 94306 USA}
-\icmlauthor{Your CoAuthor's Name}{email@coauthordomain.edu}
-\icmladdress{Their Fantastic Institute,
-            27182 Exp St., Toronto, ON M6H 2T1 CANADA}
-
-% You may provide any keywords that you
-% find helpful for describing your paper; these are used to populate
-% the "keywords" metadata in the PDF but will not be shown in the document
-\icmlkeywords{boring formatting information, machine learning, ICML}
-
-\vskip 0.3in
-]
+\maketitle
 
@@
 The purpose of this document is to provide both the basic paper template and
-submission guidelines. Abstracts should be a single paragraph, between 4--6 sentences long, ideally.  Gross violations will trigger corrections at the camera-ready phase.
+submission guidelines. Abstracts should be a single paragraph, between 4--6 sentences long, ideally.
+Gross violations will trigger corrections at the camera-ready phase.
 \end{abstract}
 
-\section{Introduction}
-\label{sec:intro}
+\section{Introduction}\label{sec:intro}
 
@@
 SVI framework permits a learning procedure that requires only unbiased estimates of the lower bound (or rather its gradient), removing the issue of the lack of analyticity of the lower bound and, for no extra cost, granting a highly data-scalable mini-batch update scheme.
-There has been much recent work which empirically demonstrates the success of ...
-despite the presence of two sources of stochasticity.
+There has been much recent work which empirically demonstrates the success of ... despite the presence of two sources of stochasticity.
 
@@
 
-l;asdf.
-We additionally introduce a kernel based approach to back-constrained amortization in which random functions are trained to capture the encoding process.
+l;asdf. We additionally introduce a kernel based approach to back-constrained amortization in which random functions are trained to capture the encoding process.
 WHY IS THIS GOOD: provable guarantees of kernel mappings, ameliorates pathological curvature??
@@
 
-\section{Background principles}
-\label{sec:background}
+\section{Background principles}\label{sec:background}
 
-\subsection{Variational Inference}
-\label{subsec:vi}
+\subsection{Variational Inference}\label{subsec:vi}
 
@@
 
-\subsection{The Bayesian GP-LVM}
-\label{subsec:gplvm}
+\subsection{The Bayesian GP-LVM}\label{subsec:gplvm}
 
@@
   p(Y|X) &= \prod_{d=1}^D p(\mathbf{y}_d | X),\\
-  p(\mathbf{y}_d) &= \mathcal{N}(\mathbf{y}_d|\mathbf{0}, K_{ff}+\beta^{-1}I)
+  p(\mathbf{y}_d|X) &= \mathcal{N}(\mathbf{y}_d|\mathbf{0}, K_{ff}+\beta^{-1}I)
 \end{align}
@@
 
-\section{GP-LVM-Conditional Mean Field Variational Inference} % (fold)
-\label{sec:gplvmcmf}
+\section{GP-LVM-Conditional Mean Field Variational Inference} \label{sec:gplvmcmf}
 
@@
 \begin{align}
-  \log p_\theta(Y) &= \mathbb{E}_{q_\phi(Z)}[\log p_\theta(Y|Z) - \log q_\phi(Z)]\notag\\
+\log p_\theta(Y) &= \Ex_{q_\phi(Z)}[\log p_\theta(Y|Z) - \log q_\phi(Z)]\notag\\
 &+ \text{KL}[q_\phi(Z)\Vert p_\theta(Z|Y)]\label{eq:var_principle}\\
-  &\geq \mathbb{E}_{q_\phi(Z)}[\log p_\theta(Y|Z) - \log q_\phi(Z)] =: \mathcal{L}_1\label{eq:bound1}
+&\geq \Ex_{q_\phi(Z)}[\log p_\theta(Y|Z) - \log q_\phi(Z)] \notag \\
+&\triangleq \mathcal{L}_1 \label{eq:bound1}
 \end{align}
@@
 $\phi$), and we denote the set of all latent variables and parameters subject to inference by $Z$ (though later we will use $\eta$ to ...).
- Integrating \ref{eq:bound1} with respect to the auxiliary variables, noting that the marginal variational distribution over $Z$ can be expressed as $q(Z) = q(Z|F(),U,X)q(F(),U,X)/q(F(),U,X|Z)$, and introducing an auxiliary model $r_\psi(F(),U,X)$, we arrive at the following \emph{auxiliary lower bound}:
+Integrating \ref{eq:bound1} with respect to the auxiliary variables, noting that the marginal variational distribution over $Z$ can be expressed as 
+\begin{equation}
+q(Z) = \frac{q(Z|\f,U,X)q(\f,U,X)}{q(\f,U,X|Z)}
+\end{equation}
+%
+and introducing an auxiliary model $r_\psi(\f,U,X)$, we arrive at the following \emph{auxiliary lower bound}:
 %
 \begin{align}
-  \mathcal{L}_1 &\geq \mathbb{E}_{q(Z,F(),U,X)}\big[ \log p(Y|Z)p(Z)r(F(),U,X|Z)\notag\\
-   &- \log q(Z|F(),X)q(F(),U,X) \big]\label{eq:lowerlower}\\
-    &+ \mathbb{E}_{q(Z)}\big[ \text{KL}[q(F(),U,X|Z)\Vert r(F(),U,X|Z)] \big] =: \mathcal{L}_{aux}.\label{eq:L_aux}
+\mathcal{L}_1 &\geq \Ex_{q(Z,\f,U,X)}\big[ \log p(Y|Z)p(Z)r(\f,U,X|Z) \notag \\
+&- \log q(Z|\f,X)q(\f,U,X) \big] \notag \\
+&+ \Ex_{q(Z)}\big[\KL[q(\f,U,X|Z)\Vert r(\f,U,X|Z)]\big] \label{eq:lowerlower} \\
+&\triangleq \mathcal{L}_{aux}. \label{eq:L_aux}
 \end{align}
 %
-To explain this bound, we find it useful to refer to $q(F(),U,X)$ (resp.
-$q(F(),U,X|Z)$) as the \emph{variational prior} (resp.) \emph{variational posterior} distributions over the auxiliary inference variables, the latter of which constitutes a posterior belief over the auxiliary variables having observed a (set of) sample(s), $Z$, generated under the measure $q$ endowed by the inference network.
+To explain this bound, we find it useful to refer to $q(\f,U,X)$ (resp.
+$q(\f,U,X|Z)$) as the \emph{variational prior} (resp.) \emph{variational posterior} distributions over the auxiliary inference variables, the latter of which constitutes a posterior belief over the auxiliary variables having observed a (set of) sample(s), $Z$, generated under the measure $q$ endowed by the inference network.
 
@@
 
-\subsection{Scalable and Amortised Inference}
-\label{subsec:scale}
+\subsection{Scalable and Amortised Inference}\label{subsec:scale}
 
@@
 \begin{align}
-  \mathcal{L}_{aux} &= \mathbb{E}_{q(Z)}[ \log p(Y|Z) ] - \mathbb{E}_{q(F(),U,X)}[ \text{KL}[q(Z|F(),U,X)\Vert p(Z)] ]\notag\\
-  &- \mathbb{E}_{q(Z)}[ \text{KL}[q(F(),U,X)\Vert r(F(),U,X|Z)] ].\label{eq:L_aux_ABC}.
+\mathcal{L}_{aux} &= \Ex_{q(Z)}[ \log p(Y|Z) ] \notag \\
+&- \Ex_{q(\f,U,X)}[ \KL[q(Z|\f,U,X)\Vert p(Z)] ]\notag\\
+&- \Ex_{q(Z)}[ \KL[q(\f,U,X)\Vert r(\f,U,X|Z)] ].\label{eq:L_aux_ABC}
 \end{align}
 %
-As noted by Kingma and Welling, in the case of Gaussian $p(Z)$ and $q(Z|F(),U,X)$, the first K-L divergence can be evaluated analytically and, but the lower lower means we can get a second analytical term...
+As noted by Kingma and Welling, in the case of Gaussian $p(Z)$ and $q(Z|\f,U,X)$, the first K-L divergence can be evaluated analytically and, but the lower lower means we can get a second analytical term...
 This was also noticed by Tran et al., in their lower lower bound...
 
-Variational compression ...
-sparse variational GPs ...
-set $r(F()|U,X,Z):=q(F()|U,X)$ which means that $r(F(),U,X|Z) = q(F()|U,X)r(U,X|Z)$.
-Full implications of this???
-Reduces configuration space...
-The aptness of this hinges on the veracity of the assumption that ...
-are sufficient statistics for ...
-in expectation under ...
+Variational compression ... sparse variational GPs ... set $r(\f|U,X,Z):=q(\f|U,X)$ which means that $r(\f,U,X|Z) = q(\f|U,X)r(U,X|Z)$.
+Full implications of this??? Reduces configuration space...
+The aptness of this hinges on the veracity of the assumption that ... are sufficient statistics for ... in expectation under ...
 
@@
 
-\subsection{Design Choices}
-\label{subsec:choices}
+\subsection{Design Choices}\label{subsec:choices}
 
@@
 
-\section{Related Work}
-\label{sec:related}
+\section{Related Work}\label{sec:related}
 
@@
 popularised the modern approach of autoencoding in the variational framework, and spurred on a great research drive in variational approaches to deep learning.
-Highly sophisticated models such as DRAW - a method to....
-- have at their core the VAE...
-Grosse et al considered an alternative variational lower bound based on importance weights to train a VAE.
+Highly sophisticated models such as DRAW - a method to.... - have at their core the VAE... Grosse et al considered an alternative variational lower bound based on importance weights to train a VAE.
 However in all these approaches, the
@@
 Salimans and Welling considered a Markov chain inference network in which the final ($T^th$) Markov transition in the chain produced the latent variable (sample), such that previous $T-1$ variables were treated as auxiliary variables.
-Collapsing over this expanded auxiliary space then resulted in a rich yet intractable mixture...
-and tractability was recovered through the use of an auxiliary distribution also of a Markov structure.
+Collapsing over this expanded auxiliary space then resulted in a rich yet intractable mixture... and tractability was recovered through the use of an auxiliary distribution also of a Markov structure.
 
@@
 This leads to coupling whereby each GP tends to be positive in similar regions ?????.
-In our approach the coupling is induced via the shared inducing coordinates...
-more faithfully to the spirit of the GP-LVM.
+In our approach the coupling is induced via the shared inducing coordinates... more faithfully to the spirit of the GP-LVM.
 
-\section{Experimental Setup}
-\label{sec:experiments}
+\section{Experimental Setup}\label{sec:experiments}
 
-...
-free form q(X) means having to optimise the auxiliary latent coordinate of the test location just to be able to evaluate.
+... free form q(X) means having to optimise the auxiliary latent coordinate of the test location just to be able to evaluate.
 Including an optimisation procedure in the predictive step is an interesting idea, but was avoided here for lack of space and clarity of discussion, so investigation was deferred to a future study.
 
-\section{Discussion and Future Work}
-\label{sec:discussion}
+\section{Discussion and Future Work}\label{sec:discussion}
 
@@
 
+change!!!!@
+
+\bibliography{paper.bib}{}
+\bibliographystyle{plainnat}
+
 \end{document}

### 2016-02-20 GP_LVM_CMF.py

@@ -103,7 +103,7 @@ class SGPDV(object):
 
         self.lowerBound = -np.inf  # Lower bound
 
-        self.numberofBatchesPerEpoch = int(np.ceil(self.N / float(self.B)))
+        self.numberofBatchesPerEpoch = int(np.ceil(np.float32(self.N) / self.B))
         numPad = self.numberofBatchesPerEpoch * self.B - self.N
 
         self.batchStream = srng.permutation(n=self.N)
@@ -167,10 +167,10 @@ class SGPDV(object):
             if encoderType_qX == 'FreeForm1':
 
                 self.Phi_full_sqrt = sharedZeroMatrix(self.N, self.N, 'Phi_full_sqrt')
-                Phi_batch_sqrt = self.Phi_full_sqrt[self.currentBatch][:, self.currentBatch]
+                Phi_batch_sqrt = self.Phi_full_sqrt[self.currentBatch][:, self.currentBatch]                
                 self.Phi = T.dot(Phi_batch_sqrt, Phi_batch_sqrt.T)
 
-                Phi_batch_sqrt.name = 'Phi_batch_sqrt'
+                Phi_batch_sqrt.name = 'Phi_batch_sqrt'                
                 self.Phi.name = 'Phi'
 
                 (self.cPhi, self.iPhi, self.logDetPhi) = cholInvLogDet(self.Phi, self.B, self.jitter)
@@ -178,16 +178,16 @@ class SGPDV(object):
                 self.qX_vars = [self.Phi_full_sqrt, self.phi_full]
 
             else:
-
-                self.Phi_full_diag = sharedZeroArray(self.N, 'Phi_full_diag')
+                
+                self.Phi_full_logdiag = sharedZeroArray(self.N, 'Phi_full_logdiag')
 
                 Phi_batch_diag = self.Phi_full_diag[self.currentBatch]
-                Phi_batch_diag.name = 'Phi_batch_diag'
+                Phi_batch_diag.name = 'Phi_batch_logdiag'
 
                 (self.Phi, self.iPhi, self.cPhi, self.logDetPhi) \
-                    = diagCholInvLogDet_fromDiag(Phi_batch_diag, 'Phi')
+                    = diagCholInvLogDet_fromLogDiag(Phi_batch_diag, 'Phi')
 
-                self.qX_vars = [self.Phi_full_diag, self.phi_full]
+                self.qX_vars = [self.Phi_full_logdiag, self.phi_full]
 
         elif self.encoderType_qX == 'MLP':
 
[…]

### 2016-02-20 Auto_encoder_model.py
Maths bug in KL_qp fixed

@@ -13,7 +13,6 @@ from theano.tensor import nlinalg
 from GP_LVM_CMF import SGPDV
 from testTools import checkgrad
 from utils import sharedZeroMatrix, sharedZeroVector, log_mean_exp_stable
-from theano.compile.nanguardmode import NanGuardMode
 
 precision = th.config.floatX
 
@@ -29,7 +28,7 @@ class VA(SGPDV):
             encoderType_qX='FreeForm2',  # 'FreeForm1', 'FreeForm2','MLP', 'Kernel'.
             encoderType_qu='Kernel',    # 'Kernel', 'MLP'
             encoderType_rX='FreeForm2',  # 'FreeForm1', 'FreeForm2', 'MLP', 'Kernel', 'NoEncoding'.
-            encoderType_ru='FreeForm2',  # 'FreeForm', 'MLP', 'NoEncoding'
+            encoderType_ru='FreeForm',  # 'FreeForm', 'MLP', 'NoEncoding'
             Xu_optimise=False,
             numHiddenUnits_encoder=0,
             numHiddentUnits_decoder=10,
@@ -110,10 +109,11 @@ class VA(SGPDV):
     def KL_qp(self):
 
         if self.continuous:
-            Kuf_Kfu_iKuu = T.dot(self.Kfu.T, T.dot(self.Kfu, self.iKuu))
+            iKuu_Kuf_Kfu_iKuu = T.dot(self.iKuu, T.dot(self.Kfu.T, T.dot(self.Kfu, self.iKuu)))
+            kappa_outer = T.dot(self.kappa.T, self.kappa)
             KL = -0.5*self.B*self.Q*(1 + T.exp(self.log_sigma)**2 - 2*self.log_sigma) \
-                 +0.5*nlinalg.trace(T.dot( self.iKuu, T.dot( Kuf_Kfu_iKuu, (T.dot(self.kappa.T, self.kappa) + self.iKuu) ) )) \
-                 +0.5*self.Q*( nlinalg.trace(self.Kff) - nlinalg.trace(Kuf_Kfu_iKuu) )
+                 +0.5*nlinalg.trace(T.dot(iKuu_Kuf_Kfu_iKuu, kappa_outer)) \
+                 +0.5*self.Q*(nlinalg.trace(self.Kff))
         else:
             raise RuntimeError("Case not implemented")
 

### 2016-02-20 GP_LVM_CMF.py
Minor bug

@@ -371,7 +371,7 @@ class SGPDV(object):
 
             self.tau  = mu_rX.T
             (self.Tau, self.cTau, self.iTau, self.logDetTau) \
-                = diagCholInvLogDet_fromLogDiag(log_sigma_rX)
+                = diagCholInvLogDet_fromLogDiag(log_sigma_rX, 'Tau')
 
             self.rX_vars = [self.W1_rX, self.W2_rX, self.W3_rX, self.b1_rX, self.b2_rX, self.b3_rX]
 

### 2016-02-20 GP_LVM_CMF.py
Randomise W4

@@ -493,7 +493,8 @@ class SGPDV(object):
                 pass
             elif var.name.startswith('W1') or \
                     var.name.startswith('W2') or \
-                    var.name.startswith('W3'):
+                    var.name.startswith('W3') or \
+                    var.name.startswith('W4'):
                 print 'Randomising ' + var.name
                 # Hidden layer weights are uniformly sampled from a symmetric interval
                 # following [Xavier, 2010]
@@ -508,7 +509,8 @@ class SGPDV(object):
 
             elif var.name.startswith('b1') or \
                     var.name.startswith('b2') or \
-                    var.name.startswith('b3'):
+                    var.name.startswith('b3') or \
+                    var.name.startswith('b4'):
                 print 'Setting ' + var.name + ' to all 0s'
                 # Offsets not randomised at all
                 var.set_value(np.zeros(var.get_value().shape, dtype=precision))

### 2016-02-22 GP_LVM_CMF.py
Half done major changes

@@ -67,7 +67,6 @@ class SGPDV(object):
                  data,                   # [NxP] matrix of observations
                  kernelType='RBF',
                  encoderType_qX='FreeForm2',  # 'FreeForm1', 'FreeForm2' 'MLP', 'Kernel'.
-                 encoderType_qu='FreeForm',   # 'FreeForm', 'MLP'
                  encoderType_rX='FreeForm2',  # 'FreeForm1', 'FreeForm2', 'MLP', 'Kernel', 'NoEncoding'.
                  encoderType_ru='FreeForm2',  # 'FreeForm1', 'FreeForm2', 'MLP', 'NoEncoding'
                  Xu_optimise=False,
@@ -89,7 +88,7 @@ class SGPDV(object):
         self.encoderType_qX = encoderType_qX
         self.encoderType_rX = encoderType_rX
         self.encoderType_ru = encoderType_ru
-        self.encoderType_qu = encoderType_qu        
+    
         self.Xu_optimise = Xu_optimise
 
         self.y = th.shared(data)
@@ -170,10 +169,9 @@ class SGPDV(object):
                 self.Phi_full_sqrt = sharedZeroMatrix(self.N, self.N, 'Phi_full_sqrt')
                 
                 Phi_batch_sqrt = self.Phi_full_sqrt[self.currentBatch][:, self.currentBatch]
+                Phi_batch_sqrt.name = 'Phi_batch_sqrt'
                 
                 self.Phi = Tdot(Phi_batch_sqrt, Phi_batch_sqrt.T, 'Phi')
-
-                Phi_batch_sqrt.name = 'Phi_batch_sqrt'
 
                 (self.cPhi, self.iPhi, self.logDetPhi) = cholInvLogDet(self.Phi, self.B, 0)
 
@@ -230,6 +228,14 @@ class SGPDV(object):
         else:
             raise RuntimeError('Unrecognised encoding for q(X): ' + self.encoderType_qX)
 
+        # Variational distribution q(u)
+        self.kappa = sharedZeroMatrix(self.Q, self.M, 'kappa')
+        self.Kappa_sqrt = sharedZeroMatrix(self.M, self.M, 'Kappa_sqrt')
+        self.Kappa = Tdot(self.Kappa_sqrt, self.Kappa_sqrt.T, 'Kappa')
[…]

### 2016-02-22 trainmnist_cond_MF.py

@@ -82,9 +82,9 @@ va.randomise()
 va.init_Xu_from_Xf()
 
 
-va.printMemberTypes()
-va.printSharedVariables()
-va.printTheanoVariables()
+#va.printMemberTypes()
+#va.printSharedVariables()
+#va.printTheanoVariables()
 
 
 va.constructUpdateFunction()

### 2016-02-23 Auto_encoder_model.py
debugged major re-write

@@ -12,7 +12,7 @@ from theano.tensor import nlinalg
 
 from GP_LVM_CMF import SGPDV
 from testTools import checkgrad
-from utils import log_mean_exp_stable, Tdot, nlinalg_trace
+from utils import log_mean_exp_stable, dot, trace, softplus
 
 precision = th.config.floatX
 
@@ -28,7 +28,7 @@ class VA(SGPDV):
             encoderType_qX='FreeForm2',  # MLP', 'Kernel'.
             encoderType_rX='FreeForm2',  # MLP', 'Kernel'.
             Xu_optimise=False,
-            numHiddenUnits_encoder=0,
+            numHiddenUnits_encoder=10,
             numHiddentUnits_decoder=10,
             continuous=True
         ):
@@ -84,8 +84,8 @@ class VA(SGPDV):
 
         if self.continuous:
 
-            h_decoder  = T.nnet.softplus(T.dot(self.W1,self.z) + self.b1)
-            mu_decoder = T.dot(self.W2, h_decoder) + self.b2
+            h_decoder  = softplus(T.dot(self.W1,self.z.T) + self.b1)
+            mu_decoder = dot(self.W2, h_decoder) + self.b2
             log_sigma_decoder = 0.5*(T.dot(self.W3, h_decoder) + self.b3)
             log_pyz    = T.sum( -(0.5 * np.log(2 * np.pi) + log_sigma_decoder) \
                                 - 0.5 * ((self.y_miniBatch.T - mu_decoder) / T.exp(log_sigma_decoder))**2 )
@@ -111,10 +111,10 @@ class VA(SGPDV):
     def KL_qp(self):
 
         if self.continuous:
-            kappa_outer = Tdot(self.kappa.T, self.kappa, 'kappa_outer')
+            kappa_outer = dot(self.kappa, self.kappa.T, 'kappa_outer')
             AtA = dot(self.A.T, self.A, 'A''.A' )
-            0.5*self.Q*trace(self.L) + 0.5*trace(dot(AtA,kappa_outer) \
-                + 0.5*self.Q+trace(dot(AtA,self.Kappa)) - 0.5*self.Q*self.B - 0.5*self.Q*self.logDetL 
[…]

### 2016-02-23 GP_LVM_CMF.py

@@ -11,7 +11,7 @@ import collections
 
 from optimisers import Adam
 from utils import cholInvLogDet, sharedZeroArray, sharedZeroMatrix, sharedZeroVector, \
-     np_log_mean_exp_stable, diagCholInvLogDet_fromLogDiag, diagCholInvLogDet_fromDiag, dot, minus, plus, mul
+     np_log_mean_exp_stable, diagCholInvLogDet_fromLogDiag, diagCholInvLogDet_fromDiag, dot, minus, plus, mul, softplus, sigmoid, trace
 
 # precision = np.float64
 precision = th.config.floatX
@@ -69,7 +69,7 @@ class SGPDV(object):
                  encoderType_qX='FreeForm2',  # 'MLP', 'Kernel'.
                  encoderType_rX='FreeForm2',  # 'MLP', 'Kernel'
                  Xu_optimise=False,
-                 numberOfEncoderHiddenUnits=0
+                 numberOfEncoderHiddenUnits=10
                  ):
 
         self.numTestSamples = 5000
@@ -86,7 +86,6 @@ class SGPDV(object):
 
         self.encoderType_qX = encoderType_qX
         self.encoderType_rX = encoderType_rX
-        self.encoderType_ru = encoderType_ru
         self.Xu_optimise = Xu_optimise
 
         self.y = th.shared(data)
@@ -136,9 +135,9 @@ class SGPDV(object):
 
         # Random variables
         self.xi    = srng.normal(size=(self.B, self.R), avg=0.0, std=1.0, ndim=None)
-        self.alpha = srng.normal(size=(self.Q, self.M), avg=0.0, std=1.0, ndim=None)
-        self.beta  = srng.normal(size=(self.Q, self.B), avg=0.0, std=1.0, ndim=None)
-        self.eta.name   = 'xi'
+        self.alpha = srng.normal(size=(self.M, self.Q), avg=0.0, std=1.0, ndim=None)
+        self.beta  = srng.normal(size=(self.B, self.Q), avg=0.0, std=1.0, ndim=None)
+        self.xi.name   = 'xi'
         self.alpha.name = 'alpha'
         self.beta.name  = 'beta'
         
@@ -163,7 +162,7 @@ class SGPDV(object):
 
                 self.Phi_full_sqrt = sharedZeroMatrix(self.N, self.N, 'Phi_full_sqrt')
 
[…]

### 2016-02-23 trainmnist_cond_MF.py

@@ -62,9 +62,7 @@ va = VA(
     x_train,                   # [NxP] matrix of observations
     kernelType=kernelType,
     encoderType_qX=encoderType_qX,  # 'FreeForm1', 'FreeForm2', 'MLP', 'Kernel'.
-    encoderType_qu=encoderType_qu,  # 'Kernel'
-    encoderType_rX=encoderType_rX,  # 'FreeForm1', 'FreeForm2', 'MLP', 'Kernel', 'NoEncoding'.
-    encoderType_ru=encoderType_ru,  # 'FreeForm', 'MLP', 'NoEncoding'
+    encoderType_rX=encoderType_rX,  # 'MLP', 'Kernel', 'NoEncoding'.
     Xu_optimise=Xu_optimise,
     numHiddenUnits_encoder=numHiddenUnits_encoder,
     numHiddentUnits_decoder=numHiddentUnits_decoder,

### 2016-02-23 utils.py

@@ -114,68 +114,63 @@ def shared_ones_like(shared_var):
     return th.shared(np.ones(shared_var.get_value(borrow=True).shape).astype(shared_var.dtype),
                          broadcastable=shared_var.broadcastable)
 
-def getname(D):
-    if type(D) == int or type(D) == float:
-        Dname = str(D)
-    elif hasattr(D,'name'):
-        Dname = Dname
+def getname(T):
+    if type(T) == int or type(T) == float:
+        name = str(T)
+    elif hasattr(T, 'name') and not T.name == None:
+        name = T.name
     else:
-        RuntimeError('Could not get name')
-    return Dname
-
-def Tname(A, B, op, name=None):
-    
-    Aname = getname(A)
-    Bname = getname(B)
-
-    if not name == None:
-        pass
-    elif not Aname == None and not Bname == None:
-        name = Aname + op + Bname
-    elif not Aname == None:
-        name = Aname + op + '?'
-    elif not Bname == None:
-        name = '?' + 'op' + Bname
-    else:
-        name = None
+        name = '?'
     return name
+
+def inName(A, B, op, name=None):
+    if name == None:
+        Aname = getname(A)
+        Bname = getname(B)
+        Cname = '(' + Aname + op + Bname + ')'
+    else:
+        Cname = name
+    return Cname
 
 def dot(A, B, name=None):
     C = T.dot(A,B)
-    C.name = Tname(A, B, '.', name)
+    C.name = inName(A, B, ' . ', name)
     return C
 
 def minus(A, B, name=None):
     C = A - B
-    C.name = Tname(A, B, '-', name)
+    C.name = inName(A, B, ' - ', name)
     return C
 
 def plus(A, B, name=None):
     C = A + B
-    C.name = Tname(A, B, '+', name)
+    C.name = inName(A, B, ' + ', name)
     return C
 
 def mul(A, B, name=None):
     C = A * B
-    C.name = Tname(A, B, '*', name)
+    C.name = inName(A, B, ' * ', name)
     return C
 
 def softplus(A, name=None):
-    return T_function(A, T.net.softplus, 'softplus', name)
+    return namedFunction(A, T.nnet.softplus, 'softplus', name)
 
 def sigmoid(A, name=None):
-    return T_function(A, T.net.sigmoid, 'sigmoid', name)
+    return namedFunction(A, T.nnet.sigmoid, 'sigmoid', name)
 
 def trace(A, name=None):
[…]

### 2016-02-23 Auto_encoder_model.py
Fixed bugs so thigs wont break for discrete data, and added option for 2 hidden layers in Auto_encoder_model.

@@ -4,7 +4,7 @@ Created on Thu Feb  4 20:44:40 2016
 
 @author: clloyd
 """
- 
+
 import numpy as np
 import theano as th
 import theano.tensor as T
@@ -31,7 +31,8 @@ class VA(SGPDV):
             encoderType_ru='FreeForm2',  # 'FreeForm', 'MLP', 'NoEncoding'
             Xu_optimise=False,
             numHiddenUnits_encoder=0,
-            numHiddentUnits_decoder=10,
+            numHiddenUnits_decoder=10,
+            numHiddenLayers_decoder=2,
             continuous=True
         ):
 
@@ -50,7 +51,7 @@ class VA(SGPDV):
             numberOfEncoderHiddenUnits=numHiddenUnits_encoder
         )
 
-        self.HU_decoder = numHiddentUnits_decoder
+        self.HU_decoder = numHiddenUnits_decoder
         self.continuous = continuous
 
         # Construct appropriately sized matrices to initialise theano shares
@@ -59,15 +60,32 @@ class VA(SGPDV):
         P_HU_mat = np.zeros((self.P, self.HU_decoder), dtype=precision)
         P_vec    = np.zeros((self.P, 1), dtype=precision)
 
-        self.W1 = th.shared(HU_Q_mat, name='W1')
-        self.W2 = th.shared(P_HU_mat, name='W2')
-        self.W3 = th.shared(P_HU_mat, name='W3')
-        self.b1 = th.shared(HU_vec,   name='b1', broadcastable=(False,True))
-        self.b2 = th.shared(P_vec,    name='b2', broadcastable=(False,True))
-        self.b3 = th.shared(P_vec,    name='b3', broadcastable=(False,True))
+        self.W_zh = th.shared(HU_Q_mat, name='W_zh')
+        self.W_hy1 = th.shared(P_HU_mat, name='W_hy')
 
-        self.likelihoodVariables = [self.W1, self.W2, self.W3, self.b1, self.b2, self.b3]
+        self.b_zh = th.shared(HU_vec,   name='b_zh', broadcastable=(False,True))
+        self.b_hy1 = th.shared(P_vec,    name='b_zh', broadcastable=(False,True))
+
+
+        self.likelihoodVariables = [self.W_zh, self.W_hy1, self.b_zh, self.b_hy1]
+
+        if numHiddenLayers_decoder == 2:
[…]

### 2016-02-23 GP_LVM_CMF.py

@@ -89,7 +89,7 @@ class SGPDV(object):
         self.encoderType_qX = encoderType_qX
         self.encoderType_rX = encoderType_rX
         self.encoderType_ru = encoderType_ru
-        self.encoderType_qu = encoderType_qu        
+        self.encoderType_qu = encoderType_qu
         self.Xu_optimise = Xu_optimise
 
         self.y = th.shared(data)
@@ -168,9 +168,9 @@ class SGPDV(object):
             if encoderType_qX == 'FreeForm1':
 
                 self.Phi_full_sqrt = sharedZeroMatrix(self.N, self.N, 'Phi_full_sqrt')
-                
+
                 Phi_batch_sqrt = self.Phi_full_sqrt[self.currentBatch][:, self.currentBatch]
-                
+
                 self.Phi = Tdot(Phi_batch_sqrt, Phi_batch_sqrt.T, 'Phi')
 
                 Phi_batch_sqrt.name = 'Phi_batch_sqrt'
@@ -204,7 +204,7 @@ class SGPDV(object):
             # [HxB] = softplus( [HxP] . [BxP]^T + repmat([Hx1],[1,B]) )
             h_qX = T.nnet.softplus(Tdot(self.W1_qX, self.y_miniBatch.T) + self.b1_qX)
             # [RxB] = sigmoid( [RxH] . [HxB] + repmat([Rx1],[1,B]) )
-            mu_qX = T.nnet.sigmoid(Tdot(self.W2_qX, h_qX) + self.b2_qX)
+            mu_qX = Tdot(self.W2_qX, h_qX) + self.b2_qX
             # [1xB] = 0.5 * ( [1xH] . [HxB] + repmat([1x1],[1,B]) )
             log_sigma_qX = 0.5 * (Tdot(self.W3_qX, h_qX) + self.b3_qX)
 
@@ -265,7 +265,7 @@ class SGPDV(object):
             # [HxQ] = softplus( [HxP] . [PxQ] + repmat([Hx1],[1,Q]) )
             h2_qu = T.nnet.softplus(Tdot(h1_qu, self.W2_qu) + self.b2_qu)
             # [MxQ] = sigmoid( [MxH] . [HxQ] + repmat([Mx1],[1,Q]) )
-            mu_qu = T.nnet.sigmoid(Tdot(self.W3_qu, h2_qu) + self.b3_qu)
+            mu_qu = Tdot(self.W3_qu, h2_qu) + self.b3_qu
[…]

### 2016-02-23 trainmnist_cond_MF.py

@@ -15,6 +15,7 @@ parser = argparse.ArgumentParser()
 
 args = parser.parse_args()
 
+
 print "Loading MNIST data"
 #Retrieved from: http://deeplearning.net/data/mnist/mnist.pkl.gz
 
@@ -22,20 +23,24 @@ f = gzip.open('mnist.pkl.gz', 'rb')
 (x_train, t_train), (x_valid, t_valid), (x_test, t_test)  = cPickle.load(f)
 f.close()
 
-print 'Loading PCA initialisation of Xf if it exists'
+[N,dimX] = x_train.shape
+
+print 'Loading PCA initialisation of Xf with dimX components if it exists'
+filename = 'PCA_init_Xf_{}_components.pkl'.format(dimX)
 try:
-    with open('PCA_init_Xf', 'rb') as f:
-        Xf = pkl.load(f)
+    with open(filename, 'rb') as f:
+        Xf = cPickle.load(f)
 except:
     'Does not exist so performing the PCA now and saving result'
-    pcaXf = PCA()
+    pcaXf = PCA(n_components=dimX)
     Xf = pcaXf.fit_transform(x_train)
-    with open('PCA_init_Xf', 'wb') as ff:
+    filename = 'PCA_init_Xf_{}_components'.format(dimX)
+    with open(filename, 'wb') as ff:
         cPickle.dump(Xf, ff)
 
 data = x_train
 
-[N,dimX] = data.shape
+
 
 dimZ = 40
 dimX = 3

### 2016-02-24 Auto_encoder_model.py
Some fixes

@@ -127,7 +127,7 @@ class VA(SGPDV):
 
         if self.continuous:
             kappa_outer = dot(self.kappa, self.kappa.T, 'kappa_outer')
-            AtA = dot(self.A.T, self.A, 'A''.A' )
+            AtA = dot(self.A.T, self.A)
             KL = 0.5*self.Q*trace(self.C) + 0.5*trace(dot(AtA,kappa_outer)) \
                 + 0.5*self.Q+trace(dot(AtA,self.Kappa)) - 0.5*self.Q*self.B - 0.5*self.Q*self.logDetC
             KL.name = 'KL_qp'
@@ -229,7 +229,7 @@ if __name__ == "__main__":
 
     print 'H_qX'
     print th.function([], va.H_qX())()
- 
+
     va.construct_L_dL_functions()
 
     for i in range(len(va.gradientVariables)):

### 2016-02-24 GP_LVM_CMF.py

@@ -140,11 +140,11 @@ class SGPDV(object):
         self.xi.name   = 'xi'
         self.alpha.name = 'alpha'
         self.beta.name  = 'beta'
-        
+
         self.sample_xi    = th.function([], self.xi)
         self.sample_alpha = th.function([], self.alpha)
         self.sample_beta  = th.function([], self.beta)
-               
+
         self.sample_batchStream = th.function([], self.batchStream)
         self.sample_padStream   = th.function([], self.padStream)
 
@@ -164,7 +164,7 @@ class SGPDV(object):
 
                 Phi_batch_sqrt = self.Phi_full_sqrt[self.currentBatch][:, self.currentBatch]
                 Phi_batch_sqrt.name = 'Phi_batch_sqrt'
-                
+
                 self.Phi = dot(Phi_batch_sqrt, Phi_batch_sqrt.T, 'Phi')
 
                 self.cPhi, _, self.logDetPhi = cholInvLogDet(self.Phi, self.B, 0)
@@ -252,7 +252,7 @@ class SGPDV(object):
         # [BxQ] = [BxM] * [MxQ]
         self.mu = dot(self.A, self.u, 'mu')
         # Sample f from q(f|u,X) = N( mu_q, C )
-        # [BxQ] =  
+        # [BxQ] =
         self.z  = plus(self.mu, (dot(self.cC, self.beta)), 'z')
 
         self.qz_vars = [self.log_theta]
@@ -270,7 +270,7 @@ class SGPDV(object):
             self.b3_rX = sharedZeroVector(self.R, 'b3_rX', broadcastable=(False, True))
 
             # [HxB] = softplus( [Hx(Q+P)] . [(Q+P)xB] + repmat([Hx1], [1,B]) )
-            h_rX = plus(softplus(dot(self.W1_rX, T.concatenate((self.z.T, self.y_miniBatch.T)))), self.b1_rX, 'h_rX')
+            h_rX = softplus(plus(dot(self.W1_rX, T.concatenate((self.z.T, self.y_miniBatch.T))), self.b1_rX), 'h_rX')
             # [RxB] = softplus( [RxH] . [HxB] + repmat([Rx1], [1,B]) )
             mu_rX = plus(dot(self.W2_rX, h_rX), self.b2_rX, 'mu_rX')
             # [RxB] = 0.5*( [RxH] . [HxB] + repmat([Rx1], [1,B]) )
@@ -294,11 +294,11 @@ class SGPDV(object):
[…]

### 2016-02-24 GP_LVM_CMF.py
name change for Xf to Xz

@@ -561,21 +561,21 @@ class SGPDV(object):
 
         return self.lowerBounds
 
-    def init_Xu_from_Xf(self):
+    def init_Xu_from_Xz(self):
 
-        Xf_min = np.zeros(self.R,)
-        Xf_max = np.zeros(self.R,)
-        Xf_locations = th.function([], self.phi, no_default_updates=True) # [B x R]
+        Xz_min = np.zeros(self.R,)
+        Xz_max = np.zeros(self.R,)
+        Xz_locations = th.function([], self.phi, no_default_updates=True) # [B x R]
         for b in range(self.numberofBatchesPerEpoch):
             self.iterator.set_value(b)
-            Xf_batch = Xf_locations()
-            Xf_min = np.min( (Xf_min, Xf_batch.min(axis=0)), axis=0)
-            Xf_max = np.max( (Xf_min, Xf_batch.max(axis=0)), axis=0)
+            Xz_batch = Xz_locations()
+            Xz_min = np.min( (Xz_min, Xz_batch.min(axis=0)), axis=0)
+            Xz_max = np.max( (Xz_min, Xz_batch.max(axis=0)), axis=0)
 
-        Xf_min.reshape(-1,1)
-        Xf_max.reshape(-1,1)
-        Df = Xf_max - Xf_min
-        Xu = np.random.rand(self.M, self.R) * Df + Xf_min # [M x R]
+        Xz_min.reshape(-1,1)
+        Xz_max.reshape(-1,1)
+        Df = Xz_max - Xz_min
+        Xu = np.random.rand(self.M, self.R) * Df + Xz_min # [M x R]
 
         self.Xu.set_value(Xu, borrow=True)
 

### 2016-02-24 Auto_encoder_model.py
Added funtionality to plot generate new data from a learned decoder, and plot in 2d space.

@@ -92,6 +92,14 @@ class VA(SGPDV):
                 if var.name.startswith('W_'):
                     var.set_value(var.get_value()/4.0)
 
+    def generate_new_data(self, z_test):
+        h_decoder  = softplus(dot(self.W_zh,z_test.T) + self.b_zh)
+        if self.numHiddenLayers_decoder == 2:
+            h_decoder = softplus(dot(self.W_hh, h_decoder) + self.b_hh)
+        mu_decoder = dot(self.W_hy1, h_decoder) + self.b_hy1
+
+        return mu_decoder
+
     def log_p_y_z(self):
 
         if self.continuous:

### 2016-02-24 trainmnist_cond_MF.py

@@ -5,6 +5,7 @@ import argparse
 import gzip, cPickle
 import os
 import theano.tensor as T
+from matplotlib import pyplot as plt
 # from sklearn.decomposition import PCA
 
 os.environ["CUDA_LAUNCH_BLOCKING"] = "1"
@@ -41,7 +42,7 @@ f.close()
 
 data = x_train
 
-dimZ = 40
+dimZ = 2
 dimX = 3
 batchSize = 200
 encoderType_qX='MLP'
@@ -80,8 +81,8 @@ va.setKernelParameters(5*np.ones((2,)),
     [1e-10,1e-10], [10,10] )
 
 va.randomise()
-
-va.init_Xu_from_Xf()
+# va.Xz.set_value()
+va.init_Xu_from_Xz()
 
 
 #va.printMemberTypes()
@@ -98,6 +99,25 @@ numberOfEpochs = 10
 
 
 va.train(numberOfEpochs=numberOfEpochs)
+
+# Generate synthesised data from the learned decoder
+# at each point in a grid of a 2d latent space
+
+nx = ny = 20
+x_values = np.linspace(-3, 3, nx)
+y_values = np.linspace(-3, 3, ny)
+
+canvas = np.empty((28*ny, 28*nx))
+for i, yi in enumerate(x_values):
+    for j, xi in enumerate(y_values):
+        z_mu = np.array([[xi, yi]])
+        x_mean = va.generate_new_data(z_mu)
+        canvas[(nx-i-1)*28:(nx-i)*28, j*28:(j+1)*28] = x_mean[0].reshape(28, 28)
+
+plt.figure(figsize=(8, 10))
+Xi, Yi = np.meshgrid(x_values, y_values)
+plt.imshow(canvas, origin="upper")
+plt.tight_layout()
 
 #for i in range(1,8):
 #

### 2016-02-24 Auto_encoder_model.py
New visualisation functionality made to work properly with Theano.

@@ -92,11 +92,13 @@ class VA(SGPDV):
                 if var.name.startswith('W_'):
                     var.set_value(var.get_value()/4.0)
 
-    def generate_new_data(self, z_test):
+    def create_new_data_function(self):
+        self.z_test = sharedZeroMatrix(self.Q,1,'z_test')
         h_decoder  = softplus(dot(self.W_zh,z_test.T) + self.b_zh)
         if self.numHiddenLayers_decoder == 2:
             h_decoder = softplus(dot(self.W_hh, h_decoder) + self.b_hh)
         mu_decoder = dot(self.W_hy1, h_decoder) + self.b_hy1
+        self.new_data_function = th.function([], mu_decoder, no_default_updates=True)
 
         return mu_decoder
 

### 2016-02-24 trainmnist_cond_MF.py

@@ -108,10 +108,12 @@ x_values = np.linspace(-3, 3, nx)
 y_values = np.linspace(-3, 3, ny)
 
 canvas = np.empty((28*ny, 28*nx))
+va.create_new_data_function()
 for i, yi in enumerate(x_values):
     for j, xi in enumerate(y_values):
         z_mu = np.array([[xi, yi]])
-        x_mean = va.generate_new_data(z_mu)
+        va.z_test.set_value(z_mu)
+        x_mean = va.new_data_function()
         canvas[(nx-i-1)*28:(nx-i)*28, j*28:(j+1)*28] = x_mean[0].reshape(28, 28)
 
 plt.figure(figsize=(8, 10))

### 2016-02-24 Auto_encoder_model.py
Fixed bug

@@ -94,7 +94,7 @@ class VA(SGPDV):
 
     def create_new_data_function(self):
         self.z_test = sharedZeroMatrix(self.Q,1,'z_test')
-        h_decoder  = softplus(dot(self.W_zh,z_test.T) + self.b_zh)
+        h_decoder  = softplus(dot(self.W_zh,self.z_test.T) + self.b_zh)
         if self.numHiddenLayers_decoder == 2:
             h_decoder = softplus(dot(self.W_hh, h_decoder) + self.b_hh)
         mu_decoder = dot(self.W_hy1, h_decoder) + self.b_hy1

### 2016-02-24 trainmnist_cond_MF.py
minor edit

@@ -94,7 +94,7 @@ va.constructUpdateFunction()
 
 print "Training"
 learning_rate = 1e-3
-numberOfEpochs = 10
+numberOfEpochs = 3
 
 
 

### 2016-02-24 trainmnist_cond_MF.py
minor

@@ -42,7 +42,7 @@ f.close()
 
 data = x_train
 
-dimZ = 2
+dimZ = 40
 dimX = 3
 batchSize = 200
 encoderType_qX='MLP'
@@ -94,32 +94,34 @@ va.constructUpdateFunction()
 
 print "Training"
 learning_rate = 1e-3
-numberOfEpochs = 10
+numberOfEpochs = 100
 
 
 
 va.train(numberOfEpochs=numberOfEpochs)
 
-# Generate synthesised data from the learned decoder
-# at each point in a grid of a 2d latent space
+makePlots=False
+if makePlots:
+    # Generate synthesised data from the learned decoder
+    # at each point in a grid of a 2d latent space
 
-nx = ny = 20
-x_values = np.linspace(-3, 3, nx)
-y_values = np.linspace(-3, 3, ny)
+    nx = ny = 20
+    x_values = np.linspace(-3, 3, nx)
+    y_values = np.linspace(-3, 3, ny)
 
-canvas = np.empty((28*ny, 28*nx))
-va.create_new_data_function()
-for i, yi in enumerate(x_values):
-    for j, xi in enumerate(y_values):
-        z_mu = np.array([[xi, yi]])
-        va.z_test.set_value(z_mu)
-        x_mean = va.new_data_function()
-        canvas[(nx-i-1)*28:(nx-i)*28, j*28:(j+1)*28] = x_mean[0].reshape(28, 28)
+    canvas = np.empty((28*ny, 28*nx))
+    va.create_new_data_function()
+    for i, yi in enumerate(x_values):
+        for j, xi in enumerate(y_values):
+            z_mu = np.array([[xi, yi]])
+            va.z_test.set_value(z_mu)
+            x_mean = va.new_data_function()
+            canvas[(nx-i-1)*28:(nx-i)*28, j*28:(j+1)*28] = x_mean[0].reshape(28, 28)
 
-plt.figure(figsize=(8, 10))
-Xi, Yi = np.meshgrid(x_values, y_values)
-plt.imshow(canvas, origin="upper")
-plt.tight_layout()
+    plt.figure(figsize=(8, 10))
+    Xi, Yi = np.meshgrid(x_values, y_values)
+    plt.imshow(canvas, origin="upper")
+    plt.tight_layout()
 
 #for i in range(1,8):
 #

### 2016-02-24 Auto_encoder_model.py
debug information output, condition number, ARD kernel added

@@ -24,7 +24,7 @@ class VA(SGPDV):
             dimX,                   # Dimensionality of the latent co-ordinates
             dimZ,                   # Dimensionality of the latent variables
             data,                   # [NxP] matrix of observations
-            kernelType='RBF',
+            kernelType='ARD',
             encoderType_qX='FreeForm2',  # MLP', 'Kernel'.
             encoderType_rX='FreeForm2',  # MLP', 'Kernel'.
             Xu_optimise=False,

### 2016-02-24 GP_LVM_CMF.py

@@ -8,6 +8,7 @@ from theano.tensor.shared_randomstreams import RandomStreams
 # import progressbar
 import time
 import collections
+from myCond import myCond
 
 from optimisers import Adam
 from utils import cholInvLogDet, sharedZeroArray, sharedZeroMatrix, sharedZeroVector, \
@@ -30,21 +31,32 @@ class kernelFactory(object):
         else:
             _X2 = X2
         if self.kernelType == 'RBF':
-            inls = T.exp(theta[0])
+            inls = T.exp(theta[0,1])
             # dist = (((X1 / theta[0])**2).sum(1)) + (((_X2 / theta[0])**2).sum(1)).T - 2*dot( X1 / theta[0], _X2.T / theta[0] )
             dist = ((X1 / inls)**2).sum(1)[:, None] + ((_X2 / inls)**2).sum(1)[None, :] - 2 * (X1 / inls).dot((_X2 / inls).T)
             # Always want kernels to be 64-bit
             if dtype == 'float64':
                 dist = T.cast(dist, dtype='float64')
-            K = T.exp(theta[1] - dist / 2.0)
+            K = T.exp(theta[0,0] - dist / 2.0)
+            if X2 is None:
+                K = K + self.eps * T.eye(X1.shape[0])
+            K.name = name_ + '(RBF)'
+        elif self.kernelType == 'ARD':
+            inls = T.exp(theta[0,1:])
+            # dist = (((X1 / theta[0])**2).sum(1)) + (((_X2 / theta[0])**2).sum(1)).T - 2*dot( X1 / theta[0], _X2.T / theta[0] )
+            dist = ((X1 / inls)**2).sum(1)[:, None] + ((_X2 / inls)**2).sum(1)[None, :] - 2 * (X1 / inls).dot((_X2 / inls).T)
+            # Always want kernels to be 64-bit
+            if dtype == 'float64':
+                dist = T.cast(dist, dtype='float64')
+            K = T.exp(theta[0,0] - dist / 2.0)
             if X2 is None:
                 K = K + self.eps * T.eye(X1.shape[0])
             K.name = name_ + '(RBF)'
         elif self.kernelType == 'RBFnn':
-            K = theta[0] + self.eps
+            K = theta[0,0] + self.eps
             K.name = name_ + '(RBFnn)'
         elif self.kernelType == 'LIN':
-            K = theta[0] * (X1.dot(_X2.T) + 1)
+            K = theta[0,0] * (X1.dot(_X2.T) + 1)
[…]

### 2016-02-24 myCond.py

@@ -1,0 +1,44 @@
+
+import theano
+from theano.gof import Op, Apply
+from theano.tensor import as_tensor_variable
+import numpy
+
+class myCond(Op):
+    """
+    Matrix condition number. Input should be a square matrix.
+    """
+
+    __props__ = ()
+
+    def make_node(self, x):
+        x = as_tensor_variable(x)
+        assert x.ndim == 2
+        o = theano.tensor.scalar(dtype=x.dtype)
+        return Apply(self, [x], [o])
+
+    def perform(self, node, inputs, outputs):
+        (x,) = inputs
+        (z,) = outputs
+        try:
+            z[0] = numpy.asarray(numpy.linalg.cond(x), dtype=x.dtype)
+        except Exception:
+            print('Failed to compute condition number', x)
+            raise
+
+    def infer_shape(self, node, shapes):
+        return [()]
+
+    def __str__(self):
+        return ""
+
+
+if __name__ == '__main__':
+
+	A = theano.shared(numpy.asarray(numpy.random.randn(5, 5)), name='A')
+	B = theano.tensor.dot(A, A.T)
+	c = myCond()(B)
+	cf = theano.function([], c, no_default_updates=True)
+	print 'conition number of c = {}'.format( cf() )
+
+

### 2016-02-24 trainmnist_cond_MF.py

@@ -6,7 +6,7 @@ import gzip, cPickle
 import os
 import theano.tensor as T
 from matplotlib import pyplot as plt
-# from sklearn.decomposition import PCA
+from sklearn.decomposition import PCA
 
 os.environ["CUDA_LAUNCH_BLOCKING"] = "1"
 os.environ["LD_LIBRARY_PATH"]   = os.path.dirname(os.path.realpath(__file__)) + '/fastlin/'
@@ -26,29 +26,28 @@ f.close()
 
 [N,dimX] = x_train.shape
 
-# print 'Loading PCA initialisation of Xf with dimX components if it exists'
-# filename = 'PCA_init_Xf_{}_components.pkl'.format(dimX)
-# try:
-#     with open(filename, 'rb') as f:
-#         Xf = cPickle.load(f)
-# except:
-#     'Does not exist so performing the PCA now and saving result'
-#     pcaXf = PCA(n_components=dimX)
-#     Xf = pcaXf.fit_transform(x_train)
-#     filename = 'PCA_init_Xf_{}_components'.format(dimX)
-#     with open(filename, 'wb') as ff:
-#         cPickle.dump(Xf, ff)
-
+print 'Loading PCA initialisation of Xf with dimX components if it exists'
+filename = 'PCA_init_Xf_{}_components.pkl'.format(dimX)
+try:
+    with open(filename, 'rb') as f:
+        Xf = cPickle.load(f)
+except:
+    'Does not exist so performing the PCA now and saving result'
+    pcaXf = PCA(n_components=dimX)
+    Xf = pcaXf.fit_transform(x_train)
+    filename = 'PCA_init_Xf_{}_components'.format(dimX)
+    with open(filename, 'wb') as ff:
+        cPickle.dump(Xf, ff)
 
 data = x_train
 
 dimZ = 40
-dimX = 3
-batchSize = 200
-encoderType_qX='MLP'
+dimX = 30
+batchSize = 400
+encoderType_qX='FreeForm2'
 encoderType_rX='MLP'
 Xu_optimise=True
-kernelType='RBF'
+kernelType='ARD'
 numHiddenUnits_encoder=400
 numHiddentUnits_decoder=400
 numberOfInducingPoints =500
@@ -74,32 +73,25 @@ va = VA(
     continuous=True
 )
 
-va.construct_L_using_r()
-
[…]

### 2016-02-24 utils.py

@@ -24,9 +24,9 @@ def invLogDet( C ):
     # Return inv(A) and log det A where A = C . C^T 
     iC = nlinalg.matrix_inverse(C)
     iC.name = 'i' + C.name
-    iA  = T.dot( iC.T, iC )
-    iA.name    = 'i' + C.name[1:]        
-    logDetA = 2.0*T.sum( T.log( T.abs_( T.diag(C) ) ) )
+    iA = T.dot(iC.T, iC)
+    iA.name = 'i' + C.name[1:]        
+    logDetA = 2.0*T.sum(T.log(T.abs_(T.diag(C))))
     logDetA.name = 'logDet' + C.name[1:]    
     return(iA, logDetA)
     
@@ -49,7 +49,8 @@ def cholInvLogDet(A, dim, jitter, fast=False):
         (iA,logDetA) = invLogDet(cA)
     else:
         iA = nlinalg.matrix_inverse(A_jitter)
-        logDetA = T.log( nlinalg.Det()(A_jitter) )
+        #logDetA = T.log( nlinalg.Det()(A_jitter) )
+        logDetA = 2.0*T.sum(T.log(T.abs_(T.diag(cA))))
         iA.name = 'i' + A.name
         logDetA.name = 'logDet' + A.name
 
@@ -91,14 +92,14 @@ def np_log_mean_exp_stable(x, axis=0):
     m = np.max(x, axis=axis, keepdims=True)
     return m + np.log(np.mean(np.exp(x - m), axis=axis, keepdims=True))
 
-def sharedZeroMatrix(M, N, name, dtype=th.config.floatX):
-    return th.shared(np.asarray(np.zeros((M, N)), dtype), name=name)
+def sharedZeroMatrix(M, N, name, dtype=th.config.floatX, broadcastable=[]):
+    if len(broadcastable) == 0:
+        return th.shared(np.asarray(np.zeros((M, N)), dtype), name=name)
+    else:
+        return th.shared(np.asarray(np.zeros((M, N)), dtype), name=name, broadcastable=broadcastable)
 
 def sharedZeroVector(M, name, dtype=th.config.floatX, broadcastable=[]):
-    if len(broadcastable) == 0:
-        return th.shared(np.asarray(np.zeros((M, 1)), dtype), name=name)
-    else:
-        return th.shared(np.asarray(np.zeros((M, 1)), dtype), name=name, broadcastable=broadcastable)
[…]

### 2016-02-25 trainmnist_cond_MF.py
Created new branch to capture code base that seems to give good bounds often

@@ -42,8 +42,8 @@ f.close()
 
 data = x_train
 
-dimZ = 40
-dimX = 3
+dimZ = 50
+dimX = 100
 batchSize = 200
 encoderType_qX='MLP'
 encoderType_rX='MLP'
@@ -94,7 +94,7 @@ va.constructUpdateFunction()
 
 print "Training"
 learning_rate = 1e-3
-numberOfEpochs = 100
+numberOfEpochs = 7
 
 
 

### 2016-02-28 Auto_encoder_model.py
more changes

@@ -61,8 +61,8 @@ class AutoEncoderModel(Printable):
         self.lowerBound = -np.inf  # Lower bound
         self.lowerBounds = []
 
+        self.jitterProtector = jitterProtector()
         if encoderType == 'Hybrid':
-            self.jitterProtector = jitterProtector()
             self.encoder = Hybrid_variational_model(
                 self.y_miniBatch,
                 self.B,
@@ -88,6 +88,7 @@ class AutoEncoderModel(Printable):
 
         self.encoder.construct_L_terms()
         self.decoder.construct_L_terms()
+
         self.L = self.encoder.L_terms + self.decoder.L_terms
 
         self.dL = T.grad(self.L, self.gradientVariables)
@@ -125,7 +126,7 @@ class AutoEncoderModel(Printable):
                 # Sample from the encoder
                 self.encoder.sample()
                 self.iterator.set_value(it)
-                lbTmp = self.jitterProtect(self.updateFunction, reset=False)
+                lbTmp = self.jitterProtector.jitterProtect(self.updateFunction, reset=False)
                 if constrain:
                     self.constrainKernelParameters()
 

### 2016-02-28 GP_LVM_CMF.py

@@ -93,9 +93,6 @@ class SGPDV(printable):
         self.R = params['dimX']
         self.M = params['numberOfInducingPoints']
         self.H = params['numberOfEncoderHiddenUnits']
-        self.encoderType_qX = params['encoderType_qX']
-        self.encoderType_rX = params['encoderType_rX']
-        self.Xu_optimise = params['Xu_optimise']
         kernelType = params['kernelType']
 
         if kernelType == 'RBF':
@@ -112,85 +109,27 @@ class SGPDV(printable):
         # kernel parameters
         self.log_theta = sharedZeroMatrix(
             1, self.numberOfKernelParameters, 'log_theta', broadcastable=(True, False))  # parameters of Kuu, Kuf, Kff
-        self.log_omega = sharedZeroMatrix(
-            1, self.numberOfKernelParameters, 'log_omega', broadcastable=(True, False))  # parameters of Kuu, Kuf, Kff
-        self.log_gamma = sharedZeroMatrix(
-            1, self.numberOfKernelParameters, 'log_gamma', broadcastable=(True, False))  # parameters of Kuu, Kuf, Kff
-
+        
         # Random variables
-        self.alpha = srng.normal(
-            size=(self.B, self.R), avg=0.0, std=1.0, ndim=None)
-        self.beta = srng.normal(
-            size=(self.B, self.Q), avg=0.0, std=1.0, ndim=None)
-        self.alpha.name = 'alpha'
-        self.beta.name = 'beta'
+        alpha = srng.normal(size=(self.B, self.R), avg=0.0, std=1.0, ndim=None)
+        beta = srng.normal(size=(self.B, self.Q), avg=0.0, std=1.0, ndim=None)
+        alpha.name = 'alpha'
+        beta.name = 'beta'
 
-        self.sample_alpha = th.function([], self.alpha)
-        self.sample_beta = th.function([], self.beta)
+        self.sample_alpha = th.function([], alpha)
+        self.sample_beta = th.function([], beta)
 
-        # Compute parameters of q(X)
-        if self.encoderType_qX == 'FreeForm1' or self.encoderType_qX == 'FreeForm2':
-            # Have a normal variational distribution over location of latent
-            # co-ordinates
-
-            self.phi_full = sharedZeroMatrix(self.N, self.R, 'phi_full')
[…]

### 2016-02-28 Hybrid_variational_model.py

@@ -10,45 +10,53 @@ import theano as th
 import theano.tensor as T
 from theano.tensor import nlinalg
 from printable import Printable
+from utils import plus, mul
 
 from GP_LVM_CMF import SGPDV
-from testTools import checkgrad
-from utils import log_mean_exp_stable, dot, trace, softplus, sharedZeroVector, sharedZeroMatrix, plus
 
-precision = th.config.floatX
-
-class Hyrid_encoder(printable):
+class Hybrid_encoder(Printable):
 
     def __init__(
 
-        def __init__(self, y_miniBatch, minbatchSize, dimY, dimZ, params):
+        def __init__(self, y_miniBatch, minbatchSize, jitterProtect, dimY, dimZ, params):
 
-        self.B = minbatchSize
-        self.P = dimY
-        self.Q = dimZ
+        HU = params['numHiddenUnits_encoder']
+        HL = params['numHiddenLayers_encoder']
 
-        numHiddenUnits_encoder = params['numHiddenUnits_encoder']
-        numHiddenLayers_encoder = params['numHiddenLayers_encoder']
+        self.gp_encoder = SGPDV(y_miniBatch,
+                                miniBatchSize,
+                                dimY,
+                                dimZ,
+                                jitterProtect,
+                                params)
 
-                 dimY,
-                 dimZ,
-                 jitterProtect,
-                 params,
-                 )
+        self.mlp_encoder = MLP_Network(self, dimY, dimZ,
+                HL, 'encoder', num_layers=HL)
 
-        self.mlp_encoder = MLP_Network(self, self.P, self.Q,
-                numHiddenUnits_encoder, 'encoder', num_layers=numHiddenLayers_encoder)
-        self.mu_encoder, self.log_sigma_encoder2 = self.mlp_encoder.setup(T.concatenate((self.gp_encoder.f, y_miniBatch))
+        self.mu_encoder, self.log_sigma_encoder2 \
+            = self.mlp_encoder.setup(T.concatenate((self.gp_encoder.f, y_miniBatch)))
          
-        self.beta = srng.normal(size=(self.Q, self.B), avg=0.0, std=1.0, ndim=None)
+        gamma = srng.normal(size=(dimZ, miniBatchSize), avg=0.0, std=1.0, ndim=None)
+        gamma.name = 'gamma'
+        self.sample_gamma = th.function([], gamma)
 
[…]

### 2016-02-28 nnet.py

@@ -8,6 +8,7 @@ from utils import tanh, sigmoid, softplus, exp
 floatX = theano.config.floatX
 
 class Linear():
+
     def __init__(self, dim_in, dim_out, name):
         self.dim_in  = dim_in
         self.dim_out = dim_out
@@ -34,7 +35,8 @@ class Linear():
             X = var.ge
 
             symInterval = 4.0 * np.sqrt(6. / (X + Y))
-            X_Y_mat = np.asarray(np.random.uniform(size=(X, Y),
+            X_Y_mat = np.asarray(np.random.uniform(size=(X, Y)))
+            
         elif type(nonlinearity) == Linear:
             raise RuntimeError('Consecutive linear layers')
 
@@ -117,7 +119,7 @@ class MLP_Network():
         h_outin = self.hidden.setup(x_in)
         if self.continuous:
             mu = self.muLinear.setup(h_inout, **kwargs)
-            logsigma = self.logsigmaLinear.setup(h_inout, **kwargs)
+            logsigma = 0.5 * self.logsigmaLinear.setup(h_inout, **kwargs)
             mu.name = 'mu_' + name
             logsigma.name = 'logsimga' + name
             return (mu, sigma)

### 2016-02-28 utils.py

@@ -176,3 +176,10 @@ def namedFunction(A, func, funcname, name):
     else:
         B.name = name 
     return B
+
+def conditionNumber(M):
+    cond = myCond()(M)
+    cond.name = 'cond(' + getname(M) + ')'
+    condNum = th.function(
+            [], cond, no_default_updates=True)
+    return condNum

### 2016-02-28 Auto_encoder_model.py
more changes

@@ -87,7 +87,7 @@ class AutoEncoderModel(Printable):
             raise RuntimeErorr('Unrecognised decoder type')
 
         self.encoder.construct_L_terms()
-        self.decoder.construct_L_terms()
+        self.decoder.construct_L_terms(self.encoder)
 
         self.L = self.encoder.L_terms + self.decoder.L_terms
 

### 2016-02-28 GP_LVM_CMF.py

@@ -233,18 +233,6 @@ class SGPDV(printable):
         self.log_theta_min = np.array(np.log(theta_min), dtype=precision)
         self.log_theta_max = np.array(np.log(theta_max), dtype=precision)
 
-        if self.encoderType_qX == 'Kernel':
-            self.log_gamma.set_value(
-                np.asarray(np.log(gamma), dtype=precision))
-            self.log_gamma_min = np.array(np.log(gamma_min), dtype=precision)
-            self.log_gamma_max = np.array(np.log(gamma_max), dtype=precision)
-
-        if self.encoderType_rX == 'Kernel':
-            self.log_omega.set_value(
-                np.asarray(np.log(omega), dtype=precision))
-            self.log_omega_min = np.array(np.log(omega_min), dtype=precision)
-            self.log_omega_max = np.array(np.log(omega_max), dtype=precision)
-
     def constrainKernelParameters(self):
 
         def constrain(variable, min_val, max_val):
@@ -266,11 +254,6 @@ class SGPDV(printable):
                     variable.set_value(vals)
 
         constrain(self.log_theta, self.log_theta_min, self.log_theta_max)
-
-        if self.encoderType_qX == 'Kernel':
-            constrain(self.log_gamma, self.log_gamma_min, self.log_gamma_max)
-        if self.encoderType_rX == 'Kernel':
-            constrain(self.log_omega, self.log_omega_min, self.log_omega_max)
 
     def init_Xu_from_Xf(self):
 

### 2016-02-28 Hybrid_variational_model.py

@@ -40,7 +40,7 @@ class Hybrid_encoder(Printable):
         gamma.name = 'gamma'
         self.sample_gamma = th.function([], gamma)
 
-        self.z = plus(self.mu_qz, mul(T.exp(self.log_sigma_qz * 0.5), gamma), 'z')
+        self.z = plus(self.mu_qz, mul(T.exp(self.log_sigma_qz), gamma), 'z')
 
         self.construct_rfXf(self.z)
 

### 2016-02-28 Hysterisis_variational_model.py

@@ -1,0 +1,66 @@
+# -*- coding: utf-8 -*-
+"""
+Created on Thu Feb  4 20:44:40 2016
+
+@author: clloyd
+"""
+
+import numpy as np
+import theano as th
+import theano.tensor as T
+
+from printable import Printable
+from utils import plus, mul
+
+precision = th.config.floatX
+
+log2pi = T.constant(np.log(2 * np.pi))
+
+class Hysterisis_encoder(Printable):
+
+    def __init__(self, y_miniBatch, minbatchSize, dimY, dimZ, params):
+
+        self.B = minbatchSize
+
+        numHiddenUnits_encoder = params['numHiddenUnits_encoder']
+        numHiddenLayers_encoder = params['numHiddenLayers_encoder']
+
+        self.mlp_f_y = MLP_Network(self, dimY, dimZ,
+                numHiddenUnits_encoder, 'encoder', num_layers=numHiddenLayers_encoder)
+        self.mu_f_y, self.log_sigma_f_y = self.mlp_f_y.setup(y_miniBatch)
+ 
+        alpha = srng.normal(size=(dimZ, minbatchSize), avg=0.0, std=1.0, ndim=None)
+        alpha.name = 'alpha'
+        self.sample_alpha = th.function([], alpha)
+
+        self.f = plus(self.mu_f_y, mul(T.exp(self.log_sigma_f_y*0.5), alpha), 'f')
+
+        self.mlp_z_fy = MLP_Network(self, dimY, dimZ,
+                numHiddenUnits_encoder, 'encoder', num_layers=numHiddenLayers_encoder)
+        self.mu_qz_fy, self.log_sigma_qz_fy = self.mlp_z_fy.setup(T.concatenate((self.f, y_miniBatch)))
+         
+        beta = srng.normal(size=(dimZ, minbatchSize), avg=0.0, std=1.0, ndim=None)
+        beta.name = 'beta'
+        self.sample_beta = th.function([], beta)
+
+        self.z = plus(self.mu_qz_fy, mul(T.exp(self.log_sigma_qz_fy*0.5), beta), 'z')
+
+        self.mlp_rf_yz = MLP_Network(self, (dimY+dimZ), dimZ,
+                numHiddenUnits_encoder, 'encoder', num_layers=numHiddenLayers_encoder)
+        self.mu_rf_yz, self.log_sigma_rf_yz = self.mlp_rf_yz.setup(T.concatenate((y_miniBatch,self.z)))
+
[…]

### 2016-02-28 MLP_likelihood_model.py

@@ -1,0 +1,53 @@
+# -*- coding: utf-8 -*-
+"""
+Created on Thu Feb  4 20:44:40 2016
+
+@author: clloyd
+"""
+
+import numpy as np
+import theano as th
+import theano.tensor as T
+from printable import printable
+
+log2pi = np.log(2 * np.pi)
+
+class MLP_likelihood_model(Printable):
+
+    def __init__(self, y_miniBatch, miniBatchSize, dimY, dimZ, encoder, params):
+
+        self.B = miniBatchSize
+
+        numHiddenUnits_decoder = params['numHiddenUnits_decoder']
+        numHiddenLayers_decoder = params['numHiddenLayers_decoder']
+        self.continuous = params['continuous']
+
+        if self.continuous:
+            self.mlp_decoder = MLP_Network(self, dimZ, dimY,
+                numHiddenUnits_decoder, 'decoder', num_layers=numHiddenLayers_decoder)
+            self.mu_decoder, self.log_sigma_decoder = self.mlp_decoder.setup(encoder.z)
+        else:
+            self.mlp_decoder = MLP_Network(self, dimZ, dimY,
+                numHiddenUnits_decoder, 'decoder', num_layers=numHiddenLayers_decoder, continuous=False)
+            self.yhat = self.mlp_decoder.setup(encoder.z)
+
+        self.gradientVariables = mlp.decoder.params
+
+    def construct_L_terms(self, encoder):
+
+        self.KL_qp = 0.5*(T.sum(T.exp(encoder.log_simga_qz_fy*2))) \
+                   + 0.5 * T.sum(encoder.mu_qz_fy**2) \
+                   - T.sum(encoder.log_simga_qz_fy)
+                   - self.B
+
+        if self.continuous:
+            self.log_p_y_z = 
+        else:
+            pass
+
+
+        self.L_terms = self.log_pyz + self.KL
+
+    def randomise(self):
+        self.mlp_decoder.randomise()
+

### 2016-02-28 MLP_variational_model.py

@@ -1,0 +1,55 @@
+# -*- coding: utf-8 -*-
+"""
+Created on Thu Feb  4 20:44:40 2016
+
+@author: clloyd
+"""
+
+import numpy as np
+import theano as th
+import theano.tensor as T
+from theano.tensor import nlinalg
+
+from testTools import checkgrad
+from utils import log_mean_exp_stable
+
+precision = th.config.floatX
+
+class MLP_variational_model(Printable):
+
+    def __init__(self, y_miniBatch, minBatchSize, dimY, dimZ, params):
+
+    	self.B = minBatchSize
+
+    	numHiddenUnits_encoder = params['numHiddenUnits_encoder']
+        numHiddenLayers_encoder = params['numHiddenLayers_encoder']
+ 
+        self.mlp_encoder = MLP_Network(self, dimY, dimZ,
+                numHiddenUnits_encoder, 'encoder', num_layers=numHiddenLayers_encoder)
+
+        self.mu_encoder, self.log_sigma_encoder = self.mlp_encoder.setup(y_miniBatch)
+         
+        alpha = srng.normal(size=(dimZ, self.B), avg=0.0, std=1.0, ndim=None)
+        alpha.name = 'alpha'
+        self.sample_alpha = th.function([], alpha)
+
+        self.gradientVariables = mlp.encoder.params
+
+        self.z = plus(self.mu_encoder, mul(T.exp(self.log_sigma_encoder*0.5), alpha))
+
+    def construct_L_terms():
+    	pass
+
+    def sample(self):
+        self.sample_alpha()
+
+        
+
+
+
+
+
+
+
+ 
+

### 2016-02-28 icml2016/paper.tex
Fixed a few bugs and implemented log_pyz for continuous and discrete.

@@
 % Just a working title/suggestion
-\title{Blackbox variational inference using \\ the Gaussian process latent variable model}
+\title{Black-box Auxiliary Variational Inference}
 
@@
 \textnormal{University of Oxford} \\
-\url{{rbeard, clloyd, sjrob}@robots.ox.ac.uk}
+\url{{rory, clloyd, sjrob}@robots.ox.ac.uk}
 }
@@
 \begin{abstract}
-The purpose of this document is to provide both the basic paper template and
-submission guidelines. Abstracts should be a single paragraph, between 4--6 sentences long, ideally.
-Gross violations will trigger corrections at the camera-ready phase.
+We approach the problem of capturing the statistical dependence structure of the conditionally independent variational factors using a flexible, auxiliary variable method.
+We introduce two flexible auxiliary inference models; a deep neural net and a Bayesian non-paramteric model.
+We compare the performance of these models against each other and also consider a hybrid approach.
+The power of the approach lies in its ability to combine bottom-up and top-down learning in a fully Bayesian manner via a bidirectional augmented inference network.
 \end{abstract}
@@
 
-We approach the problem of capturing the statistical dependence structure of the conditionally independent variational factors in a flexible, Bayesian nonparametric manner.
 Considering at first a latent variable model with a latent variable for each datum as well as some global parameters, a traditional (mean field) variational approach is to learn a factored approximation to the posterior in which each latent variable is independently distributed and governed by its own variational parameter.
@@
 The dependence is induced by a fixed number of latent coordinates, the locations of which are learned so as to induce the optimal dependence structure.
-We refer to this scheme as the GP-LVM-conditional mean field
+We refer to this scheme as GP-LVM-conditional mean field (GP-LVM-CMF) inference.
+
+In addition to the GP-LVM, we introduce an auxiliary neural net approach which naturally extends the variational auto-encoder to a hierarchical version which is able to the bottom-up stochastic message passing of the variational encoder with top-down stochastic messages from a variational auxiliary decoder.
 
@@
 
-l;asdf. We additionally introduce a kernel based approach to back-constrained amortization in which random functions are trained to capture the encoding process.
+We additionally introduce a kernel based approach to back-constrained amortization in which random functions are trained to capture the encoding process.
 WHY IS THIS GOOD: provable guarantees of kernel mappings, ameliorates pathological curvature??
@@
 \end{itemize}
+
+\subsubsection{Variational Auto-encoding}
+
+The VAE (2 REFERENCES) was introduced as a framework to train (potentially) deep neural net generative models in a scalable, Bayesian manner.
+The important insight of these authors was to reparameterise the variational variables as a deterministic transformation of a random variable drawn from a standard (parameterless) distribution so that gradient information relating to the variational parameters can be included in the stochastic gradients used to optimise the bound.
+Importantly this allows for variance control of the stochastic gradients, permitting practical convergence profiles, unlike the previously proposed sleep wake algorithm (REFERENCES).
+\\ |
+\\ |
+\\ |
+IMPORTANT VAE EQUATIONS
+\\ |
+\\ |
+\\ |
 
@@
 However the motivations for this conditional inducing structure are distinct in these three cases.
-For the sparse GP the desire is scalability; in particular to circumvent the costly $\mathcal{O}(N^3$ complexity of the covariance matrix inversion, which is then limited to $\mathcal{O}(N^2M)$ through variational compression.
+For the sparse GP the desire is scalability; in particular to circumvent the costly $\mathcal{O}(N^3)$ complexity of the covariance matrix inversion, which is then limited to $\mathcal{O}(N^2M)$ through variational compression.
 For the Bayesian GP-LVM the motivation is to be able to tractably integrate over the uncertainty associated with the unknown latent input locations.
@@
   \item Algorithm in section \ref{sec:gplvmcmf}
+  \item Discussion of the aux variables acting as a stochastic hidden layer which provides smoothing...
+  \item Moreover, the additional KL penalisation between the variational posterior and the auxiliary distribution prevents the variational model trying to fit directly to the true posterior and instead does so via fitting its denominator (the var posterior) to the aux dist.
+The aux params are generated from both the latent variables and the data, acting as an intermediate between the generative likelihood model and the marginal variational model.
+  \item Links from above to batch normalisation - the auxiliary variables act like pseudo data which are global (not batch-specific).
 \end{itemize}

### 2016-02-28 GP_LVM_CMF.py

@@ -9,6 +9,7 @@ import time
 import collections
 from myCond import myCond
 from printable import printable
+import nnet
 
 from utils import cholInvLogDet, sharedZeroArray, sharedZeroMatrix, sharedZeroVector, \
     np_log_mean_exp_stable, diagCholInvLogDet_fromLogDiag, diagCholInvLogDet_fromDiag, \
@@ -109,7 +110,7 @@ class SGPDV(printable):
         # kernel parameters
         self.log_theta = sharedZeroMatrix(
             1, self.numberOfKernelParameters, 'log_theta', broadcastable=(True, False))  # parameters of Kuu, Kuf, Kff
-        
+
         # Random variables
         alpha = srng.normal(size=(self.B, self.R), avg=0.0, std=1.0, ndim=None)
         beta = srng.normal(size=(self.B, self.Q), avg=0.0, std=1.0, ndim=None)
@@ -121,7 +122,7 @@ class SGPDV(printable):
 
         self.mlp_qX = MLP_Network(self.P, self.R, self.H, 'qX')
         self.mu_qX, self.log_sigma_qX = self.mlp1_qX.setup(self.y_miniBatch.T)
-        
+
         # Variational distribution q(u)
         self.kappa = sharedZeroMatrix(self.M, self.Q, 'kappa')
         self.Kappa_sqrt = sharedZeroMatrix(self.M, self.M, 'Kappa_sqrt')
@@ -162,7 +163,7 @@ class SGPDV(printable):
         self.Kuu_conditionNumber   = conditionNumber(self.Kuu)
         self.Sigma_conditionNumber = conditionNumber(self.Sigma)
 
-    def self.construct_rX(z):
+    def construct_rX(z):
 
         self.rfXf_mlp = MLP_Network(self.Q + self.P, self.Q + self.R, 1, self.H, Softplus, 'rfXf'):
         self.mu_rfXf, self.log_sigma_rfXf = self.rfXf_mlp.setup(T.concatenate((z.T, self.y_miniBatch.T)))
@@ -182,7 +183,7 @@ class SGPDV(printable):
         fXf = T.concatenate((self.f, self.Xf), axis=1).T
 
         fX_m_mu = minus(fXf, self.mu_rfXf)
-        
+
         self.log_rfXf_zy = -0.5 * (self.R+self.Q) * self.B * log2pi \
             - T.sum(self.log_sigma_rfXf)
             - 0.5 * T.sum( div(fX_m_mu**2, T.exp(2*self.log_sigma_rfXf)))

### 2016-02-28 Hybrid_variational_model.py

@@ -11,6 +11,7 @@ import theano.tensor as T
 from theano.tensor import nlinalg
 from printable import Printable
 from utils import plus, mul
+import nnet
 
 from GP_LVM_CMF import SGPDV
 
@@ -35,7 +36,7 @@ class Hybrid_encoder(Printable):
 
         self.mu_encoder, self.log_sigma_encoder2 \
             = self.mlp_encoder.setup(T.concatenate((self.gp_encoder.f, y_miniBatch)))
-         
+
         gamma = srng.normal(size=(dimZ, miniBatchSize), avg=0.0, std=1.0, ndim=None)
         gamma.name = 'gamma'
         self.sample_gamma = th.function([], gamma)

### 2016-02-28 Hysterisis_variational_model.py

@@ -11,6 +11,7 @@ import theano.tensor as T
 
 from printable import Printable
 from utils import plus, mul
+import nnet
 
 precision = th.config.floatX
 
@@ -28,7 +29,7 @@ class Hysterisis_encoder(Printable):
         self.mlp_f_y = MLP_Network(self, dimY, dimZ,
                 numHiddenUnits_encoder, 'encoder', num_layers=numHiddenLayers_encoder)
         self.mu_f_y, self.log_sigma_f_y = self.mlp_f_y.setup(y_miniBatch)
- 
+
         alpha = srng.normal(size=(dimZ, minbatchSize), avg=0.0, std=1.0, ndim=None)
         alpha.name = 'alpha'
         self.sample_alpha = th.function([], alpha)
@@ -38,7 +39,7 @@ class Hysterisis_encoder(Printable):
         self.mlp_z_fy = MLP_Network(self, dimY, dimZ,
                 numHiddenUnits_encoder, 'encoder', num_layers=numHiddenLayers_encoder)
         self.mu_qz_fy, self.log_sigma_qz_fy = self.mlp_z_fy.setup(T.concatenate((self.f, y_miniBatch)))
-         
+
         beta = srng.normal(size=(dimZ, minbatchSize), avg=0.0, std=1.0, ndim=None)
         beta.name = 'beta'
         self.sample_beta = th.function([], beta)
@@ -52,7 +53,7 @@ class Hysterisis_encoder(Printable):
         self.gradientVariables = mlp.mlp_f_y.params + mlp.mlp_z_fy.params + self.mlp_rf_yz.params
 
     def construct_L_terms():
-        
+
         self.H_f_y = 0.5 * self.B * (1+log2pi) + T.sum(log_sigma_f_y)
         self.L_terms =  self.H_f_y
 

### 2016-02-28 MLP_likelihood_model.py

@@ -9,6 +9,7 @@ import numpy as np
 import theano as th
 import theano.tensor as T
 from printable import printable
+from utils import plus, div
 
 log2pi = np.log(2 * np.pi)
 
@@ -41,9 +42,10 @@ class MLP_likelihood_model(Printable):
                    - self.B
 
         if self.continuous:
-            self.log_p_y_z = 
+            self.log_pyz = T.sum( -(0.5*log2pi + T.sum(self.log_sigma_decoder)) \
+            - 0.5 * ((y_miniBatch.T - self.mu_decoder) / T.exp(self.log_sigma_decoder))**2 )
         else:
-            pass
+            self.log_pyz = -T.nnet.binary_crossentropy(self.mu_decoder, y_miniBatch).sum()
 
 
         self.L_terms = self.log_pyz + self.KL

### 2016-02-28 MLP_likelihood_model.py
Bug in log_pyz

@@ -42,7 +42,7 @@ class MLP_likelihood_model(Printable):
                    - self.B
 
         if self.continuous:
-            self.log_pyz = T.sum( -(0.5*log2pi + T.sum(self.log_sigma_decoder)) \
+            self.log_pyz = T.sum( -(0.5*log2pi + self.log_sigma_decoder) \
             - 0.5 * ((y_miniBatch.T - self.mu_decoder) / T.exp(self.log_sigma_decoder))**2 )
         else:
             self.log_pyz = -T.nnet.binary_crossentropy(self.mu_decoder, y_miniBatch).sum()

### 2016-02-28 Auto_encoder_model.py
Bugs bugs bugs, started unittesting.

@@ -13,6 +13,10 @@ from optimisers import Adam
 from GP_LVM_CMF import SGPDV
 from testTools import checkgrad
 from utils import log_mean_exp_stable, dot, trace, softplus, sharedZeroVector, sharedZeroMatrix, plus
+from Hybrid_variational_model import MLP_encoder
+from MLP_variational_model import MLP_variational_model
+from Hysterisis_variational_model import Hysterisis_encoder
+from jitterProjected import jitterProjected
 
 precision = th.config.floatX
 
@@ -29,7 +33,7 @@ class AutoEncoderModel(Printable):
         # set the data
         data = np.asarray(data, dtype=precision)
         self.N = data.shape[0]
-        self.P = data.shape[1] 
+        self.P = data.shape[1]
         self.B = params['miniBatchSize']
 
         self.numberofBatchesPerEpoch = int(np.ceil(np.float32(self.N) / self.B))
@@ -61,28 +65,28 @@ class AutoEncoderModel(Printable):
         self.lowerBound = -np.inf  # Lower bound
         self.lowerBounds = []
 
-        self.jitterProtector = jitterProtector()
+        self.jitterProtector = jitterProtected()
         if encoderType == 'Hybrid':
             self.encoder = Hybrid_variational_model(
                 self.y_miniBatch,
                 self.B,
                 self.jitterProtector,
-                encoderParams)
+                encoderParameters)
         elif encoderType == 'MLP':
             self.encoder = MLP_variational_model(
                 self.y_minBatch,
                 self.B,
-                encoderParams)
+                encoderParameters)
         elif encoderType == 'Hysterisis':
                 self.encoder = Hysterisis_variational_model(
                 self.y_minBatch,
                 self.B,
-                encoderParams)
+                encoderParameters)
         else:
             raise RuntimeErorr('Unrecognised encoder type')
 
         if decoderType == 'MLP':
-            self.decoder(decoderParams, self.y_miniBatch, decoderParams)
+            self.decoder(decoderParameters, self.y_miniBatch, decoderParameters)
         else:
             raise RuntimeErorr('Unrecognised decoder type')
 
@@ -186,3 +190,33 @@ class AutoEncoderModel(Printable):
                 c += 1
 
         return np_log_mean_exp_stable(ll)
+
+
+if __name__ == "__main__":
+
+    params['miniBatchSize'] = 2
[…]

### 2016-02-28 Hybrid_variational_model.py

@@ -12,14 +12,13 @@ from theano.tensor import nlinalg
 from printable import Printable
 from utils import plus, mul
 import nnet
+from jitterProjected import jitterProjected
 
 from GP_LVM_CMF import SGPDV
 
 class Hybrid_encoder(Printable):
 
-    def __init__(
-
-        def __init__(self, y_miniBatch, minbatchSize, jitterProtect, dimY, dimZ, params):
+    def __init__(self, y_miniBatch, minbatchSize, jitterProtect, dimY, dimZ, params):
 
         HU = params['numHiddenUnits_encoder']
         HL = params['numHiddenLayers_encoder']
@@ -59,5 +58,21 @@ class Hybrid_encoder(Printable):
     def randomise(self):
         self.gp_encoder.randomise()
         self.mlp_encoder.randomise()
+
+
+if __name__ == "__main__":
+    params['numHiddenUnits_encoder'] = 10
+    params['numHiddenLayers_encoder'] = 1
+    y_miniBatch = np.ones((2,2))
+    miniBatchSize = 2
+    jitterProtect = jitterProjected()
+    dimY = 2
+    dimZ = 2
+
+    hybrid = Hybrid_encoder(y_miniBatch, minbatchSize, jitterProtect, dimY, dimZ, params)
+
+    hybrid.construct_L_terms()
+    hybrid.sample()
+    hybrid.randomise()
 
 

### 2016-02-28 Hysterisis_variational_model.py

@@ -65,3 +65,23 @@ class Hysterisis_encoder(Printable):
         self.mlp_f_y.randomise()
         self.mlp_z_fy.randomise()
         self.mlp_rf_yz.randomise()
+
+
+if __name__ == "__main__":
+    params['numHiddenUnits_encoder'] = 10
+    params['numHiddenLayers_encoder'] = 1
+    y_miniBatch = np.ones((2,2))
+    miniBatchSize = 2
+    jitterProtect = jitterProjected()
+    dimY = 2
+    dimZ = 2
+
+    hyst = Hysterisis_encoder(y_miniBatch, minbatchSize, dimY, dimZ, params)
+
+    hyst.construct_L_terms()
+    hyst.sample()
+    hyst.randomise()
+
+
+
+

### 2016-02-28 MLP_likelihood_model.py

@@ -53,3 +53,28 @@ class MLP_likelihood_model(Printable):
     def randomise(self):
         self.mlp_decoder.randomise()
 
+
+if __name__ == "__main__":
+    enc_params['numHiddenUnits_encoder'] = 10
+    enc_params['numHiddenLayers_encoder'] = 1
+    enc_params['continuous'] = True
+    y_miniBatch = np.ones((2,2))
+    miniBatchSize = 2
+    dimY = 2
+    dimZ = 2
+
+    from MLP_variational_model import MLP_variational_model
+    encoder = MLP_variational_model(y_miniBatch, minBatchSize, dimY, dimZ, enc_params)
+
+    dec_params['numHiddenUnits_decoder'] = 10
+    dec_params['numHiddenLayers_decoder'] = 1
+
+    decoder = MLP_likelihood_model(y_miniBatch, miniBatchSize, dimY, dimZ, encoder, dec_params)
+
+    decoder.construct_L_terms()
+    decoder.randomis()
+
+
+
+
+

### 2016-02-28 MLP_variational_model.py

@@ -12,23 +12,24 @@ from theano.tensor import nlinalg
 
 from testTools import checkgrad
 from utils import log_mean_exp_stable
+import nnet
 
 precision = th.config.floatX
 
 class MLP_variational_model(Printable):
 
-    def __init__(self, y_miniBatch, minBatchSize, dimY, dimZ, params):
+    def __init__(self, y_miniBatch, miniBatchSize, dimY, dimZ, params):
 
-    	self.B = minBatchSize
+    	self.B = miniBatchSize
 
     	numHiddenUnits_encoder = params['numHiddenUnits_encoder']
         numHiddenLayers_encoder = params['numHiddenLayers_encoder']
- 
+
         self.mlp_encoder = MLP_Network(self, dimY, dimZ,
                 numHiddenUnits_encoder, 'encoder', num_layers=numHiddenLayers_encoder)
 
         self.mu_encoder, self.log_sigma_encoder = self.mlp_encoder.setup(y_miniBatch)
-         
+
         alpha = srng.normal(size=(dimZ, self.B), avg=0.0, std=1.0, ndim=None)
         alpha.name = 'alpha'
         self.sample_alpha = th.function([], alpha)
@@ -43,7 +44,19 @@ class MLP_variational_model(Printable):
     def sample(self):
         self.sample_alpha()
 
-        
+
+if __name__ == "__main__":
+    y_miniBatch = np.ones((2,2))
+    miniBatchSize = 2
+    dimY = 2
+    dimZ = 2
+    enc_params['numHiddenUnits_encoder'] = 10
+    enc_params['numHiddenLayers_encoder'] = 1
+
+    encoder = MLP_variational_model(y_miniBatch, miniBatchSize, dimY, dimZ, enc_params)
+
+    encoder.construct_L_terms()
+    encoder.sample()
 
 
 
@@ -51,5 +64,6 @@ class MLP_variational_model(Printable):
 
 
 
- 
+
+
 

### 2016-02-28 MLP_likelihood_model.py
More bugs.

@@ -47,7 +47,6 @@ class MLP_likelihood_model(Printable):
         else:
             self.log_pyz = -T.nnet.binary_crossentropy(self.mu_decoder, y_miniBatch).sum()
 
-
         self.L_terms = self.log_pyz + self.KL
 
     def randomise(self):

### 2016-02-28 MLP_variational_model.py

@@ -34,12 +34,14 @@ class MLP_variational_model(Printable):
         alpha.name = 'alpha'
         self.sample_alpha = th.function([], alpha)
 
-        self.gradientVariables = mlp.encoder.params
+        self.gradientVariables = self.encoder.params
 
         self.z = plus(self.mu_encoder, mul(T.exp(self.log_sigma_encoder*0.5), alpha))
 
     def construct_L_terms():
-    	pass
+        self.H = 0.5 * self.B * (1+log2pi) + T.sum(self.log_sigma_encoder)
+
+        self.L_terms =  self.H
 
     def sample(self):
         self.sample_alpha()

### 2016-02-28 Hybrid_variational_model.py
MLP_encoder seems to work ok

@@ -19,8 +19,8 @@ class Hybrid_variational_model(Printable):
 
     def __init__(self, y_miniBatch, minbatchSize, jitterProtect, dimY, dimZ, params, srng):
 
-        HU = params['numHiddenUnits_encoder']
-        HL = params['numHiddenLayers_encoder']
+        num_units = params['numHiddenUnits_encoder']
+        num_layers = params['numHiddenLayers_encoder']
 
         self.gp_encoder = SGPDV(y_miniBatch,
                                 miniBatchSize,
@@ -30,14 +30,13 @@ class Hybrid_variational_model(Printable):
                                 params,
                                 srng)
 
-        self.srng = srng
-
-        self.mlp_encoder = MLP_Network(self, dimY, dimZ, HU, 'encoder', num_layers=HL)
+        self.mlp_encoder = MLP_Network(self, dimY, dimZ, name='Hybrid_encoder', 
+            num_units=num_units, num_layers=num_layers)
 
         self.mu_qz, self.log_sigma_qz \
             = self.mlp_encoder.setup(T.concatenate((self.gp_encoder.f, y_miniBatch)))
 
-        gamma = self.srng.normal(size=(dimZ, miniBatchSize), avg=0.0, std=1.0, ndim=None)
+        gamma = srng.normal(size=(dimZ, miniBatchSize), avg=0.0, std=1.0, ndim=None)
         gamma.name = 'gamma'
         self.sample_gamma = th.function([], gamma)
 

### 2016-02-28 Hysteresis_variational_model.py

@@ -26,11 +26,11 @@ class Hysteresis_variational_model(Printable):
 
         self.B = minbatchSize
 
-        numHiddenUnits_encoder = params['numHiddenUnits_encoder']
-        numHiddenLayers_encoder = params['numHiddenLayers_encoder']
+        num_units = params['numHiddenUnits_encoder']
+        num_layers = params['numHiddenLayers_encoder']
 
-        self.mlp_f_y = MLP_Network(self, dimY, dimZ,
-                numHiddenUnits_encoder, 'encoder', num_layers=numHiddenLayers_encoder)
+        self.mlp_f_y = MLP_Network(dimY, dimZ, name='Hyster_hidden',
+                num_units=num_units, num_layers=num_layers)
         self.mu_f_y, self.log_sigma_f_y = self.mlp_f_y.setup(y_miniBatch)
 
         alpha = self.srng.normal(size=(dimZ, minbatchSize), avg=0.0, std=1.0, ndim=None)
@@ -39,9 +39,10 @@ class Hysteresis_variational_model(Printable):
 
         self.f = plus(self.mu_f_y, mul(T.exp(self.log_sigma_f_y*0.5), alpha), 'f')
 
-        self.mlp_z_fy = MLP_Network(self, dimY, dimZ,
-                numHiddenUnits_encoder, 'encoder', num_layers=numHiddenLayers_encoder)
-        self.mu_qz, self.log_sigma_qz = self.mlp_z_fy.setup(T.concatenate((self.f, y_miniBatch)))
+        self.mlp_z_fy = MLP_Network(dimY, dimZ, name='Hyster_output',
+                num_units=num_units, num_layers=num_layers)
+        self.mu_qz, self.log_sigma_qz \
+            = self.mlp_z_fy.setup(T.concatenate((self.f, y_miniBatch)))
 
         beta = self.srng.normal(size=(dimZ, minbatchSize), avg=0.0, std=1.0, ndim=None)
         beta.name = 'beta'
@@ -49,9 +50,10 @@ class Hysteresis_variational_model(Printable):
 
         self.z = plus(self.mu_qz, mul(T.exp(self.log_sigma_qz*0.5), beta), 'z')
 
-        self.mlp_rf_yz = MLP_Network(self, (dimY+dimZ), dimZ,
-                numHiddenUnits_encoder, 'encoder', num_layers=numHiddenLayers_encoder)
[…]

### 2016-02-28 MLP_likelihood_model.py

@@ -20,17 +20,17 @@ class MLP_likelihood_model(Printable):
         self.B = miniBatchSize
         self.y_miniBatch = y_miniBatch
         
-        numHiddenUnits_decoder = params['numHiddenUnits_decoder']
-        numHiddenLayers_decoder = params['numHiddenLayers_decoder']
+        num_units  = params['numHiddenUnits_decoder']
+        num_layers = params['numHiddenLayers_decoder']
         self.continuous = params['continuous']
 
         if self.continuous:
-            self.mlp_decoder = MLP_Network(dimZ, dimY,
-                numHiddenUnits_decoder, 'decoder', num_layers=numHiddenLayers_decoder, continuous=True)
+            self.mlp_decoder = MLP_Network(dimZ, dimY, name='mlp_decoder',
+                num_units=num_units, num_layers=num_layers, continuous=True)
             self.mu_decoder, self.log_sigma_decoder = self.mlp_decoder.setup(encoder.z)
         else:
-            self.mlp_decoder = MLP_Network(dimZ, dimY,
-                numHiddenUnits_decoder, 'decoder', num_layers=numHiddenLayers_decoder, continuous=False)
+            self.mlp_decoder = MLP_Network(dimZ, dimY, name='mlp_decoder',
+                num_units=num_units, num_layers=num_layers, continuous=False)
             self.yhat = self.mlp_decoder.setup(encoder.z)
 
         self.gradientVariables = self.mlp_decoder.params

### 2016-02-28 MLP_variational_model.py

@@ -21,33 +21,30 @@ class MLP_variational_model(Printable):
 
     def __init__(self, y_miniBatch, miniBatchSize, dimY, dimZ, params, srng):
 
-        self.srng = srng
-
         self.B = miniBatchSize
-        numHiddenUnits_encoder = params['numHiddenUnits_encoder']
-        numHiddenLayers_encoder = params['numHiddenLayers_encoder']
+        num_units  = params['numHiddenUnits_encoder']
+        num_layers = params['numHiddenLayers_encoder']
 
-        self.mlp_encoder = MLP_Network(dimY, dimZ,
-                numHiddenUnits_encoder, 'encoder', num_layers=numHiddenLayers_encoder)
+        self.mlp_encoder = MLP_Network(dimY, dimZ, name='MLP_encoder',
+                num_units=num_units, num_layers=num_layers)
 
-        self.mu_qz, self.log_sigma_qz = self.mlp_encoder.setup(y_miniBatch)
+        self.mu_qz, self.log_sigma_qz = self.mlp_encoder.setup(y_miniBatch.T)
 
-        alpha = self.srng.normal(size=(dimZ, self.B), avg=0.0, std=1.0, ndim=None)
+        alpha = srng.normal(size=(dimZ, self.B), avg=0.0, std=1.0, ndim=None)
         alpha.name = 'alpha'
         self.sample_alpha = th.function([], alpha)
 
-        self.gradientVariables = self.mlp_encoder.params
-
         self.z = plus(self.mu_qz, mul(exp(self.log_sigma_qz), alpha), 'z')
 
-    def construct_L_terms(self):
-        self.H = 0.5 * self.B * (1+log2pi) + T.sum(self.log_sigma_qz)
+        self.gradientVariables = self.mlp_encoder.params
 
-        self.L_terms =  self.H
+    def construct_L_terms(self):
+        self.H_qz = 0.5 * self.B * (1+log2pi) + T.sum(self.log_sigma_qz)
+
+        self.L_terms =  self.H_qz
 
     def sample(self):
         self.sample_alpha()
-
 
 if __name__ == "__main__":
     y_miniBatch = np.ones((2,2))

### 2016-02-28 nnet.py

@@ -1,9 +1,9 @@
 '''Classes used for describing and creating neural networks'''
 
 import theano
-import theano.tensor as T
 import numpy as np
-from utils import tanh, sigmoid, softplus, exp, plus, mul, dot, sharedZeroMatrix
+from utils import tanh, sigmoid, softplus, exp, plus, dot
+from utils import sharedZeroMatrix, sharedZeroVector
 
 floatX = theano.config.floatX
 
@@ -12,8 +12,8 @@ class Linear():
     def __init__(self, dim_in, dim_out, name):
         self.dim_in  = dim_in
         self.dim_out = dim_out
-        self.W = sharedZeroMatrix(dim_in, dim_out, 'W_' + name)
-        self.b = sharedZeroMatrix(1, dim_out, 'b_' + name, broadcastable=(True, False))
+        self.W = sharedZeroMatrix(dim_out, dim_in, 'W_' + name)
+        self.b = sharedZeroVector(dim_out, 'b_' + name, broadcastable=(False, True))
         self.params = [self.W, self.b]
 
     def setup(self, x_in, **kwargs):
@@ -94,25 +94,25 @@ class NNet():
 
 class MLP_Network():
 
-    def __init__(self, dim_in, dim_out, num_hidden, name, num_layers=1, continuous=True, nonlinearity=Softplus):
+    def __init__(self, dim_in, dim_out, name='', num_units=10, num_layers=1, continuous=True, nonlinearity=Softplus):
 
         self.nonlinearity = nonlinearity()
         self.name = name
         self.continuous = continuous
         self.hidden = NNet()
-        self.hidden.addLayer(Linear(dim_in, num_hidden,'hidden_'+str(0)+'_'+name))
+        self.hidden.addLayer(Linear(dim_in, num_units,'hidden_'+str(0)+'_'+name))
         self.hidden.addLayer(self.nonlinearity)
 
         for i in range(1,num_layers):
-            self.hidden.addLayer(Linear(num_hidden, num_hidden, 'hidden_'+str(i)+'_'+name))
+            self.hidden.addLayer(Linear(num_units, num_units, 'hidden_'+str(i)+'_'+name))
             self.hidden.addLayer(self.nonlinearity)
 
         if self.continuous:
-            self.muLinear = Linear(num_hidden, dim_out, 'muLinear_' + name)
-            self.logsigmaLinear = Linear(num_hidden, dim_out, 'logsigmaLinear_' + name)
[…]

### 2016-02-29 Auto_encoder_model.py
More bugs fixed, few utils added

@@ -39,8 +39,10 @@ class AutoEncoderModel(Printable):
         self.P = data.shape[1]
         self.B = params['miniBatchSize']
         self.Q = params['dimZ']
+        theanoRandomSeed = params['theanoRandomSeed']
+        numpyRandomSeed = params['numpyRandomSeed']
 
-        self.srng = createSrng(seed=123)
+        self.srng = createSrng(seed=theanoRandomSeed)
 
         self.numberofBatchesPerEpoch = int(np.ceil(np.float32(self.N) / self.B))
         numPad = self.numberofBatchesPerEpoch * self.B - self.N
@@ -116,7 +118,12 @@ class AutoEncoderModel(Printable):
         self.dL = T.grad(self.L, self.gradientVariables)
         for i in range(len(self.dL)):
             self.dL[i].name = 'dL_d' + self.gradientVariables[i].name
-
+        
+        # Initialise the variables in the networks
+        rnd = np.random.RandomState(seed=numpyRandomSeed)
+        self.encoder.randomise(rnd)
+        self.decoder.randomise(rnd)
+            
     def sample(self):
         self.sample_batchStream()
         self.sample_padStream()
@@ -201,10 +208,10 @@ class AutoEncoderModel(Printable):
         for i in range(self.numberofBatchesPerEpoch):
             print '{} of {}, {} samples'.format(i, self.numberofBatchesPerEpoch, numberOfTestSamples)
             self.iterator.set_value(i)
-            self.jitter.set_value(self.jitterDefault)
+            self.jitterProtector.reset()
             for k in range(numberOfTestSamples):
                 self.sample()
-                ll[c] = self.jitterProtect(self.L_func, reset=False)
+                ll[c] = self.jitterProtector.jitterProtect(self.L_func, reset=False)
                 c += 1
 
         return np_log_mean_exp_stable(ll)

### 2016-02-29 Hybrid_variational_model.py

@@ -55,7 +55,7 @@ class Hybrid_variational_model(Printable):
         self.gp_encoder.sample_beta()
         self.gamma.sample()
 
-    def randomise(self):
+    def randomise(self, rnd):
         self.gp_encoder.randomise()
         self.mlp_encoder.randomise()
 

### 2016-02-29 Hysteresis_variational_model.py

@@ -10,7 +10,7 @@ import theano as th
 import theano.tensor as T
 
 from printable import Printable
-from utils import plus, mul, createSrng
+from utils import plus, mul, createSrng, log_elementwiseNormal, elementwiseNormalEntropy
 from nnet import MLP_Network
 from jitterProtect import JitterProtect
 
@@ -25,6 +25,7 @@ class Hysteresis_variational_model(Printable):
         self.srng = srng
 
         self.B = minbatchSize
+        self.Q = dimZ
 
         num_units = params['numHiddenUnits_encoder']
         num_layers = params['numHiddenLayers_encoder']
@@ -59,8 +60,14 @@ class Hysteresis_variational_model(Printable):
 
     def construct_L_terms(self):
 
-        self.H_f_y = 0.5 * self.B * (1+log2pi) + T.sum(self.log_sigma_f_y)
-        self.L_terms =  self.H_f_y
+        self.H_f_y = elementwiseNormalEntropy(self.log_sigma_f_y, 'H_f_y')
+        
+        self.log_r_f_zy = log_elementwiseNormal(self.f,
+                                                self.mu_rf_yz,
+                                                self.log_sigma_rf_yz,
+                                                'log_sigma_rf_yz')
+        
+        self.L_terms = plus(self.log_r_f_zy, self.H_f_y)
 
     def sample(self):
         self.sample_alpha()

### 2016-02-29 MLP_likelihood_model.py

@@ -8,7 +8,7 @@ Created on Thu Feb  4 20:44:40 2016
 import numpy as np
 import theano.tensor as T
 from printable import Printable
-from utils import plus, exp, minus
+from utils import plus, exp, minus, mul, log_elementwiseNormal
 from nnet import MLP_Network
 
 log2pi = np.log(2 * np.pi)
@@ -18,6 +18,7 @@ class MLP_likelihood_model(Printable):
     def __init__(self, y_miniBatch, miniBatchSize, dimY, dimZ, encoder, params):
 
         self.B = miniBatchSize
+        self.Q = dimZ
         self.y_miniBatch = y_miniBatch
         
         num_units  = params['numHiddenUnits_decoder']
@@ -37,21 +38,23 @@ class MLP_likelihood_model(Printable):
 
     def construct_L_terms(self, encoder):
 
-        self.KL_qp = 0.5*(T.sum(exp(encoder.log_sigma_qz*2))) \
+        self.KL_qp = 0.5*(T.sum(exp(mul(encoder.log_sigma_qz,2)))) \
                    + 0.5 * T.sum(encoder.mu_qz**2) \
                    - T.sum(encoder.log_sigma_qz) \
-                   - self.B
+                   - 0.5*self.Q*self.B
 
         if self.continuous:
-            self.log_pyz = T.sum( -(0.5*log2pi + self.log_sigma_decoder) \
-            - 0.5 * ((self.y_miniBatch.T - self.mu_decoder) / T.exp(self.log_sigma_decoder))**2 )
+             self.log_pyz = log_elementwiseNormal(self.y_miniBatch.T,
+                                                  self.mu_decoder,
+                                                  self.log_sigma_decoder,
+                                                  'log_pyz')
         else:
             self.log_pyz = -T.nnet.binary_crossentropy(self.mu_decoder, y_miniBatch).sum()
 
-        self.L_terms = self.log_pyz + self.KL_qp
+        self.L_terms = self.log_pyz - self.KL_qp
 
-    def randomise(self, srng):
-        self.mlp_decoder.randomise(srng)
+    def randomise(self, rng):
+        self.mlp_decoder.randomise(rng)
 
 
 if __name__ == "__main__":

### 2016-02-29 MLP_variational_model.py

@@ -22,6 +22,7 @@ class MLP_variational_model(Printable):
     def __init__(self, y_miniBatch, miniBatchSize, dimY, dimZ, params, srng):
 
         self.B = miniBatchSize
+        self.Q = dimZ
         num_units  = params['numHiddenUnits_encoder']
         num_layers = params['numHiddenLayers_encoder']
 
@@ -39,12 +40,13 @@ class MLP_variational_model(Printable):
         self.gradientVariables = self.mlp_encoder.params
 
     def construct_L_terms(self):
-        self.H_qz = 0.5 * self.B * (1+log2pi) + T.sum(self.log_sigma_qz)
-
-        self.L_terms =  self.H_qz
+        self.L_terms = 0
 
     def sample(self):
         self.sample_alpha()
+        
+    def randomise(self, rng):
+        self.mlp_encoder.randomise(rng)
 
 if __name__ == "__main__":
     y_miniBatch = np.ones((2,2))

### 2016-02-29 jitterProtect.py

@@ -22,4 +22,7 @@ class JitterProtect(object):
             self.jitter.set_value(self.jitterDefault)
         return val
 
+    def reset(self):
+        self.jitter.set_value(self.jitterDefault)
+
 

### 2016-02-29 nnet.py

@@ -19,7 +19,7 @@ class Linear():
     def setup(self, x_in, **kwargs):
         return plus(dot(self.W, x_in), self.b)
 
-    def randomise(self, rnd, factor=1., nonlinearity=None):
+    def randomise(self, rnd, factor=1.0, nonlinearity=None):
         '''A randomly initialized linear layer.
         When factor is 1, the initialization is uniform as in Glorot, Bengio, 2010,
         assuming the layer is intended to be followed by the tanh nonlinearity.'''
@@ -30,10 +30,11 @@ class Linear():
                                      size=(self.n_in, self.n_out)))
             self.b.set_value(np.zeros((1, self.n_out)))
         elif type(nonlinearity) == Softplus:
-            # Hidden layer weights are uniformly sampled from a symmetric interval
-            # following [Xavier, 2010] those for the sigmoid transform
-            pass
-
+            scale = factor * np.sqrt(6./(self.dim_in+self.dim_out))
+            self.W.set_value(rnd.uniform(low=-scale,
+                                     high=scale,
+                                     size=(self.n_in, self.n_out)))
+            self.b.set_value(np.zeros((1, self.n_out)))
         elif type(nonlinearity) == Linear:
             raise RuntimeError('Consecutive linear layers')
 
@@ -82,7 +83,7 @@ class NNet():
             y = layer.setup(y, **kwargs)
         return y
 
-    def randomise(self, factor, rnd):
+    def randomise(self, rnd, factor=1.0):
         for i in range(len(self.layers)):
             if type(self.layers[i]) == Linear:
                 if i < len(self.layers):
@@ -122,7 +123,7 @@ class MLP_Network():
             mu = self.muLinear.setup(h_outin, **kwargs)
             logsigma = 0.5 * self.logsigmaLinear.setup(h_outin, **kwargs)
             mu.name = 'mu_' + self.name
-            logsigma.name = 'logsimga' + self.name
+            logsigma.name = 'logsimga_' + self.name
             return (mu, logsigma)
         else:
             h2 = self.yhatLinear.setup(h_outin, **kwargs)
@@ -131,9 +132,9 @@ class MLP_Network():
             return yhat
 
     def randomise(self, rnd):
-        self.hidden.randomise()
+        self.hidden.randomise(rnd)
         if self.continuous:
-            self.muLinear.randomise()
-            self.logsigmaLinear.randomise()
+            self.muLinear.randomise(rnd)
+            self.logsigmaLinear.randomise(rnd)
         else:
             self.yhatLinear.randomise()

### 2016-02-29 utils.py

@@ -3,8 +3,12 @@ import theano as th
 from theano import tensor as T
 from theano.tensor import nlinalg, slinalg
 from fastlin.myCholesky import myCholesky
+from fastlin.myCond import myCond
 from theano.sandbox.rng_mrg import MRG_RandomStreams
 from theano.tensor.shared_randomstreams import RandomStreams
+
+log2pi = T.constant(np.log(2 * np.pi))
+log2pi.name = 'log(2pi)'
 
 def t_repeat(x, num_repeats, axis):
     '''Repeats x along an axis num_repeats times. Axis has to be 0 or 1, x has to be a matrix.'''
@@ -189,3 +193,22 @@ def conditionNumber(M):
     condNum = th.function(
             [], cond, no_default_updates=True)
     return condNum
+
+
+def log_elementwiseNormal(x, mu, log_sigma, name):
+    
+    d = minus(x, mu)
+    d2 = mul(d,d)
+    sigma2 = exp(mul(log_sigma,2))
+    
+    lg = T.sum( -0.5*log2pi - log_sigma - 0.5 * div(d2, sigma2) )
+    lg.name = name
+    return lg
+
+def elementwiseNormalEntropy(log_sigma, name):
+    H =  T.sum(0.5 * (1+log2pi) + log_sigma)
+    H.name = name
+    return H
+         
+         
+         

### 2016-02-29 trainmnist_VAE.py
minor change to set random seeds

@@ -23,7 +23,8 @@ learning_rate = 1e-3
 
 data = x_train
 dimZ = 40
-params = {'miniBatchSize' : 200, 'dimZ':400}
+params = {'miniBatchSize' : 200, 'dimZ':400,
+         'theanoRandomSeed':123, 'numpyRandomSeed':123}
 encoderType='MLP'
 decoderType='MLP'
 encoderParameters = {'numHiddenUnits_encoder' : 400, 'numHiddenLayers_encoder' : 1}

### 2016-02-29 trainmnist_VAE.py
changed num epochs

@@ -16,7 +16,7 @@ f = gzip.open('mnist.pkl.gz', 'rb')
 f.close()
 
 
-numberOfEpochs = 1
+numberOfEpochs = 10
 learning_rate = 1e-3
 
 [N,dimY] = x_train.shape

### 2016-03-04 icml2016/paper.tex
Added parmesan and my first attempt at vae with 2 stochastic hidden layers

@@
 \newcommand{\theHalgorithm}{\arabic{algorithm}}
+\newcommand{\chris}[1]{\textcolor{ForestGreen}{#1}}
+\newcommand{\Tau}{\mathcal{T}}
+\newcommand{\Kappa}{\mathcal{K}}
+\newcommand{\xb}{\mathbf{x}}
+\newcommand{\ub}{\mathbf{u}}
+\newcommand{\zb}{\mathbf{z}}
+\newcommand{\fb}{\mathbf{f}}
+\newcommand{\Cb}{\mathbf{C}}
+\newcommand{\Ab}{\mathbf{A}}
+\newcommand{\yb}{\mathbf{y}}
+\newcommand{\dd}{\mathrm{d}}
+\newcommand{\Lo}{\mathcal{L}_1}
+\newcommand{\Laux}{\mathcal{L}_{\mathrm{aux}}}
+\newcommand{\Kff}{\mathbf{K}_{ff}}
+\newcommand{\Kuu}{\mathbf{K}_{uu}}
+\newcommand{\Kuf}{\mathbf{K}_{uf}}
+\newcommand{\Kfu}{\mathbf{K}_{fu}}
 \newcommand{\Ex}{\mathbb{E}}
-\newcommand{\KL}{\textrm{KL}}
-% I would prefer f or \mathbf{f} rather than F()
-\newcommand{\f}{f}
+\newcommand{\KL}{\mathrm{KL}}
+\newcommand{\No}{\mathcal{N}}
+\newcommand{\chol}{\mathrm{chol}}
+\newcommand{\tr}{\mathrm{tr}}
+\newcommand{\Wb}{\mathbf{W}}
+\newcommand{\hb}{\mathbf{h}}
+\newcommand{\bb}{\mathbf{b}}
+\newcommand{\bx}{\mathbf{x}}
 
@@
 
-In addition to the GP-LVM, we introduce an auxiliary neural net approach which naturally extends the variational auto-encoder to a hierarchical version which is able to the bottom-up stochastic message passing of the variational encoder with top-down stochastic messages from a variational auxiliary decoder.
-
-From the perspective of the variational principle, we propose a drop-in replacement for mean field inference in which a rich, variational distribution is specified in the expanded space of variables we wish to infer augmented with auxiliary inference variables.
+From the perspective of the variational principle, we propose a drop-in replacement for mean field inference in which a rich, variational distribution is specified in the expanded space of variables we wish to infer, augmented with auxiliary inference variables.
 This can be seen as specifying a non-parametric hierarchical variational model (CITE RANAGNATH 2015)...
@@
 
-The marginal variational distribution over the variables of interest, $q(Z,\eta)$, is itself intractable precluding the analyticity of the variational entropy... therefore we lower bound again...
+In addition to this encoder model, we introduce a framework for training variational auto-encoders and their more complex derived forms.
+Of particular interest are deep VAE architectures that consist of stacked stochastic hidden layers in both the decoding and encoding arms.
+While the depth of the generative model allows for richer latent representations and generative capacity, learning the model parameters via an equivalent deep encoder leads to returns which sharply diminish as more stochastic layers are added.
+We demonstrate that it is the unidirectional nature of the flow of information during training that hampers the learning process, causing the hidden layers closest to the data to fit rigidly to the data, rendering model hidden units further upstream effectively inactive.
+To remedy this pathology we propose an auxiliary inference scheme which combines the bottom-up stochastic message passing of the vanilla encoder with top-down flowing information.
+We show that this Auxiliary Variational Inference (AVI) scheme significantly outperform the now standard stochastic backpropagation procedure for architectures of varying depth.
+The focus of this work is to inform on how to correctly learn deep, expressive generative models, rather than on the design of such generative models themselves.
+Therefore the experiments documented here were designed to showcase the ability of the proposed inference scheme to boost the performance of a \emph{given} model.
+
+Pleasingly the technique of furnishing the variational encoder with an auxiliary model is not a heuristic or hack like e.g.
+the popular tricks of dropout, layer-wise pre-training, gradient clipping, and more recently batch normalisation, rather it is manifest as the result of natural application of the Bayesian variational principle.
+In particular the technique involves further lower bounding the variational lower bound of the model evidence in a way which recovers the auxiliary lower bound considered in RERERENCE SALIMANS.
+
+The marginal variational distribution over the variables of interest, $q(Z)$, is itself intractable precluding analyticity of the variational entropy in the lower bound therefore we lower bound again...
 
@@
 
+\begin{itemize}
+  \item Provide a Bayesian regularisation
+\end{itemize}
 The VAE (2 REFERENCES) was introduced as a framework to train (potentially) deep neural net generative models in a scalable, Bayesian manner.
@@
 
+Burda et al. REFERENCE extended the VAE by including one or more stochastic hidden layers between the data and the top layer latent variables, resulting in the following hierarchical specification:
+%
+\begin{align}
+  p_\theta(\zb_l|\zb_{l+1}) &= \No(\zb_l\mid \mu_l(\zb_{l+1}), \sigma^2_l(\zb_{l+1}))\quad l\in[1,L] \\
+  p_\theta(\zb_L) &= \No(\zb_L|0, I)
+\end{align}
+
+where $\mu_l$ and $\sigma^2_l$ are deterministic mappings consisting of a (often 2 hidden layer) MLP.
+Chaining together these simple dense mappings permits a highly correlated latent dependence structure without forgoing the computational efficiency and scalability of fully factorised models.
+
+The inference network (encoder) used in the IWAE has simply the same architecture as the generative network but with mappings in the reverse direction (figure FIGURE) i.e.
+%
+\begin{align}
+  q_\phi(\zb_1|\yb) &= \No(\zb_1\mid \mu(\yb), \sigma^2(\yb)) \\
+  q_\phi(\zb_l|\zb_{l-1}) &= \No(\zb_l\mid \mu(\zb_{l-1}), \sigma^2(\zb_{l-1}))\quad l\in[2,L].
+\end{align}
+%
+Applying the reparameterisation trick at each stochastic layer now amounts to first sampling from a standard normal $\boldsymbol{\xi}_l\sim\No(0,I)$ and then applying a deterministic mapping as follows:
+%
+\begin{equation}
+  \zb_l(\xi_l,\zb_{l-1},\phi) = \mu(\zb_{l-1},\phi) + \sigma(\zb_{l-1},\phi)\odot\boldsymbol{\xi}_l
+\end{equation}
+%
+Burda et al.
+optimise an evidence lower bound which differs from the traditional free energy in that it cannot be written as the difference between the evidence and a K-L divergence.
+This alternative bound highlights the fundamental similarity between variational inference and importance sampling, and in doing so begins to address the issues of the heavy penalisation under the standard bound of samples which poorly explain the data.
+However this penalisation worsens for layers which are further from the data as they do not receive a learning signal strong enough to overcome the K-L divergence penalisation that encourages variational posteriors towards their model priors.
+%
+\subsection{Auxiliary Variational Inference}\label{subsec:aux}
+
+To overcome the pathologies of the standard inference scheme, we introduce auxiliary variables $\xb$ into the inference model as shown in figure FIGURE, which mediate a direct forward pass sampling route from the data to the top level latent variables.
+Since these variables do not appear in the generative model their effect must be marginalised out of the inference model.
+This marginalisation is however intractable and so we must further lower the bound of equation EQUATION by substituting in the Bayesian identity $q_\phi(\zb_1,\zb_2|\yb) = \frac{q_\phi(\zb_1|\yb)q_\phi(\xb|\yb)q_\phi(\zb_2|\zb_1,\xb)}{q_\phi(\xb|\zb_1,\zb_2,\yb)}$
+%
 \subsection{The Bayesian GP-LVM}\label{subsec:gplvm}
@@
 For the Bayesian GP-LVM the motivation is to be able to tractably integrate over the uncertainty associated with the unknown latent input locations.
-For the GP-LVM-CMF however there are no observations of the GP function values since they do not exist in the generative model, so their purpose here is to act as \emph{variational (pseudo) data} upon which the rest of the function values can be conditioned in the standard GP formulation.
+For the GP-LVM-CMF however there are no observations of the GP function values since they do not exist in the generative model, so their purpose here is to act as \emph{pseudo-data} upon which the rest of the function values can be conditioned via the standard GP formulation.
 
@@
 \begin{equation}
-q(Z) = \frac{q(Z|\f,U,X)q(\f,U,X)}{q(\f,U,X|Z)}
+q(Z) = \frac{q(Z|\fb,U,X)q(\fb,U,X)}{q(\fb,U,X|Z)}
 \end{equation}
 %
-and introducing an auxiliary model $r_\psi(\f,U,X)$, we arrive at the following \emph{auxiliary lower bound}:
+and introducing an auxiliary model $r_\psi(\fb,U,X)$, we arrive at the following \emph{auxiliary lower bound}:
 %
 \begin{align}
-\mathcal{L}_1 &\geq \Ex_{q(Z,\f,U,X)}\big[ \log p(Y|Z)p(Z)r(\f,U,X|Z) \notag \\
-&- \log q(Z|\f,X)q(\f,U,X) \big] \notag \\
-&+ \Ex_{q(Z)}\big[\KL[q(\f,U,X|Z)\Vert r(\f,U,X|Z)]\big] \label{eq:lowerlower} \\
+\mathcal{L}_1 &\geq \Ex_{q(Z,\fb,U,X)}\big[ \log p(Y|Z)p(Z)r(\fb,U,X|Z) \notag \\
+&- \log q(Z|\fb,X)q(\fb,U,X) \big] \notag \\
+&+ \Ex_{q(Z)}\big[\KL[q(\fb,U,X|Z)\Vert r(\fb,U,X|Z)]\big] \label{eq:lowerlower} \\
 &\triangleq \mathcal{L}_{aux}. \label{eq:L_aux}
@@
 %
-To explain this bound, we find it useful to refer to $q(\f,U,X)$ (resp.
-$q(\f,U,X|Z)$) as the \emph{variational prior} (resp.) \emph{variational posterior} distributions over the auxiliary inference variables, the latter of which constitutes a posterior belief over the auxiliary variables having observed a (set of) sample(s), $Z$, generated under the measure $q$ endowed by the inference network.
+To explain this bound, we find it useful to refer to $q(\fb,U,X)$ (resp.
+$q(\fb,U,X|Z)$) as the \emph{variational prior} (resp.) \emph{variational posterior} distributions over the auxiliary inference variables, the latter of which constitutes a posterior belief over the auxiliary variables having observed a (set of) sample(s), $Z$, generated under the measure $q$ endowed by the inference network.
 
@@
 \mathcal{L}_{aux} &= \Ex_{q(Z)}[ \log p(Y|Z) ] \notag \\
-&- \Ex_{q(\f,U,X)}[ \KL[q(Z|\f,U,X)\Vert p(Z)] ]\notag\\
-&- \Ex_{q(Z)}[ \KL[q(\f,U,X)\Vert r(\f,U,X|Z)] ].\label{eq:L_aux_ABC}
+&- \Ex_{q(\fb,U,X)}[ \KL[q(Z|\fb,U,X)\Vert p(Z)] ]\notag\\
+&- \Ex_{q(Z)}[ \KL[q(\fb,U,X)\Vert r(\fb,U,X|Z)] ].\label{eq:L_aux_ABC}
 \end{align}
 %
-As noted by Kingma and Welling, in the case of Gaussian $p(Z)$ and $q(Z|\f,U,X)$, the first K-L divergence can be evaluated analytically and, but the lower lower means we can get a second analytical term...
+As noted by Kingma and Welling, in the case of Gaussian $p(Z)$ and $q(Z|\fb,U,X)$, the first K-L divergence can be evaluated analytically and, but the lower lower means we can get a second analytical term...
 This was also noticed by Tran et al., in their lower lower bound...
 
-Variational compression ... sparse variational GPs ... set $r(\f|U,X,Z):=q(\f|U,X)$ which means that $r(\f,U,X|Z) = q(\f|U,X)r(U,X|Z)$.
+Variational compression ... sparse variational GPs ... set $r(\fb|U,X,Z):=q(\fb|U,X)$ which means that $r(\fb,U,X|Z) = q(\fb|U,X)r(U,X|Z)$.
 Full implications of this??? Reduces configuration space...

### 2016-03-04 parmesan_dev/examples/iw_vae.py

@@ -1,0 +1,447 @@
+# This code implementents a variational autoencoder using importance weighted
+# sampling as described in Burda et al. 2015 "Importance Weighted Autoencoders"
+import theano
+theano.config.floatX = 'float32'
+import matplotlib
+matplotlib.use('Agg')
+import theano.tensor as T
+import numpy as np
+import lasagne
+from parmesan.distributions import log_stdnormal, log_normal2, log_bernoulli
+from parmesan.layers import SampleLayer, NormalizeLayer, ScaleAndShiftLayer
+from parmesan.datasets import load_mnist_realval, load_mnist_binarized
+import matplotlib.pyplot as plt
+import shutil, gzip, os, cPickle, time, math, operator, argparse
+
+filename_script = os.path.basename(os.path.realpath(__file__))
+
+parser = argparse.ArgumentParser()
+parser.add_argument("-dataset", type=str,
+        help="sampled or fixed binarized MNIST, sample|fixed", default="sample")
+parser.add_argument("-eq_samples", type=int,
+        help="number of samples for the expectation over q(z|x)", default=1)
+parser.add_argument("-iw_samples", type=int,
+        help="number of importance weighted samples", default=1)
+parser.add_argument("-lr", type=float,
+        help="learning rate", default=0.001)
+parser.add_argument("-anneal_lr_factor", type=float,
+        help="learning rate annealing factor", default=0.9995)
+parser.add_argument("-anneal_lr_epoch", type=float,
+        help="larning rate annealing start epoch", default=1000)
+parser.add_argument("-batch_norm", type=str,
+        help="batch normalization", default='true')
+parser.add_argument("-outfolder", type=str,
+        help="output folder", default=os.path.join("results", os.path.splitext(filename_script)[0]))
+parser.add_argument("-nonlin_enc", type=str,
+        help="encoder non-linearity", default="rectify")
+parser.add_argument("-nonlin_dec", type=str,
+        help="decoder non-linearity", default="rectify")
+parser.add_argument("-nhidden", type=int,
+        help="number of hidden units in deterministic layers", default=500)
[…]

### 2016-03-04 parmesan_dev/examples/iw_vae_normflow.py

@@ -1,0 +1,488 @@
+# This code implementents a variational autoencoder using importance weighted
+# sampling as described in Burda et al. 2015 "Importance Weighted Autoencoders"
+# and the planar normalizing flow described in Rezende et al. 2015
+# "Variational Inference with Normalizing Flows"
+import theano
+theano.config.floatX = 'float32'
+import matplotlib
+matplotlib.use('Agg')
+import theano.tensor as T
+import numpy as np
+import lasagne
+from parmesan.distributions import log_stdnormal, log_normal2, log_bernoulli
+from parmesan.layers import SampleLayer, NormalizingPlanarFlowLayer, ListIndexLayer, NormalizeLayer, ScaleAndShiftLayer
+from parmesan.datasets import load_mnist_realval, load_mnist_binarized
+import matplotlib.pyplot as plt
+import shutil, gzip, os, cPickle, time, math, operator, argparse
+
+filename_script = os.path.basename(os.path.realpath(__file__))
+
+parser = argparse.ArgumentParser()
+parser.add_argument("-dataset", type=str,
+        help="sampled or fixed binarized MNIST, sample|fixed", default="sample")
+parser.add_argument("-eq_samples", type=int,
+        help="number of samples for the expectation over q(z|x)", default=1)
+parser.add_argument("-iw_samples", type=int,
+        help="number of importance weighted samples", default=1)
+parser.add_argument("-lr", type=float,
+        help="learning rate", default=0.001)
+parser.add_argument("-anneal_lr_factor", type=float,
+        help="learning rate annealing factor", default=0.9995)
+parser.add_argument("-anneal_lr_epoch", type=float,
+        help="larning rate annealing start epoch", default=1000)
+parser.add_argument("-batch_norm", type=str,
+        help="batch normalization", default='true')
+parser.add_argument("-outfolder", type=str,
+        help="output folder", default=os.path.join("results", os.path.splitext(filename_script)[0]))
+parser.add_argument("-nonlin_enc", type=str,
+        help="encoder non-linearity", default="rectify")
+parser.add_argument("-nonlin_dec", type=str,
+        help="decoder non-linearity", default="rectify")
[…]

### 2016-03-04 parmesan_dev/examples/vae_2_sLayer.py

@@ -1,0 +1,251 @@
+# Implements a variational autoencoder as described in Kingma et al. 2013
+# "Auto-Encoding Variational Bayes"
+import theano
+theano.config.floatX = 'float32'
+import theano.tensor as T
+import numpy as np
+import lasagne
+from parmesan.distributions import log_stdnormal, log_normal2, log_bernoulli, kl_normal2_stdnormal, normalEntropy2
+from parmesan.layers import SimpleSampleLayer
+from parmesan.datasets import load_mnist_realval, load_mnist_binarized
+import time, shutil, os
+
+filename_script = os.path.basename(os.path.realpath(__file__))
+
+#settings
+dataset = 'fixed'
+batch_size = 100
+nhidden = 200
+nonlin_enc = T.nnet.softplus
+nonlin_dec = T.nnet.softplus
+latent_size = 100
+latent_ext_size = 50
+analytic_kl_term = True
+lr = 0.0003
+num_epochs = 1000
+results_out = os.path.join("results", os.path.splitext(filename_script)[0])
+
+np.random.seed(1234) # reproducibility
+
+# Setup outputfolder logfile etc.
+if not os.path.exists(results_out):
+    os.makedirs(results_out)
+shutil.copy(os.path.realpath(__file__), os.path.join(results_out, filename_script))
+logfile = os.path.join(results_out, 'logfile.log')
+
+#SYMBOLIC VARS
+sym_x = T.matrix()
+sym_lr = T.scalar('lr')
+
+
+#Helper functions
+def bernoullisample(x):
+    return np.random.binomial(1,x,size=x.shape).astype(theano.config.floatX)
+
+
+### LOAD DATA
+if dataset is 'sample':
+    print "Using real valued MNIST dataset to binomial sample dataset after every epoch "
+    train_x, train_t, valid_x, valid_t, test_x, test_t = load_mnist_realval()
+    del train_t, valid_t, test_t
+    preprocesses_dataset = bernoullisample
+else:
+    print "Using fixed binarized MNIST data"
+    train_x, valid_x, test_x = load_mnist_binarized()
+    preprocesses_dataset = lambda dataset: dataset #just a dummy function
+
+#concatenate train and validation set
+train_x = np.concatenate([train_x, valid_x])
+
+train_x = train_x.astype(theano.config.floatX)
+test_x = test_x.astype(theano.config.floatX)
+
[…]

### 2016-03-04 parmesan_dev/parmesan/distributions.py

@@ -1,0 +1,157 @@
+import math
+import theano.tensor as T
+c = - 0.5 * math.log(2*math.pi)
+
+
+def log_normal(x, mean, sd):
+    """
+    Compute log pdf of a Gaussian distribution with diagonal covariance, at values x.
+
+        .. math:: \log p(x) = \log \mathcal{N}(x; \mu, \sigma^2I)
+
+    Parameters
+    ----------
+    x : Theano tensor
+        Values at which to evaluate pdf.
+    mean : Theano tensor
+        Mean of the Gaussian distribution.
+    sd : Theano tensor
+        Standard deviation of the diagonal covariance Gaussian.
+
+    Returns
+    -------
+    Theano tensor
+        Element-wise log probability, this has to be summed for multi-variate distributions.
+    """
+    return c - T.log(T.abs_(sd)) - (x - mean)**2 / (2 * sd**2)
+
+
+def log_normal2(x, mean, log_var):
+    """
+    Compute log pdf of a Gaussian distribution with diagonal covariance, at values x.
+    Here variance is parameterized in the log domain, which ensures :math:`\sigma > 0`.
+
+        .. math:: \log p(x) = \log \mathcal{N}(x; \mu, \sigma^2I)
+
+    Parameters
+    ----------
+    x : Theano tensor
+        Values at which to evaluate pdf.
+    mean : Theano tensor
+        Mean of the Gaussian distribution.
+    log_var : Theano tensor
+        Log variance of the diagonal covariance Gaussian.
+
+    Returns
+    -------
+    Theano tensor
+        Element-wise log probability, this has to be summed for multi-variate distributions.
+    """
+    return c - log_var/2 - (x - mean)**2 / (2 * T.exp(log_var))
+
+def log_stdnormal(x):
+    """
+    Compute log pdf of a standard Gaussian distribution with zero mean and unit variance, at values x.
+
+        .. math:: \log p(x) = \log \mathcal{N}(x; 0, I)
+
+    Parameters
+    ----------
+    x : Theano tensor
+        Values at which to evaluate pdf.
+
+    Returns
+    -------
+    Theano tensor
+        Element-wise log probability, this has to be summed for multi-variate distributions.
+    """
+    return c - x**2 / 2
+
+
+def log_bernoulli(x, p):
+    """
+    Compute log pdf of a Bernoulli distribution with success probability p, at values x.
+
+        .. math:: \log p(x; p) = \log \mathcal{B}(x; p)
+
+    Parameters
+    ----------
+    x : Theano tensor
[…]

### 2016-03-04 parmesan_dev/parmesan/layers/sample.py

@@ -1,0 +1,137 @@
+import lasagne
+from theano.sandbox.rng_mrg import MRG_RandomStreams as RandomStreams
+import theano.tensor as T
+import theano
+
+class SimpleSampleLayer(lasagne.layers.MergeLayer):
+    """
+    Simple sampling layer drawing a single Monte Carlo sample to approximate
+    E_q [log( p(x,z) / q(z|x) )]. This is the approach described in [KINGMA]_.
+
+    Parameters
+    ----------
+    mu, log_var : :class:`Layer` instances
+        Parameterizing the mean and log(variance) of the distribution to sample
+        from as described in [KINGMA]_. The code assumes that these have the
+        same number of dimensions.
+
+    seed : int
+        seed to random stream
+
+    Methods
+    ----------
+    seed : Helper function to change the random seed after init is called
+
+    References
+    ----------
+        ..  [KINGMA] Kingma, Diederik P., and Max Welling.
+            "Auto-Encoding Variational Bayes."
+            arXiv preprint arXiv:1312.6114 (2013).
+    """
+    def __init__(self, mean, log_var,
+                 seed=lasagne.random.get_rng().randint(1, 2147462579),
+                 **kwargs):
+        super(SimpleSampleLayer, self).__init__([mean, log_var], **kwargs)
+
+        self._srng = RandomStreams(seed)
+
+    def seed(self, seed=lasagne.random.get_rng().randint(1, 2147462579)):
+       self._srng.seed(seed)
+
+    def get_output_shape_for(self, input_shapes):
+        return input_shapes[0]
+
+    def get_output_for(self, input, **kwargs):
+        mu, log_var = input
+        eps = self._srng.normal(mu.shape)
+        z = mu + T.exp(0.5 * log_var) * eps
+        return z
+
+
+class SampleLayer(lasagne.layers.MergeLayer):
+    """
+    Sampling layer supporting importance sampling as described in [BURDA]_ and
+    multiple Monte Carlo samples for the approximation of
+    E_q [log( p(x,z) / q(z|x) )].
+
+    Parameters
+    ----------
+    mu : class:`Layer` instance
+        Parameterizing the mean of the distribution to sample
+        from as described in [BURDA]_.
+
+    log_var : class:`Layer` instance
+        By default assumed to parametrize log(sigma^2) of the distribution to
+        sample from as described in [BURDA]_ which is transformed to sigma using
[…]

### 2016-03-04 parmesan_dev/parmesan/utils.py

@@ -1,0 +1,156 @@
+import numpy as np
+
+
+class ConfusionMatrix:
+    """
+       Simple confusion matrix class
+       row is the true class, column is the predicted class
+    """
+    def __init__(self, n_classes, class_names=None):
+        self.n_classes = n_classes
+        if class_names is None:
+            self.class_names = map(str, range(n_classes))
+        else:
+            self.class_names = class_names
+
+        # find max class_name and pad
+        max_len = max(map(len, self.class_names))
+        self.max_len = max_len
+        for idx, name in enumerate(self.class_names):
+            if len(self.class_names) < max_len:
+                self.class_names[idx] = name + " "*(max_len-len(name))
+
+        self.mat = np.zeros((n_classes, n_classes), dtype='int')
+
+    def __str__(self):
+        # calucate row and column sums
+        col_sum = np.sum(self.mat, axis=1)
+        row_sum = np.sum(self.mat, axis=0)
+
+        s = []
+
+        mat_str = self.mat.__str__()
+        mat_str = mat_str.replace('[', '').replace(']', '').split('\n')
+
+        for idx, row in enumerate(mat_str):
+            if idx == 0:
+                pad = " "
+            else:
+                pad = ""
+            class_name = self.class_names[idx]
+            class_name = " " + class_name + " |"
+            row_str = class_name + pad + row
+            row_str += " |" + str(col_sum[idx])
+            s.append(row_str)
+
+        row_sum = [(self.max_len+4)*" "+" ".join(map(str, row_sum))]
+        hline = [(1+self.max_len)*" "+"-"*len(row_sum[0])]
+
+        s = hline + s + hline + row_sum
+
+        # add linebreaks
+        s_out = [line+'\n' for line in s]
+
+        return "".join(s_out)
+
+    def batchadd(self, y_pred, y_true):
+        assert y_true.shape == y_pred.shape
+        assert len(y_true) == len(y_pred)
+        assert max(y_true) < self.n_classes
+        y_true = y_true.flatten()
+        y_pred = y_pred.flatten()
+        for i in range(len(y_true)):
+                self.mat[y_true[i], y_pred[i]] += 1
+
+    def batchaddmask(self, y_true, y_pred, mask):
+        assert y_true.shape == y_pred.shape
+        assert y_true.shape == mask.shape
[…]

### 2016-03-04 parmesan_dev/utils_RB.py

@@ -1,0 +1,5 @@
+def normalEntropy2(log_var):
+    '''given vector of the diag'''
+    numberOfElements = log_var.shape[0]
+    H = numberOfElements * (0.5 - c) + 0.5*T.sum(log_var)
+    return H
